---
title: "Piece 1 -- Building the Entrant-Composition Panel"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

This document builds `entrant_composition_panel.rds` from raw MINEDUC files and the
shared coauthor panels, in one pass. It is the single source of the panel the analysis
file `1_effect_sae_alloc.Rmd` reads; nothing downstream depends on any intermediate CSV.
Each stage below is documented so the construction can be audited end to end.

**Inputs.** Two kinds.

*Packed raw MINEDUC files* -- three rds files, built once from the public releases by
`00a_pack_raw_data.R` (which documents the sources and the raw-csv handling; repoint its
paths to your own copies to repack). They live in the shared data folder alongside the
canonical panels; this package's `data/` folder works as a fallback location.

- `sae_admissions.rds` -- the SAE offer (A1, processes 2016--2025), application (C1,
  2017--2025) and assignment (D1, 2017--2025) tables. Feed the rollout doors and the
  over-demand tag; A1 seats verified identical to the canonical `sae_data` `VACANTES`
  (2026-08-15).
- `matricula_2011_2023.rds` -- the annual enrolment census, one row per student per
  school per grade code. Identifies entrants: the was-this-student-here-last-year check
  must run against one consistent enrolment register. The graded panel is the wrong
  instrument for that check (it implies enrolment from year-end academic records, and
  its disagreement with the census drifts across years; see Stage 4).
- `alumnos_sep_2012_2023.rds` -- the annual SEP-prioritario (low-income) roster, the
  official beneficiary designation each year (`PRIORITARIO_ALU` from 2016; beneficiary
  roster through 2015). Needed for the low-income share of entrants (the canonical
  `panel_9-12` flag exists only for SAE applicants and cannot tag the pre-SAE baseline).

*Shared coauthor panels* (in `Dropbox/Effect SAE on SUA/data`, = the project `code/data`).

- `panel_performance_2002_2023.rds` -- the canonical graded-enrolment panel: supplies each
  entrant's prior-year GPA and the pre-SAE enrolment for the fill test.
- `panel_ee_2004_2025_homogenized.rds` -- school region.
- `sae_data_2016_2023.rds` -- the (COD\_ENSE, COD\_GRADO) -> SAE level map and declared capacity.

**Output.** One row per always-over-demanded school-grade cell x academic year, columns
`rbd, cod_nivel, yr, region, is_entry, door, n_ent, n_gpa, gpa_z, sh_prio, slack_fail,
door_year_tag`.

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Libraries ---------------------------------------------------------------
library(data.table)

setDTthreads(4); options(warn = 1)
t0 <- proc.time()

# Paths -------------------------------------------------------------------
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
ROOT <- file.path(home, "Dropbox", "Effect of Centralization")
# paths relative to this file (knitting sets the working directory here), so the
# package runs wherever the folder is placed: the panel writes next to the code,
# and data/ is the local fallback for the packed files
DATA_RAW <- "data"
OUTF <- "entrant_composition_panel.rds"
# every data file resolves through one chain: the common shared Dropbox folder,
# then the project's code/data mirror, then this package's own data/ folder
data_path <- function(f) {
  for (p in c(file.path(home, "Dropbox", "Effect SAE on SUA", "data", f),
              file.path(ROOT, "code", "data", f),
              file.path(DATA_RAW, f)))
    if (file.exists(p)) return(p)
  stop(sprintf("input file not found in any data location: %s", f))
}

say <- function(...) { cat(sprintf(...), "\n"); flush.console() }
hdr <- function(x) cat("\n", strrep("=", 92), "\n", x, "\n", strrep("=", 92), "\n", sep = "")

NIVLAB <- c("-1"="PreK","0"="Kinder","1"="1bas","2"="2bas","3"="3bas","4"="4bas","5"="5bas",
            "6"="6bas","7"="7bas","8"="8bas","9"="1med","10"="2med","11"="3med","12"="4med")
ENTRY <- c(-1L, 0L, 1L, 7L, 9L)          # grades a student can ENTER a school at

# SAE tables come from the packed file: one call per table x process year.
# (The packing script 00a_pack_raw_data.R handles the raw-csv quirks once:
# case-resolved headers, Latin-1, course ids forced to character.)
RAW_SAE <- readRDS(data_path("sae_admissions.rds"))
rd_sae <- function(tb, pr) {
  d <- RAW_SAE[[tolower(tb)]][proc == pr]
  if (nrow(d)) copy(d) else NULL
}

# school -> region, and the (COD_ENSE, COD_GRADO) -> SAE level map, from shared panels
ee  <- as.data.table(readRDS(data_path("panel_ee_2004_2025_homogenized.rds")))
REG <- unique(ee[!is.na(cod_reg_rbd), .(rbd, region = cod_reg_rbd)], by = "rbd")
rm(ee); invisible(gc())
SD  <- as.data.table(readRDS(data_path("sae_data_2016_2023.rds")))
SD  <- SD[!is.na(NIVEL) & !is.na(COD_ENSE) & !is.na(COD_GRADO)]
MAP <- unique(SD[, .(COD_ENSE, COD_GRADO, cod_nivel = NIVEL)])
if (nrow(MAP[, .N, by = .(COD_ENSE, COD_GRADO)][N > 1L]))
  stop("a (COD_ENSE, COD_GRADO) pair maps to more than one SAE level; mapping is not a function")
rm(SD); invisible(gc())
```

## Stage 1 -- the region x grade rollout door

SAE phased in by region and, within a region, sometimes by grade. A (region, grade-level)
cell is under SAE in the first process year any school in that region offers that level in
the A1 file. This first year is the cell's *door*, read from the offer files. Reading it
from first appearance in a later file would time the earliest-treated region wrong; the
offer files time Magallanes at process 2016 correctly.

```{r echo=FALSE, warning=FALSE, message=FALSE}
ALL <- list()
# the rollout map is fully determined by processes 2016-2023 (every door year is
# 2016-2020); A1 2024-2025 feed Stage 2's demand window only
for (pr in 2016L:2023L) {
  d <- rd_sae("A1", pr)
  stopifnot("A1 rows must exist for every rollout-map year" = !is.null(d))
  niv <- d$cod_nivel   # 2016's `nivel` column is normalised to cod_nivel in the packed file
  ALL[[as.character(pr)]] <- data.table(proc = pr, rbd = d$rbd, nivel = as.integer(niv))
  rm(d); invisible(gc())
}
A0 <- rbindlist(ALL); rm(ALL); invisible(gc())
A0 <- merge(A0, REG, by = "rbd", all.x = TRUE)
A0 <- A0[!is.na(nivel) & !is.na(region)]
ROLL <- A0[, .(first_proc = min(proc)), by = .(region, nivel)]
stopifnot("rollout map must be one row per region-grade" =
            !anyDuplicated(ROLL, by = c("region", "nivel")),
          "process 2016 must have loaded (it times the Magallanes door)" =
            min(ROLL$first_proc) == 2016L,
          "every door year must be 2016-2020" =
            all(ROLL$first_proc %in% 2016L:2020L))
say("rollout doors: %s region-grade cells | door years %s",
    format(nrow(ROLL), big.mark = ","), paste(sort(unique(ROLL$first_proc)), collapse = ", "))
rm(A0); invisible(gc())
```

## Stage 2 -- effective demand and the over-demand tag

A child *effectively demanded* a school-grade if they made a genuine (non-continuity)
application to it and ranked it at least as high as the school they were finally assigned.
Under deferred acceptance a student is rejected from a school only when it filled with
students it ranked higher, so anyone preferring it to their outcome genuinely lost a
contest for one of its seats. A school-grade is *over-demanded* (`tag_eff`) when effective
demand exceeds the seats offered to the algorithm.

Four constructions matter and are enforced here. Seats are `vacantes`, net of students
continuing inside the school, and continuity application rows never create demand, so
numerator and denominator sit on the same base (they otherwise diverge across the panel and
manufacture a spurious trend). Zero-seat cells are kept in the demand panel and flagged
(`zero_seats`); Stage 3 excludes them from the study set. The placement rank is located
against the unfiltered application file, so every placement is recoverable. And effective
demand is computed by two independent paths and asserted equal cell by cell, with an
economic seat-base check that genuine admits fit inside `vacantes` in at least 98% of
positive-seat cells each year (late-round seat additions make a small overage normal).

```{r echo=FALSE, warning=FALSE, message=FALSE}
CELLS_ED <- list()
for (pr in 2017L:2025L) {
  A <- rd_sae("A1", pr); C <- rd_sae("C1", pr); D <- rd_sae("D1", pr)
  if (is.null(A) || is.null(C) || is.null(D)) { say("process %d: missing table", pr); next }
  if (!"agregada_por_continuidad" %in% names(C)) C[, agregada_por_continuidad := 0L]
  C[is.na(agregada_por_continuidad), agregada_por_continuidad := 0L]
  setnames(C, "preferencia_postulante", "pref")
  stopifnot("preference rank must be positive" = C[, all(pref >= 1L)])

  if (anyDuplicated(A, by = c("rbd", "cod_nivel", "cod_curso")))
    stop(sprintf("process %d: A1 has duplicate rbd-nivel-curso rows; seats would double-count", pr))
  if (!"cupos_totales" %in% names(A)) A[, cupos_totales := NA_integer_]
  SEATS <- A[, .(seats = sum(vacantes, na.rm = TRUE),
                 capacity = sum(cupos_totales, na.rm = TRUE)), by = .(rbd, cod_nivel)]

  # placement rank: the best (lowest) preference the child gave to ANY course at the
  # school-grade they were placed in, located against the UNFILTERED application file
  BEST_ALL <- C[, .(best_pref = min(pref)), by = .(mrun, cod_nivel, rbd)]
  DA <- unique(D[!is.na(rbd_admitido), .(mrun, cod_nivel, rbd = rbd_admitido)],
               by = c("mrun", "cod_nivel"))
  PL <- merge(DA, BEST_ALL, by = c("mrun", "cod_nivel", "rbd"), all.x = TRUE)
  if (PL[, mean(!is.na(best_pref))] < 1)
    stop(sprintf("process %d: a placement rank was not recoverable in the application file", pr))
  PL[, a_rank := as.double(best_pref)]   # double, so Inf survives (into an int column it becomes NA)

  AP <- merge(C[, .(mrun, rbd, cod_nivel, cod_curso, pref, cont = agregada_por_continuidad)],
              SEATS[, .(rbd, cod_nivel)], by = c("rbd", "cod_nivel"))
  AP <- merge(AP, PL[, .(mrun, cod_nivel, a_rank)], by = c("mrun", "cod_nivel"), all.x = TRUE)
  AP[, a_rank := as.double(a_rank)]
  AP[is.na(a_rank), a_rank := Inf]        # unassigned: every genuine listing contested
  stopifnot("a_rank must be double so Inf survives" = is.double(AP$a_rank),
            "no NA may remain in a_rank" = !anyNA(AP$a_rank))
  AP <- merge(AP, DA[, .(mrun, cod_nivel, placed_rbd = rbd)],
              by = c("mrun", "cod_nivel"), all.x = TRUE)
  stopifnot("unassigned must agree across the D1 and C1 routes" =
              AP[is.na(placed_rbd) != !is.finite(a_rank), .N] == 0L)

  # PATH 1: a genuine listing at or above the placement is a contest; distinct children
  AP[, contested := (cont != 1L) & (pref <= a_rank)]
  M1 <- AP[contested == TRUE, .(eff_demand = uniqueN(mrun)), by = .(rbd, cod_nivel)]
  # PATH 2: partition into took-a-seat / spurned / unassigned, then re-sum
  G <- AP[cont != 1L & pref <= a_rank]
  adm  <- G[!is.na(placed_rbd) & placed_rbd == rbd, .(n_adm  = uniqueN(mrun)), by = .(rbd, cod_nivel)]
  spur <- G[!is.na(placed_rbd) & placed_rbd != rbd, .(n_spur = uniqueN(mrun)), by = .(rbd, cod_nivel)]
  unas <- G[is.na(placed_rbd),                      .(n_unas = uniqueN(mrun)), by = .(rbd, cod_nivel)]
  M2 <- Reduce(function(x, y) merge(x, y, by = c("rbd", "cod_nivel"), all = TRUE), list(adm, spur, unas))
  for (v in c("n_adm", "n_spur", "n_unas")) M2[is.na(get(v)), (v) := 0L]
  M2[, eff2 := n_adm + n_spur + n_unas]
  CHK <- merge(M1, M2, by = c("rbd", "cod_nivel"), all = TRUE)
  for (v in c("eff_demand", "eff2", "n_adm", "n_spur", "n_unas")) CHK[is.na(get(v)), (v) := 0L]
  if (nrow(CHK[eff_demand != eff2])) stop("effective demand computed two ways does not agree")

  # economic seat-base check: genuine admits must fit inside vacantes
  SB <- merge(SEATS, CHK[, .(rbd, cod_nivel, n_adm)], by = c("rbd", "cod_nivel"), all.x = TRUE)
  SB[is.na(n_adm), n_adm := 0L]
  if (SB[seats > 0, mean(n_adm <= seats)] < 0.98) stop("seat base mismatched; numerator/denominator not comparable")

  cell <- merge(SEATS, CHK[, .(rbd, cod_nivel, eff_demand)], by = c("rbd", "cod_nivel"), all.x = TRUE)
  cell[is.na(eff_demand), eff_demand := 0L]
  cell[, proc := pr]
  CELLS_ED[[as.character(pr)]] <- cell
  say("process %d | school-grades %s | seat-base OK", pr, format(nrow(SEATS), big.mark = ","))
  rm(A, C, D, AP, G, DA, PL, BEST_ALL, M1, M2, CHK, SB, SEATS); invisible(gc())
}
X <- rbindlist(CELLS_ED); rm(CELLS_ED); invisible(gc())
X <- merge(X, REG, by = "rbd", all.x = TRUE)
X[, is_entry := as.integer(cod_nivel %in% ENTRY)]
X[, level := NIVLAB[as.character(cod_nivel)]]
X <- merge(X, ROLL, by.x = c("region", "cod_nivel"), by.y = c("region", "nivel"), all.x = TRUE)
setnames(X, "first_proc", "door")
X[, `:=`(zero_seats = as.integer(seats == 0),
         tag_eff = as.integer(eff_demand > seats),
         cell = paste(rbd, cod_nivel, sep = "|"))]
stopifnot("no cell may be observed before its door (e < 0)" = X[proc - door < 0L, .N] == 0L)

# first-observed-year tag: was the cell over-demanded at its first observed process
# year? (= the door year for every wave except Magallanes: Stage 2 begins at process
# 2017, the lineage window, so door-2016 cells are first observed at 2017. The 2016
# C1/D1 files ship with the package but are not ingested here.)
X[, snap := min(proc), by = cell]
door_year_cells <- X[proc == snap & tag_eff == 1L, unique(cell)]
say("effective-demand panel: %s school-grade-years | %s tagged | %s door-year-tagged cells",
    format(nrow(X), big.mark = ","), format(X[tag_eff == 1L, .N], big.mark = ","),
    format(length(door_year_cells), big.mark = ","))
rm(RAW_SAE); invisible(gc())   # the SAE tables are not needed past this stage
```

## Stage 3 -- the always-over-demanded study cells

The design tags cells on post-SAE demand and then assumes they were over-demanded before
SAE too. That extrapolation is credible only where over-demand is a stable school trait; a
cell that flips from year to year does not qualify. The study set is therefore the cells
over-demanded in *every* year we observe them. The primary sample keeps positive-seat cells only; zero-seat cells turn away
all demand but have no seats for SAE to reallocate.

```{r echo=FALSE, warning=FALSE, message=FALSE}
XP <- X[zero_seats == 0L]
P  <- XP[, .(years = .N, n_tag = sum(tag_eff)), by = .(cell, rbd, cod_nivel, level, is_entry, region, door)]
P[, always := as.integer(n_tag == years)]
ALW <- P[always == 1L, .(rbd, cod_nivel, door, region, is_entry, cell)]
say("always-over-demanded cells: %s of %s positive-seat cells (%.1f%%)",
    format(nrow(ALW), big.mark = ","), format(nrow(P), big.mark = ","), 100 * nrow(ALW) / nrow(P))
```

## Stage 4 -- entrants and their composition

An *entrant* to a cell in academic year `t` is a student enrolled there in `t` with no
enrolment at that school in `t-1`. Enrolment is read from the annual matricula census,
restricted to the enrolment streams that map to SAE grade levels (regular education; the
same restriction as the validated construction). The `t-1` presence check runs against that
register, so a continuing student is recognised whether or not they have a grade record; a
student whose `t-1` enrolment at the school was in a stream outside the map (adult or
special education) is counted as an entrant. An earlier version of this build ran the
presence check against the graded panel instead. The two registers time enrolment
differently: the census records who is enrolled at the annual snapshot; the graded panel
implies enrolment from year-end academic records. Mid-year movers therefore land in
different classes under the two constructions -- a student who arrived at the school
during `t-1` has a year-end grade record there and reads as continuing under the graded
panel, and reads as an entrant in `t` under the census -- and this disagreement class
grew substantially across the panel window (student-level audit 2026-08-16,
`_phantom_audit.R`; students enrolled with no grade record anywhere are a material class
only at PreK/Kinder/1st, which carry no GPA). The event study of the discrepancy between
the two constructions carries a pre-trend lead structure concentrated at the non-entry
grades and also shifts the post-treatment entry-door estimate (`_delta_did_diag.R`), so
the graded-panel construction imported a time-drifting entrant definition and
manufactured the pre-trend; the census construction is one fixed instrument, uniform
across years and regions, and is the validated original. Each entrant's prior-year
GPA is their grade in `t-1` from the canonical `panel_performance_2002_2023.rds`, z-scored
nationally within year. Entrants at grades 2-12 have a graded prior year and carry a GPA;
1st grade is void for GPA (its prior year, Kinder, is ungraded, so the sole 1st-grade
entrants with a prior grade are repeaters); PreK/Kinder carry no GPA. Each entrant's
low-income status is their SEP-prioritario designation in the enrolment year, from the
annual Alumnos-SEP roster (verified 2026-08-15 against the canonical `panel_9-12` flag:
same concept, ~85% year-aligned agreement, the remainder a snapshot-timing difference
between two official files).

```{r echo=FALSE, warning=FALSE, message=FALSE}
YRS <- 2012:2023   # panel window of record (as the 2026-08-15 build); every prior-GPA
                   # lookup (t-1 <= 2022) sits inside panel_performance; matricula read from 2011
CELLS_T <- ALW[, .(rbd, cod_nivel, door, region, is_entry)]   # tagged cells only

# prior-GPA lookup: national within-year z of PROM_GRAL in [1,7], from the canonical panel
PF <- as.data.table(readRDS(data_path("panel_performance_2002_2023.rds")))
PF <- PF[AGNO %in% (min(YRS) - 1L):2023L, .(MRUN, AGNO, PROM_GRAL)]
PF <- PF[!is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7]
PF[, z := (PROM_GRAL - mean(PROM_GRAL)) / sd(PROM_GRAL), by = AGNO]
GPA <- PF[, .(mrun = as.character(MRUN), agno = AGNO, z)]; rm(PF); invisible(gc())
# ~0.5% of student-years are graded at two schools; collapse to one row per (mrun, agno)
GPA <- GPA[, .(z = mean(z)), by = .(mrun, agno)]
setkey(GPA, mrun, agno)
stopifnot("GPA lookup unique on (mrun, agno)" = !anyDuplicated(GPA, by = c("mrun", "agno")))

# SEP prioritario set for a given year (roster <=2015, PRIORITARIO_ALU flag from 2016;
# the pre-2016 files have no PRIORITARIO_ALU column, packed as all-NA)
SEPD <- readRDS(data_path("alumnos_sep_2012_2023.rds"))
prio_set <- function(y) {
  D <- SEPD[agno == y]
  stopifnot("Alumnos-SEP year must be in the packed file" = nrow(D) > 0)
  if (any(!is.na(D$PRIORITARIO_ALU)))
    as.character(unique(D[PRIORITARIO_ALU == 1L & !is.na(RBD), MRUN]))
  else as.character(unique(D[!is.na(CRITERIO_SEP) & CRITERIO_SEP != 0, MRUN]))
}

# entrants from the matricula census, year by year; prev = (student, school) pairs
# enrolled in t-1 in SAE-mappable streams, so continuing students are recognised
# regardless of grades (a t-1 enrolment in adult/special ed is outside prev)
MATC <- readRDS(data_path("matricula_2011_2023.rds"))
prev <- NULL
RES <- list()
for (t in (min(YRS) - 1L):max(YRS)) {
  M <- MATC[agno == t, .(MRUN, RBD, COD_ENSE, COD_GRADO)]
  stopifnot("matricula year must be in the packed file" = nrow(M) > 0)
  M <- merge(M, MAP, by = c("COD_ENSE", "COD_GRADO"))          # -> cod_nivel; drops adult/special ed
  M <- unique(M[, .(mrun = as.character(MRUN), rbd = RBD, cod_nivel)],
              by = c("mrun", "rbd", "cod_nivel"))
  if (t >= min(YRS) && !is.null(prev)) {
    cur <- merge(M, CELLS_T, by = c("rbd", "cod_nivel"))       # enrolled in a tagged cell
    cur[, in_prev := !is.na(prev[cur, on = c("mrun", "rbd"), which = TRUE])]
    E <- cur[in_prev == FALSE]                                 # entrants
    ps <- prio_set(t)
    E[, prio := as.integer(mrun %chin% ps)]
    E[, agno_prev := t - 1L]
    E[, z := GPA[E, on = .(mrun, agno = agno_prev), x.z]]
    cell <- E[, .(n_ent = .N, sh_prio = mean(prio), gpa_z = mean(z, na.rm = TRUE),
                  n_gpa = sum(!is.na(z))),
              by = .(rbd, cod_nivel, door, region, is_entry)]
    cell[, yr := t]
    RES[[as.character(t)]] <- cell
    say("year %d | entrant cells %s | entrants %s | prio roster %s | %.1f min", t,
        format(nrow(cell), big.mark = ","), format(nrow(E), big.mark = ","),
        format(length(ps), big.mark = ","), (proc.time() - t0)[["elapsed"]] / 60)
    rm(cur, E); invisible(gc())
  }
  prev <- unique(M[, .(mrun, rbd)]); setkey(prev, mrun, rbd)
  rm(M); invisible(gc())
}
PANEL <- rbindlist(RES); rm(RES, prev, GPA, MATC, SEPD); invisible(gc())
say("entrant panel: %s cell-years | %s cells", format(nrow(PANEL), big.mark = ","),
    format(uniqueN(PANEL[, paste(rbd, cod_nivel)]), big.mark = ","))
```

## Stage 5 -- the pre-SAE slack flag

The pre-SAE fill test asks whether an always-over-demanded cell was already full before SAE:
mean enrolment over the four academic years up to the door, as a share of mean enrolment over
the three years after it. A ratio near one is consistent with the cell having been equally
full before; a ratio well below one is either genuine slack (the grade had empty seats and
filled up) or school growth (the whole school got bigger). The two are separated by comparing
each cell's ratio with its own school's ratio over the same window. Cells whose verdict is
genuine slack or borderline are flagged; the analysis reports the estimate with these cells
excluded as one robustness column.

```{r echo=FALSE, warning=FALSE, message=FALSE}
N_PRE <- 4L; N_POST <- 3L
PFe <- as.data.table(readRDS(data_path("panel_performance_2002_2023.rds")))
PFe <- merge(PFe[, .(AGNO, RBD, MRUN, COD_ENSE, COD_GRADO)], MAP, by = c("COD_ENSE", "COD_GRADO"))
E   <- PFe[, .(enr = uniqueN(MRUN)), by = .(rbd = RBD, cod_nivel, AGNO)]
SCH <- E[, .(school_enr = sum(enr)), by = .(rbd, AGNO)]
rm(PFe); invisible(gc())

# cell-level pre/post enrolment ratio for the always cells (graded levels only)
G  <- ALW[cod_nivel >= 1L]
W  <- merge(G, E, by = c("rbd", "cod_nivel"), allow.cartesian = TRUE)
W[, per := fifelse(AGNO <= door, "pre", fifelse(AGNO <= door + N_POST, "post", "later"))]
W  <- W[(per == "pre" & AGNO > door - N_PRE) | per == "post"]
R  <- dcast(W[, .(n = .N, m = mean(enr)), by = .(rbd, cod_nivel, door, per)],
            rbd + cod_nivel + door ~ per, value.var = c("n", "m"))
for (v in c("n_pre", "n_post")) R[is.na(get(v)), (v) := 0L]
R  <- R[n_pre >= 3L & n_post >= 2L & m_post > 0][, ratio := m_pre / m_post]

# school-level ratio over the same window, to net out school growth
K  <- unique(R[, .(rbd, door)])
Ms <- merge(K, SCH, by = "rbd", allow.cartesian = TRUE)
Ms[, per := fifelse(AGNO <= door & AGNO > door - N_PRE, "pre",
             fifelse(AGNO > door & AGNO <= door + N_POST, "post", NA_character_))]
Ms <- Ms[!is.na(per)]
Zs <- dcast(Ms[, .(n = .N, m = mean(school_enr)), by = .(rbd, door, per)],
            rbd + door ~ per, value.var = c("n", "m"))
Zs <- Zs[n_pre >= 3L & n_post >= 2L & m_post > 0][, sch_ratio := m_pre / m_post]
R  <- merge(R, Zs[, .(rbd, door, sch_ratio)], by = c("rbd", "door"))
R[, excess := ratio / sch_ratio]
R[, verdict := fifelse(ratio >= 0.90, "pass",
                fifelse(sch_ratio < 0.90, "fail-growth",
                 fifelse(excess < 0.90, "fail-slack", "fail-borderline")))]
slack_cells <- R[verdict %in% c("fail-slack", "fail-borderline"), paste(rbd, cod_nivel, sep = "|")]
say("testable cells %s | genuine-slack/borderline flagged %s",
    format(nrow(R), big.mark = ","), format(length(slack_cells), big.mark = ","))
```

## Assemble and save

The panel is restricted to the always-over-demanded cells and carries the two robustness
flags (`slack_fail`, `door_year_tag`) as columns, so every column of the analysis table is a
filter on this one file.

```{r echo=FALSE, warning=FALSE, message=FALSE}
OUT <- PANEL[, .(rbd, cod_nivel, yr, region, is_entry, door, n_ent, n_gpa, gpa_z, sh_prio)]
# 1st grade has no graded prior year (Kinder is ungraded), so the only 1st-grade
# entrants with a prior GPA are repeaters -- a non-representative sliver that produces
# single-student cells with extreme z. Void gpa_z there; grades 2-12 all have a real
# graded prior year and stay. n_ent and sh_prio are kept for 1st-grade cells.
# (Deliberate deviation from the 94_quick_look construction, which kept these GPAs;
# user decision 2026-08-15.)
OUT[cod_nivel == 1L, `:=`(gpa_z = NA_real_, n_gpa = 0L)]
OUT[, cell := paste(rbd, cod_nivel, sep = "|")]
OUT[, slack_fail    := cell %chin% slack_cells]
OUT[, door_year_tag := cell %chin% door_year_cells]
OUT[, cell := NULL]
setcolorder(OUT, c("rbd", "cod_nivel", "yr", "region", "is_entry", "door",
                   "n_ent", "n_gpa", "gpa_z", "sh_prio", "slack_fail", "door_year_tag"))
saveRDS(OUT, OUTF)
say("wrote %s", OUTF)
say("panel: %s cell-years | %s cells | years %s",
    format(nrow(OUT), big.mark = ","), format(uniqueN(OUT[, paste(rbd, cod_nivel)]), big.mark = ","),
    paste(range(OUT$yr), collapse = "-"))
say("gpa_z non-missing cell-years: %s | door-year-tagged: %s | slack-flagged: %s",
    format(OUT[!is.na(gpa_z), .N], big.mark = ","),
    format(OUT[door_year_tag == TRUE, .N], big.mark = ","),
    format(OUT[slack_fail == TRUE, .N], big.mark = ","))
say("total elapsed %.1f min", (proc.time() - t0)[["elapsed"]] / 60)
```
