---
title: "Effect of SAE on Allocation - Entrant Composition"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(data.table)
library(did)
# did 2.1.2 is the SE vintage of record for this panel (the region-clustered bootstrap SE
# is version-specific; 2.5.1 reports a much tighter one). Stars come from randomization
# inference, not this SE, but the reported SE must be the pinned vintage -- so halt on any other.
stopifnot("this analysis requires the did package at version 2.1.2" =
            as.character(packageVersion("did")) == "2.1.2")

seed_head <- 20260815
seed_batt <- 20260806
biters    <- 10000
B_ri      <- 9999
B_rob     <- 2999   # RI draws for the robustness-column stars (resolves significance)

# The analysis panel (built by 00_build_entrant_panel.Rmd). A copy next to this file
# wins (a fresh local rebuild); otherwise the shared data folder's copy of record.
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
panel_file <- "entrant_composition_panel.rds"
if (!file.exists(panel_file))
  panel_file <- file.path(home, "Dropbox", "Effect SAE on SUA", "data",
                          "entrant_composition_panel.rds")
stopifnot("panel not found next to the code or in the shared data folder" =
            file.exists(panel_file))
dir.create("tables", showWarnings = FALSE)
dir.create("figures", showWarnings = FALSE)

# helpers: table formatting
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# one CS-DiD fit, simple aggregate, region- or cell-clustered
fit_cs <- function(D, yvar, clustervar = "region", anticipation = 0L,
                   weightsname = NULL, bit = biters) {
  d <- D[!is.na(get(yvar)), c(.(cell_id = cell_id, yr = yr, G = G, region = region,
                                y = get(yvar)),
                              if (!is.null(weightsname)) .(w = get(weightsname)))]
  # cell clustering is did's default (idname = cell_id); region clustering is explicit
  at <- att_gt(yname = "y", tname = "yr", idname = "cell_id", gname = "G",
               data = d, control_group = "notyettreated", panel = TRUE,
               allow_unbalanced_panel = TRUE, base_period = "universal",
               anticipation = anticipation,
               weightsname = if (!is.null(weightsname)) "w",
               bstrap = TRUE, biters = bit,
               clustervars = if (clustervar == "region") "region" else NULL,
               print_details = FALSE)
  # aggregate immediately after the fit so the bootstrap is independent of run order
  s <- aggte(at, type = "simple", na.rm = TRUE)
  list(att = s$overall.att, se = s$overall.se,
       n_cellyears = nrow(d), n_cells = uniqueN(d$cell_id))
}

# treated cells' own pre-door means
pre_mean <- function(D, yvar) D[yr < G & !is.na(get(yvar)), mean(get(yvar))]

# randomization inference: permute the region-to-wave assignment, recompute the
# hand-rolled simple ATT under each draw. wave sizes are held fixed.
prep_ri <- function(D, yvar, minyr = -Inf, wcol = NULL) {
  d <- D[!is.na(get(yvar)) & yr >= minyr]
  d <- d[, c("cell", "region", "is_entry", "yr", yvar, wcol), with = FALSE]
  setnames(d, yvar, "y")
  Wd <- dcast(d, cell + region + is_entry ~ yr, value.var = "y")
  yrs <- as.integer(setdiff(names(Wd), c("cell", "region", "is_entry")))
  w <- NULL
  if (!is.null(wcol)) {
    wc <- unique(d[, .(cell, wv = get(wcol))])
    w <- wc$wv[match(Wd$cell, wc$cell)]
  }
  list(M = as.matrix(Wd[, as.character(yrs), with = FALSE]),
       region = Wd$region, is_entry = Wd$is_entry, yrs = yrs, w = w)
}
# hand-rolled CS simple ATT; base period shifts with anticipation, cells weighted if pp$w set
simple_att_hand <- function(pp, wave_by_region, antic = 0L) {
  G <- wave_by_region[as.character(pp$region)] + (1L - pp$is_entry) + 1L
  yrs <- pp$yrs; M <- pp$M; W <- pp$w
  num <- 0; den <- 0
  for (g in sort(unique(G))) {
    bi <- match(g - 1L - antic, yrs); if (is.na(bi)) next
    tr <- G == g
    for (t in yrs[yrs >= g]) {
      ti <- match(t, yrs)
      ok_t <- tr & !is.na(M[, ti]) & !is.na(M[, bi])
      # control pool respects anticipation, matching did's notyettreated rule (G > t + antic)
      ok_c <- (G > t + antic) & !is.na(M[, ti]) & !is.na(M[, bi])
      if (!any(ok_t) || !any(ok_c)) next
      if (is.null(W)) {
        att <- mean(M[ok_t, ti] - M[ok_t, bi]) - mean(M[ok_c, ti] - M[ok_c, bi]); wt <- sum(ok_t)
      } else {
        att <- weighted.mean(M[ok_t, ti] - M[ok_t, bi], W[ok_t]) -
               weighted.mean(M[ok_c, ti] - M[ok_c, bi], W[ok_c]); wt <- sum(W[ok_t])
      }
      num <- num + wt * att; den <- den + wt
    }
  }
  if (den == 0) NA_real_ else num / den
}
wave_map <- function(D) {
  WM <- unique(D[, .(region, door, is_entry)])
  WM[, wave := door - (1L - is_entry)]
  stopifnot("one wave per region" = WM[, uniqueN(wave), by = region][, all(V1 == 1L)])
  W <- unique(WM[, .(region, wave)])[order(region)]
  setNames(W$wave, as.character(W$region))
}
# event-time lead vector for the exact pre-trend test. Pre-period lead of cohort g
# at t = g+e versus base g-1: controls must be untreated THROUGH the base year
# (G > g-1), matching did's universal-base construction.
ES_LEADS <- -6L:-2L
leads_hand <- function(pp, wv) {
  G <- wv[as.character(pp$region)] + (1L - pp$is_entry) + 1L
  yrs <- pp$yrs; M <- pp$M
  out <- setNames(rep(NA_real_, length(ES_LEADS)), as.character(ES_LEADS))
  for (e in ES_LEADS) {
    num <- 0; den <- 0
    for (g in sort(unique(G))) {
      bi <- match(g - 1L, yrs); ti <- match(g + e, yrs)
      if (is.na(bi) || is.na(ti)) next
      tr <- G == g
      ok_t <- tr & !is.na(M[, ti]) & !is.na(M[, bi])
      # controls: untreated through the base year AND not the treated cohort itself
      # (G > g-1 alone admits G == g, which attenuates every lead toward zero)
      ok_c <- (G > g - 1L) & !tr & !is.na(M[, ti]) & !is.na(M[, bi])
      if (!any(ok_t) || !any(ok_c)) next
      att <- mean(M[ok_t, ti] - M[ok_t, bi]) - mean(M[ok_c, ti] - M[ok_c, bi])
      num <- num + sum(ok_t) * att; den <- den + sum(ok_t)
    }
    out[as.character(e)] <- if (den > 0) num / den else NA_real_
  }
  out
}
# exact joint pre-trend p: family-max studentized leads over region-to-wave permutations.
# check = the did dynamic leads for the same sample; the hand statistic must track them
# or the run halts (guards the control-pool construction against drift from did's)
ri_pretrend <- function(D, check = NULL, B = 2999L) {
  set.seed(seed_head)
  wv <- wave_map(D); pp <- prep_ri(D, "gpa_z")
  obs <- leads_hand(pp, wv)
  if (!is.null(check))
    stopifnot("hand leads must track the did dynamic leads" =
                max(abs(obs[names(check)] - check), na.rm = TRUE) < 0.02)
  regions <- as.integer(names(wv)); waves <- unname(wv)
  P <- matrix(NA_real_, B, length(obs))
  for (b in seq_len(B)) P[b, ] <- leads_hand(pp, setNames(sample(waves), as.character(regions)))
  sdp <- apply(P, 2, sd, na.rm = TRUE)
  fmax <- apply(abs(P) / rep(sdp, each = B), 1, max, na.rm = TRUE)
  (1 + sum(fmax >= max(abs(obs) / sdp, na.rm = TRUE))) / (1 + length(fmax))
}
ri_p <- function(pps, wv, B = B_ri) {
  obs <- vapply(pps, simple_att_hand, numeric(1), wave_by_region = wv)
  regions <- as.integer(names(wv)); waves <- unname(wv)
  P <- matrix(NA_real_, B, length(pps), dimnames = list(NULL, names(pps)))
  for (b in seq_len(B)) {
    wp <- setNames(sample(waves), as.character(regions))
    for (nm in names(pps)) P[b, nm] <- simple_att_hand(pps[[nm]], wp)
  }
  p2 <- vapply(names(pps), function(nm) {
    tp <- P[, nm]; tp <- tp[!is.na(tp)]
    (1 + sum(abs(tp) >= abs(obs[nm]))) / (1 + length(tp))
  }, numeric(1))
  list(obs = obs, p2 = p2)
}
# single-outcome RI p on its own wave map (used per robustness column)
ri_one <- function(pp, wv, antic = 0L, B = B_rob, seed = seed_head) {
  set.seed(seed)
  obs <- simple_att_hand(pp, wv, antic)
  regions <- as.integer(names(wv)); waves <- unname(wv)
  tp <- numeric(B)
  for (b in seq_len(B)) tp[b] <- simple_att_hand(pp, setNames(sample(waves), as.character(regions)), antic)
  tp <- tp[!is.na(tp)]
  (1 + sum(abs(tp) >= abs(obs))) / (1 + length(tp))
}
```

## Data and treatment

The unit is a school-grade cell, one school by one entry grade level, observed each
year. A cell is always over-demanded when, in every year we observe it, the children who
made a genuine non-continuity application ranking the cell at least as high as the school
they were finally assigned outnumber the seats the school offered. Treatment is the year
the region's grade level entered SAE; a cell is treated from the following entering
cohort, so the treatment cohort is `G = door + 1`. The entering class of year `t` is the
students enrolled in the cell in `t` with no enrolment at that school the year before,
identified from the annual matricula enrolment census (the full register, so a continuing
student is recognised whether or not they have a grade record; see the build file). The
two composition outcomes are the entering class's mean prior-year GPA,
standardized nationally within year, and the share of the entering class carrying
SEP-prioritario status, the national low-income tag. The analysis reads one panel,
`entrant_composition_panel.rds`, built and documented in `00_build_entrant_panel.Rmd`.

```{r echo=FALSE, warning=FALSE, message=FALSE}
CP <- as.data.table(readRDS(panel_file))
CP[, cell := paste(rbd, cod_nivel, sep = "|")]
CP[, cell_id := as.integer(factor(cell))]
CP[, G := door + 1L]
CP[, w_ent := mean(n_ent), by = cell_id]
stopifnot("every cell must carry a region" = !anyNA(CP$region),
          "every cell must carry a treatment door" = !anyNA(CP$door))
wv_real <- wave_map(CP)
```

## Headline: effect on entering-class composition

Callaway and Sant'Anna group-time ATT with not-yet-treated cells as controls, a
universal base period, aggregated to one average effect. Standard errors are from a
region-clustered multiplier bootstrap. Significance is from within-rollout randomization
inference: 9,999 permutations of the region-to-wave assignment holding the wave-size
multiset fixed, with the two outcomes sharing one permutation stream. The
share-prioritario column uses the 2016 and later subpanel, because the SEP record
definition changes at calendar 2016.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(seed_head)
g_reg <- fit_cs(CP,              "gpa_z",   "region")
p_reg <- fit_cs(CP[yr >= 2016L], "sh_prio", "region")

# the randomization stream is seeded independently of the did bootstrap, so the
# stars depend only on the permutations -- not on how many draws any fit consumed
set.seed(seed_head)
ri <- ri_p(list(gpa_z   = prep_ri(CP, "gpa_z"),
                sh_prio = prep_ri(CP, "sh_prio", minyr = 2016L)), wv_real)
# two-path check: the hand-rolled statistic must track the did estimand
# (tolerance scaled per outcome: sh_prio is a share whose effect is an order smaller)
stopifnot("hand statistic must track the did point" =
            abs(ri$obs["gpa_z"] - g_reg$att) < 0.02 && abs(ri$obs["sh_prio"] - p_reg$att) < 0.005)

att  <- c(g_reg$att, p_reg$att)
se   <- c(g_reg$se,  p_reg$se)
pri  <- ri$p2[c("gpa_z", "sh_prio")]
prem <- c(pre_mean(CP, "gpa_z"), pre_mean(CP[yr >= 2016L], "sh_prio"))
nobs <- c(g_reg$n_cellyears, p_reg$n_cellyears)

tab <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect of SAE on Entering-Class Composition}",
  "\\label{tab: alloc headline}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lcc}\n      \\toprule\n",
  "                   & Entrant GPA (z) & Share prioritario\\\\\n",
  "                   & (1) & (2)\\\\\n      \\midrule\n",
  row_tex("Treated", paste0(fmt(att), star(pri))),
  row_tex("", paste0("(", fmt(se), ")")),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(prem)),
  row_tex("Observations", format(nobs, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: Callaway--Sant'Anna ATT, not-yet-treated ",
  "controls. Region-clustered multiplier-bootstrap standard errors in parentheses (did ",
  "2.1.2 vintage). Stars from randomization inference over the region-to-wave assignment ",
  "(9,999 permutations): $^{***}$p$<$0.01, $^{**}$p$<$0.05, $^{*}$p$<$0.1. Share ",
  "prioritario uses the 2016+ subpanel (SEP definition break).}\n\\end{table}\n")
writeLines(tab, "tables/alloc_headline.tex")
cat(tab)
```

## Robustness

One specification per column, same outcome as column (1) of the headline; the Baseline
column repeats the headline fit so every check reads against it. *Drop Santiago*: the
Metropolitan region is the last adopter and therefore the dominant not-yet-treated
control in the late identifying comparisons, and it holds a third of the country's
students; this column removes it entirely, treated and control alike, so the estimate
rests on the other fifteen regions — anything specific to Santiago in 2019--2020 cannot
drive a result that survives here. *$n_{gpa}\geq 5$*: a cell-year's outcome is the mean
prior GPA of its graded entrants, and a mean over one or two students is noise; this
column keeps only cell-years with at least five graded entrants, so thin cells cannot
carry the unweighted estimator. *Slack excl.*: the over-demand tag is measured in SAE
years and extrapolated to the pre-SAE period; Stage 5 of the build compares each cell's
pre-door enrolment with its post-door fill and flags cells that look like they had empty
seats before SAE. A cell with slack had no selection to lose, so this column drops the
flagged cells and defends the extrapolation directly. *Entrant-wtd.*: the baseline
weights every cell equally (the class-average estimand); this column weights cells by
their average entrant count, moving the estimand to the average entering student, so the
result is shown to hold where the students actually are. *Anticip.\ = 1*: the base year
is the door year, when SAE was already announced, so the final pre-SAE intake could
itself be distorted by anticipation; this column moves the base one year earlier, before
any announcement behaviour. Region-clustered bootstrap standard errors are in
parentheses; stars are from randomization inference on each column's own specification.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(seed_batt)
# label, panel, weight column, anticipation, outcome column
specs <- list(
  list(lab = "Baseline",         D = CP,                       w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "Drop Santiago",    D = CP[region != 13L],        w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "$n_{gpa}\\geq 5$", D = CP[n_gpa >= 5L],          w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "Slack excl.",      D = CP[slack_fail == FALSE],  w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "Entrant-wtd.",     D = CP,                       w = "w_ent", a = 0L, y = "gpa_z"),
  list(lab = "Anticip.\\ = 1",   D = CP,                       w = NULL,    a = 1L, y = "gpa_z"))

fitb   <- lapply(specs, function(s) fit_cs(s$D, s$y, "region", anticipation = s$a,
                                           weightsname = s$w, bit = 999))
fitb[[1]] <- g_reg   # Baseline repeats the headline spec: reuse the headline fit (biters 10000)
b_att  <- vapply(fitb, `[[`, 0,  "att")
b_se   <- vapply(fitb, `[[`, 0,  "se")
b_nobs <- vapply(fitb, `[[`, 0L, "n_cellyears")
b_pre  <- vapply(specs, function(s) pre_mean(s$D, s$y), 0)
# stars from randomization inference, one p per column (statistic matches each spec)
b_pri  <- vapply(specs, function(s)
  ri_one(prep_ri(s$D, s$y, wcol = s$w), wave_map(s$D), antic = s$a), 0)
b_pri[1] <- ri$p2[["gpa_z"]]   # and the headline RI p (9,999 shared-stream draws)

labs <- vapply(specs, `[[`, "", "lab"); k <- length(labs)
tab_b <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Robustness: Effect on Entrant Prior-GPA}",
  "\\label{tab: alloc robustness}\n",
  "   \\centerline{\\scalebox{0.8}{\\begin{tabular}{l", strrep("c", k), "}\n      \\toprule\n",
  "                   & ", paste(labs, collapse = " & "), "\\\\\n",
  "                   & ", paste0("(", seq_len(k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
  row_tex("Treated", paste0(fmt(b_att), star(b_pri))),
  row_tex("", paste0("(", fmt(b_se), ")")),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(b_pre)),
  row_tex("Observations", format(b_nobs, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: one specification change per column; Baseline ",
  "repeats the headline fit. Region-clustered multiplier-bootstrap standard errors in ",
  "parentheses; stars from randomization inference on each column's own specification: ",
  "$^{***}$p$<$0.01, $^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab_b, "tables/alloc_robustness.tex")
cat(tab_b)
```

## Parallel trends: event studies

Dynamic ATT on entrant prior-GPA, the same configuration as the headline. The reference
period is the year before treatment. The full dynamic path is printed with the figure;
the figure window ends at event time +1, where every point is identified by several
rollout waves (the +2 estimate rests on the earliest adopters against the final wave
alone, and no not-yet-treated control remains at +3). Two formal statistics are printed
under the figure. The exact randomization joint test of the leads is the inference of
record: the region-to-wave permutation, family-max studentized statistic, with the
hand-computed leads asserted against the estimator's dynamic path at every run. The
estimator's own Wald pre-test is printed beside it; its rejection is driven by the 2017
cohort's disaggregated cells, which are identified by a single region (Magallanes)
against its door-year base and carry single-region t-statistics of the kind this
project's inference standard rejects; the identical rejection appears on the superseded
build's panel (`_pretest_diag.R`, 2026-08-16), and the anticipation-1 robustness column
addresses the door-year base directly.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
es_one <- function(D, fname) {
  set.seed(seed_head)
  d <- D[!is.na(gpa_z), .(cell_id, yr, G, region, y = gpa_z)]
  at <- att_gt(yname = "y", tname = "yr", idname = "cell_id", gname = "G", data = d,
               control_group = "notyettreated", panel = TRUE, allow_unbalanced_panel = TRUE,
               base_period = "universal", anticipation = 0L, bstrap = TRUE, biters = biters,
               clustervars = "region", print_details = FALSE)
  ed <- aggte(at, type = "dynamic", na.rm = TRUE)
  es <- data.table(e = ed$egt, att = ed$att.egt, se = ed$se.egt)
  # did's pre-test of parallel trends: joint Wald on the pre-treatment group-time
  # ATTs with their full covariance (the leads share a base period, so a
  # sum-of-squared-t approximation would get the size wrong)
  Wstat <- as.numeric(at$W); Wp <- as.numeric(at$Wpval)
  # full dynamic path printed below; the FIGURE window ends at e = +1, where every
  # point is identified by several waves. e = +2 is identified only by the earliest
  # adopters against the final wave (SE ~5x the +1 value) and e = +3 has no
  # not-yet-treated control left; both stay in the printed table.
  print(es[is.finite(att), .(e, att = round(att, 4), se = round(se, 4))])
  leads_did <- es[e %in% -6:-2, setNames(att, as.character(e))]
  es <- es[e >= -6 & e <= 1 & is.finite(att) & is.finite(se)]
  lo <- es$att - 1.96 * es$se
  hi <- es$att + 1.96 * es$se
  pdf(fname, width = 7.2, height = 4.8, family = "serif")
  par(mar = c(4.6, 5.0, 1.2, 1.2), las = 1, mgp = c(3.1, 0.8, 0),
      cex.axis = 1.05, cex.lab = 1.2)
  plot(es$e, es$att, type = "n",
       ylim = range(pretty(c(lo, hi))), xlim = c(min(es$e) - 0.4, max(es$e) + 0.4),
       xlab = "Event time (years relative to SAE adoption)",
       ylab = "Effect on entrant prior-GPA (z)", xaxt = "n")
  axis(1, at = min(es$e):max(es$e))
  abline(h = 0, col = "black", lwd = 1)
  abline(v = -0.5, col = "black", lty = 2)
  arrows(es$e, lo, es$e, hi, length = 0.045, angle = 90, code = 3,
         col = "black", lwd = 1.3)
  points(es$e, es$att, pch = 21, bg = "white", col = "black", cex = 1.25, lwd = 1.4)
  points(-1, 0, pch = 21, bg = "white", col = "black", cex = 1.25, lwd = 1.4)  # reference period
  invisible(dev.off())
  list(W = Wstat, p = Wp, leads = leads_did)
}
j_all  <- es_one(CP, "figures/alloc_event_study.pdf")
rp_all <- ri_pretrend(CP, check = j_all$leads)
cat(sprintf(paste0("`alloc_event_study.pdf`: exact RI joint lead test $p = %.2f$; ",
                   "Wald pre-test $W = %.1f$, $p = %.2f$.\n"),
            rp_all, j_all$W, j_all$p))
```

## Interpretation guards

Two checks that the headline is a change in who enters, with the measurement itself
standing still. The first outcome is the number of entrants per cell. Seats at
over-demanded schools are set by the school and were full before and after SAE, so the
reform should change the identity of the entrants and leave their count alone; a
door-aligned move in the count would mean the entrant definition or the seat base shifts
with treatment. The second outcome is the share of a cell's entrants who carry a
prior-year GPA. The GPA outcome can only be computed for students with a graded prior
year inside Chile, and a student who just arrived in the country has none. Migration
into Chile rose sharply over 2017--2019; if it tracked the rollout, the entering-class
GPA could move because the measurable share of the class moved. A flat share closes
that door. The share is computed for grades 2--12 (the grades where a prior GPA can
exist). Same estimator, standard errors, and randomization inference as the headline.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
CP[, sh_gpa := fifelse(n_ent > 0L, n_gpa / n_ent, NA_real_)]
CPg <- CP[cod_nivel >= 2L]
set.seed(seed_head)
g_ent <- fit_cs(CP,  "n_ent",  "region")
g_cov <- fit_cs(CPg, "sh_gpa", "region")
p_ent <- ri_one(prep_ri(CP,  "n_ent"),  wave_map(CP))
p_cov <- ri_one(prep_ri(CPg, "sh_gpa"), wave_map(CPg))

att_g  <- c(g_ent$att, g_cov$att)
se_g   <- c(g_ent$se,  g_cov$se)
pri_g  <- c(p_ent, p_cov)
prem_g <- c(pre_mean(CP, "n_ent"), pre_mean(CPg, "sh_gpa"))
nobs_g <- c(g_ent$n_cellyears, g_cov$n_cellyears)

tab_g <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Interpretation Guards: Entrant Counts and GPA Coverage}",
  "\\label{tab: alloc guards}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lcc}\n      \\toprule\n",
  "                   & Entrants per cell & Share with prior GPA\\\\\n",
  "                   & (1) & (2)\\\\\n      \\midrule\n",
  row_tex("Treated", paste0(fmt(att_g), star(pri_g))),
  row_tex("", paste0("(", fmt(se_g), ")")),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(prem_g)),
  row_tex("Observations", format(nobs_g, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: Callaway--Sant'Anna ATT, not-yet-treated ",
  "controls, as in the headline. Column (2) covers grades 2--12, where a prior-year GPA ",
  "can exist. Region-clustered multiplier-bootstrap standard errors in parentheses; ",
  "stars from randomization inference: $^{***}$p$<$0.01, $^{**}$p$<$0.05, ",
  "$^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab_g, "tables/alloc_guards.tex")
cat(tab_g)
```
