---
title: "Effect of SAE on Allocation - Entrant Composition"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(data.table)
library(did)
library(parallel)
library(tools)
# did 2.1.2 is the SE vintage of record for this panel; stars come from randomization
# inference, but the reported SE must be the pinned vintage -- halt on any other.
stopifnot("this analysis requires the did package at version 2.1.2" =
            as.character(packageVersion("did")) == "2.1.2",
          "R 4.5.1 required (default RNG; seeds reproduce the permutation draws)" =
            paste(R.version$major, R.version$minor, sep = ".") == "4.5.1")

seed_head <- 20260815   # headline, robustness, and guard permutation streams
seed_batt <- 20260806   # robustness SE fits
seed_es   <- 20260830   # share-prioritario event study and its lead test
biters    <- 10000
B_ri      <- 9999L      # permutation draws, headline
B_rob     <- 2999L      # permutation draws per robustness column and per guard
N_CORES   <- min(16L, max(1L, parallel::detectCores() - 2L))
                        # permutation scoring only; assignments are drawn sequentially
                        # in the main process, so results are worker-count invariant

# The analysis panel (built by Piece 1's 00_build_entrant_panel.Rmd). A copy next to
# this file wins; otherwise the shared data folder's copy of record. MD5-pinned.
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
panel_file <- "entrant_composition_panel.rds"
if (!file.exists(panel_file))
  panel_file <- file.path(home, "Dropbox", "Effect of Centralization",
                          "Updated Paper Spine", "Piece 1", "entrant_composition_panel.rds")
if (!file.exists(panel_file))
  panel_file <- file.path(home, "Dropbox", "Effect SAE on SUA", "data",
                          "entrant_composition_panel.rds")
stopifnot("panel not found" = file.exists(panel_file),
          "panel MD5 must match the Piece 1 record" =
            unname(md5sum(panel_file)) == "8ccec73b5955b723485195d5085b669d")
dir.create("tables", showWarnings = FALSE)
dir.create("figures", showWarnings = FALSE)

# helpers: table formatting
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# one CS-DiD fit, simple aggregate, region-clustered
fit_cs <- function(D, yvar, clustervar = "region", anticipation = 0L,
                   weightsname = NULL, bit = biters) {
  d <- D[!is.na(get(yvar)), c(.(cell_id = cell_id, yr = yr, G = G, region = region,
                                y = get(yvar)),
                              if (!is.null(weightsname)) .(w = get(weightsname)))]
  at <- att_gt(yname = "y", tname = "yr", idname = "cell_id", gname = "G",
               data = d, control_group = "notyettreated", panel = TRUE,
               allow_unbalanced_panel = TRUE, base_period = "universal",
               anticipation = anticipation,
               weightsname = if (!is.null(weightsname)) "w",
               bstrap = TRUE, biters = bit,
               clustervars = if (clustervar == "region") "region" else NULL,
               print_details = FALSE)
  s <- aggte(at, type = "simple", na.rm = TRUE)
  list(att = s$overall.att, se = s$overall.se,
       n_cellyears = nrow(d), n_cells = uniqueN(d$cell_id))
}
pre_mean <- function(D, yvar) D[yr < G & !is.na(get(yvar)), mean(get(yvar))]

# randomization machinery
wave_map <- function(D) {
  WM <- unique(D[, .(region, door, is_entry)])
  WM[, wave := door - (1L - is_entry)]
  stopifnot("one wave per region" = WM[, uniqueN(wave), by = region][, all(V1 == 1L)])
  W <- unique(WM[, .(region, wave)])[order(region)]
  setNames(W$wave, as.character(W$region))
}
prep_ri <- function(D, yvar, minyr = -Inf, wcol = NULL) {
  d <- D[!is.na(get(yvar)) & yr >= minyr]
  d <- d[, c("cell", "region", "is_entry", "yr", yvar, wcol), with = FALSE]
  setnames(d, yvar, "y")
  Wd <- dcast(d, cell + region + is_entry ~ yr, value.var = "y")
  yrs <- as.integer(setdiff(names(Wd), c("cell", "region", "is_entry")))
  w <- NULL
  if (!is.null(wcol)) {
    wc <- unique(d[, .(cell, wv = get(wcol))])
    w <- wc$wv[match(Wd$cell, wc$cell)]
  }
  list(M = as.matrix(Wd[, as.character(yrs), with = FALSE]),
       region = Wd$region, is_entry = Wd$is_entry, yrs = yrs, w = w)
}
simple_att_hand <- function(pp, wave_by_region, antic = 0L) {
  G <- wave_by_region[as.character(pp$region)] + (1L - pp$is_entry) + 1L
  yrs <- pp$yrs; M <- pp$M; W <- pp$w
  num <- 0; den <- 0
  for (g in sort(unique(G))) {
    bi <- match(g - 1L - antic, yrs); if (is.na(bi)) next
    tr <- G == g
    for (t in yrs[yrs >= g]) {
      ti <- match(t, yrs)
      ok_t <- tr & !is.na(M[, ti]) & !is.na(M[, bi])
      ok_c <- (G > t + antic) & !is.na(M[, ti]) & !is.na(M[, bi])
      if (!any(ok_t) || !any(ok_c)) next
      if (is.null(W)) {
        att <- mean(M[ok_t, ti] - M[ok_t, bi]) - mean(M[ok_c, ti] - M[ok_c, bi]); wt <- sum(ok_t)
      } else {
        att <- weighted.mean(M[ok_t, ti] - M[ok_t, bi], W[ok_t]) -
               weighted.mean(M[ok_c, ti] - M[ok_c, bi], W[ok_c]); wt <- sum(W[ok_t])
      }
      num <- num + wt * att; den <- den + wt
    }
  }
  if (den == 0) NA_real_ else num / den
}
ES_LEADS <- -6L:-2L
leads_hand <- function(pp, wv) {
  G <- wv[as.character(pp$region)] + (1L - pp$is_entry) + 1L
  yrs <- pp$yrs; M <- pp$M
  out <- setNames(rep(NA_real_, length(ES_LEADS)), as.character(ES_LEADS))
  for (e in ES_LEADS) {
    num <- 0; den <- 0
    for (g in sort(unique(G))) {
      bi <- match(g - 1L, yrs); ti <- match(g + e, yrs)
      if (is.na(bi) || is.na(ti)) next
      tr <- G == g
      ok_t <- tr & !is.na(M[, ti]) & !is.na(M[, bi])
      ok_c <- (G > g - 1L) & !tr & !is.na(M[, ti]) & !is.na(M[, bi])
      if (!any(ok_t) || !any(ok_c)) next
      att <- mean(M[ok_t, ti] - M[ok_t, bi]) - mean(M[ok_c, ti] - M[ok_c, bi])
      num <- num + sum(ok_t) * att; den <- den + sum(ok_t)
    }
    out[as.character(e)] <- if (den > 0) num / den else NA_real_
  }
  out
}
# exact joint pre-trend p: family-max studentized leads over region-to-wave draws
ri_pretrend_y <- function(D, yvar, seed, check = NULL, B = B_rob) {
  set.seed(seed)
  wv <- wave_map(D); pp <- prep_ri(D, yvar)
  obs <- leads_hand(pp, wv)
  if (!is.null(check))
    stopifnot("hand leads must track the did dynamic leads" =
                max(abs(obs[names(check)] - check), na.rm = TRUE) < 0.02)
  regions <- as.integer(names(wv)); waves <- unname(wv)
  P <- matrix(NA_real_, B, length(obs))
  for (b in seq_len(B)) P[b, ] <- leads_hand(pp, setNames(sample(waves), as.character(regions)))
  sdp <- apply(P, 2, sd, na.rm = TRUE)
  fmax <- apply(abs(P) / rep(sdp, each = B), 1, max, na.rm = TRUE)
  (1 + sum(fmax >= max(abs(obs) / sdp, na.rm = TRUE))) / (1 + length(fmax))
}

# the test statistic: the simple ATT scored against its own delete-one-region
# jackknife scale, at the observed and at every permuted wave map
jk_stat <- function(pp, w, regs, idx, antic = 0L) {
  th <- simple_att_hand(pp, w, antic)
  loo <- vapply(seq_along(regs), function(i) {
    k <- idx[[i]]
    simple_att_hand(list(M = pp$M[k, , drop = FALSE], region = pp$region[k],
                         is_entry = pp$is_entry[k], yrs = pp$yrs,
                         w = if (is.null(pp$w)) NULL else pp$w[k]), w, antic)
  }, numeric(1))
  ok <- !is.na(loo); Rp <- sum(ok)
  v  <- if (Rp >= 2L) (Rp - 1) / Rp * sum((loo[ok] - mean(loo[ok]))^2) else NA_real_
  c(th = th, t = th / sqrt(v))
}
# one permutation test: seed, then B sequential draws (the statistic consumes no
# randomness); only the per-draw scoring is distributed to workers
score_stream <- function(pp, wv, seed, B, antic = 0L) {
  regs <- sort(unique(pp$region))
  idx  <- lapply(regs, function(r) which(pp$region != r))
  set.seed(seed)
  obs <- jk_stat(pp, wv, regs, idx, antic)
  regions <- as.integer(names(wv)); waves <- unname(wv)
  WPS <- vector("list", B)
  for (b in seq_len(B)) WPS[[b]] <- setNames(sample(waves), as.character(regions))
  RES <- do.call(rbind, parLapplyLB(cl, WPS, function(w) jk_stat(pp, w, regs, idx, antic),
                                    chunk.size = 64L))
  TT <- RES[, "t"]
  p  <- (1 + sum(abs(TT[is.finite(TT)]) >= abs(obs["t"]))) / (1 + sum(is.finite(TT)))
  stopifnot("non-finite draws must stay under 5%" = mean(!is.finite(TT)) <= 0.05)
  list(th = unname(obs["th"]), t = unname(obs["t"]), p = p,
       n_dropped = sum(!is.finite(TT)))
}
cl <- makeCluster(N_CORES)
clusterExport(cl, c("simple_att_hand", "jk_stat"))
```

## Data and treatment

The unit is a school-grade cell, one school by one grade level, observed each year
from 2012 to 2023. A cell is always over-demanded when, in every year we observe
it, the children who made a genuine non-continuity application ranking the cell at
least as high as their final placement outnumber the seats the school offered.
Treatment is the year the region's grade level entered SAE; a cell is treated from
the following entering cohort. The entering class of year $t$ is the students
enrolled in the cell in $t$ with no enrolment at that school the year before,
identified from the annual matricula enrolment census. The two composition
outcomes are the entering class's mean prior-year GPA, standardized nationally
within year, and share prioritario, the share of the entering class carrying the
government's low-income designation, each entrant tagged by that year's published
rule. The analysis reads one panel, `entrant_composition_panel.rds`, built and
documented in the Piece 1 package.

```{r echo=FALSE, warning=FALSE, message=FALSE}
CP <- as.data.table(readRDS(panel_file))
CP[, cell := paste(rbd, cod_nivel, sep = "|")]
CP[, cell_id := as.integer(factor(cell))]
CP[, G := door + 1L]
CP[, w_ent := mean(n_ent), by = cell_id]
stopifnot("every cell must carry a region" = !anyNA(CP$region),
          "every cell must carry a treatment door" = !anyNA(CP$door))
wv_real <- wave_map(CP)
stopifnot("wave multiset must be {1,4,10,1}" =
            identical(as.integer(sort(as.vector(table(wv_real)))), c(1L, 1L, 4L, 10L)))
```

## Inference

Significance in every table comes from randomization inference over the
region-to-wave assignment: permute which regions received which rollout wave,
holding the wave-size multiset fixed, with 9,999 draws for the headline and 2,999
per robustness column and guard. The test statistic is the hand-computed simple
ATT divided by its own delete-one-region jackknife standard error, computed at the
observed assignment and at every permuted one; the two-sided p compares the
observed studentized value with the permuted ones. Studentized randomization
statistics are the standard treatment when the units being reassigned differ
greatly in size, as they do here with the Metropolitan region holding about a
third of the cells: Young (2019, QJE) adopts the randomization-t as the default
statistic, and MacKinnon and Webb (2020, Journal of Econometrics) show for
difference-in-differences that randomization inference on the raw coefficient is
distorted by heterogeneous cluster sizes while the studentized version holds its
size in their designs. The joint pre-trend test applies the same principle,
scaling each lead by its permutation spread. Exactness under the sharp null is
unaffected by the choice of statistic. Region-clustered multiplier-bootstrap
standard errors (did 2.1.2 vintage) are reported for scale.

## Headline: effect on entering-class composition

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(seed_head)
g_reg <- fit_cs(CP, "gpa_z",   "region")
p_reg <- fit_cs(CP, "sh_prio", "region")

pp_g <- prep_ri(CP, "gpa_z")
pp_p <- prep_ri(CP, "sh_prio")
set.seed(seed_head)   # one shared permutation stream scores both outcomes
regs16 <- sort(unique(pp_g$region))
idx_g <- lapply(regs16, function(r) which(pp_g$region != r))
idx_p <- lapply(regs16, function(r) which(pp_p$region != r))
obs_g <- jk_stat(pp_g, wv_real, regs16, idx_g)
obs_p <- jk_stat(pp_p, wv_real, regs16, idx_p)
stopifnot("hand gpa statistic must track the did point"  = abs(obs_g["th"] - g_reg$att) < 0.02,
          "hand prio statistic must track the did point" = abs(obs_p["th"] - p_reg$att) < 0.005)
regions <- as.integer(names(wv_real)); waves <- unname(wv_real)
WPS <- vector("list", B_ri)
for (b in seq_len(B_ri)) WPS[[b]] <- setNames(sample(waves), as.character(regions))
clusterExport(cl, c("pp_g", "pp_p", "regs16", "idx_g", "idx_p"))
RES <- do.call(rbind, parLapplyLB(cl, WPS, function(w)
  c(jk_stat(pp_g, w, regs16, idx_g), jk_stat(pp_p, w, regs16, idx_p)), chunk.size = 64L))
stopifnot("no draw may be non-finite on the headline stream" = all(is.finite(RES)))
p_g <- (1 + sum(abs(RES[, 2]) >= abs(obs_g["t"]))) / (1 + B_ri)
p_p <- (1 + sum(abs(RES[, 4]) >= abs(obs_p["t"]))) / (1 + B_ri)
cat(sprintf("Randomization test: gpa T %+.2f, p %.4f; share prioritario T %+.2f, p %.4f.\n\n",
            obs_g["t"], p_g, obs_p["t"], p_p))

att  <- c(g_reg$att, p_reg$att)
se   <- c(g_reg$se,  p_reg$se)
pri  <- c(p_g, p_p)
prem <- c(pre_mean(CP, "gpa_z"), pre_mean(CP, "sh_prio"))
nobs <- c(g_reg$n_cellyears, p_reg$n_cellyears)
note_core <- paste0(
  "each draw scored against its own delete-one-region jackknife scale following ",
  "Young (2019) and MacKinnon and Webb (2020): ",
  "$^{***}$p$<$0.01, $^{**}$p$<$0.05, $^{*}$p$<$0.1.")
note_head <- paste0(
  "Stars from randomization inference over the region-to-wave assignment (9,999 ",
  "permutations), ", note_core)
note_rob <- paste0(
  "Stars from randomization inference over the region-to-wave assignment (9,999 ",
  "permutations for the Baseline column, 2,999 per other column), ", note_core)
note_gua <- paste0(
  "Randomization inference over the region-to-wave assignment (2,999 permutations ",
  "per column), ", note_core)
tab <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect of SAE on Entering-Class Composition}",
  "\\label{tab: alloc headline}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lcc}\n      \\toprule\n",
  "                   & Entrant GPA (z) & Share prioritario\\\\\n",
  "                   & (1) & (2)\\\\\n      \\midrule\n",
  row_tex("Treated", paste0(fmt(att), star(pri))),
  row_tex("", paste0("(", fmt(se), ")")),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(prem)),
  row_tex("Observations", format(nobs, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: Callaway--Sant'Anna ATT, not-yet-treated ",
  "controls. Region-clustered multiplier-bootstrap standard errors in parentheses (did ",
  "2.1.2 vintage). Both outcomes cover 2012--2023; each entrant's prioritario status is ",
  "that year's published designation. ", note_head, "}\n\\end{table}\n")
writeLines(tab, "tables/alloc_headline.tex")
cat(tab)
```

## Robustness

One specification per column, same outcome as column (1) of the headline; the
Baseline column repeats the headline fit so every check reads against it. The
column definitions follow the Piece 1 package: dropping Santiago entirely,
requiring at least five graded entrants per cell-year, excluding cells whose
pre-SAE fill test flags slack, weighting cells by their average entrant count, and
moving the base period one year before the door.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(seed_batt)
specs <- list(
  list(lab = "Baseline",         D = CP,                       w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "Drop Santiago",    D = CP[region != 13L],        w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "$n_{gpa}\\geq 5$", D = CP[n_gpa >= 5L],          w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "Slack excl.",      D = CP[slack_fail == FALSE],  w = NULL,    a = 0L, y = "gpa_z"),
  list(lab = "Entrant-wtd.",     D = CP,                       w = "w_ent", a = 0L, y = "gpa_z"),
  list(lab = "Anticip.\\ = 1",   D = CP,                       w = NULL,    a = 1L, y = "gpa_z"))
fitb <- lapply(specs, function(s) fit_cs(s$D, s$y, "region", anticipation = s$a,
                                         weightsname = s$w, bit = 999))
fitb[[1]] <- g_reg   # Baseline repeats the headline spec: reuse the headline fit
b_att <- vapply(fitb, `[[`, 0, "att"); b_se <- vapply(fitb, `[[`, 0, "se")
b_nobs <- vapply(fitb, `[[`, 0L, "n_cellyears")
b_pre  <- vapply(specs, function(s) pre_mean(s$D, s$y), 0)
b_p <- numeric(6L); b_p[1] <- p_g
for (i in 2:6) {
  s <- specs[[i]]
  z <- score_stream(prep_ri(s$D, s$y, wcol = s$w), wave_map(s$D),
                    seed_head, B_rob, antic = s$a)
  b_p[i] <- z$p
}
cat(sprintf("Randomization p by column: %s.\n\n", paste(sprintf("%.4f", b_p), collapse = ", ")))

labs <- vapply(specs, `[[`, "", "lab"); k <- length(labs)
tab_b <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Robustness: Effect on Entrant Prior-GPA}",
  "\\label{tab: alloc robustness}\n",
  "   \\centerline{\\scalebox{0.8}{\\begin{tabular}{l", strrep("c", k), "}\n      \\toprule\n",
  "                   & ", paste(labs, collapse = " & "), "\\\\\n",
  "                   & ", paste0("(", seq_len(k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
  row_tex("Treated", paste0(fmt(b_att), star(b_p))),
  row_tex("", paste0("(", fmt(b_se), ")")),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(b_pre)),
  row_tex("Observations", format(b_nobs, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: one specification change per column; Baseline ",
  "repeats the headline fit. Region-clustered multiplier-bootstrap standard errors in ",
  "parentheses. ", note_rob, "}\n\\end{table}\n")
writeLines(tab_b, "tables/alloc_robustness.tex")
cat(tab_b)
```

## Parallel trends: event studies, both outcomes

Dynamic ATT in the headline configuration, one figure per outcome, each cut at
event time $+1$ where several rollout waves identify every point; the full paths
print below. The exact randomization joint test of the leads (event times $-6$
through $-2$) is the inference of record for the pre-adoption years. The share
prioritario deep leads ($-8$, $-7$) are identified by the latest adopters; the
printed path shows them, and the joint test of leads $-6$ through $-2$ does not
reject.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
es_one <- function(D, yvar, ylab, fname, seed) {
  set.seed(seed)
  d <- D[!is.na(get(yvar)), .(cell_id, yr, G, region, y = get(yvar))]
  at <- att_gt(yname = "y", tname = "yr", idname = "cell_id", gname = "G", data = d,
               control_group = "notyettreated", panel = TRUE, allow_unbalanced_panel = TRUE,
               base_period = "universal", anticipation = 0L, bstrap = TRUE, biters = biters,
               clustervars = "region", print_details = FALSE)
  ed <- aggte(at, type = "dynamic", na.rm = TRUE)
  es <- data.table(e = ed$egt, att = ed$att.egt, se = ed$se.egt)
  Wstat <- as.numeric(at$W); Wp <- as.numeric(at$Wpval)
  print(es[is.finite(att), .(e, att = round(att, 4), se = round(se, 4))])
  leads_did <- es[e %in% -6:-2, setNames(att, as.character(e))]
  es <- es[e >= -6 & e <= 1 & is.finite(att) & is.finite(se)]
  lo <- es$att - 1.96 * es$se; hi <- es$att + 1.96 * es$se
  pdf(fname, width = 7.2, height = 4.8, family = "serif")
  par(mar = c(4.6, 5.0, 1.2, 1.2), las = 1, mgp = c(3.1, 0.8, 0),
      cex.axis = 1.05, cex.lab = 1.2)
  plot(es$e, es$att, type = "n",
       ylim = range(pretty(c(lo, hi))), xlim = c(min(es$e) - 0.4, max(es$e) + 0.4),
       xlab = "Event time (years relative to SAE adoption)", ylab = ylab, xaxt = "n")
  axis(1, at = min(es$e):max(es$e))
  abline(h = 0, col = "black", lwd = 1)
  abline(v = -0.5, col = "black", lty = 2)
  arrows(es$e, lo, es$e, hi, length = 0.045, angle = 90, code = 3, col = "black", lwd = 1.3)
  points(es$e, es$att, pch = 21, bg = "white", col = "black", cex = 1.25, lwd = 1.4)
  points(-1, 0, pch = 21, bg = "white", col = "black", cex = 1.25, lwd = 1.4)
  invisible(dev.off())
  list(W = Wstat, p = Wp, leads = leads_did)
}
j_g <- es_one(CP, "gpa_z", "Effect on entrant prior-GPA (z)",
              "figures/alloc_event_study.pdf", seed_head)
rp_g <- ri_pretrend_y(CP, "gpa_z", seed_head, check = j_g$leads)
cat(sprintf(paste0("`alloc_event_study.pdf`: exact RI joint lead test $p = %.2f$; ",
                   "Wald pre-test $W = %.1f$, $p = %.2f$.\n\n"), rp_g, j_g$W, j_g$p))
j_p <- es_one(CP, "sh_prio", "Effect on share prioritario",
              "figures/alloc_event_study_prio.pdf", seed_es)
rp_p <- ri_pretrend_y(CP, "sh_prio", seed_es, check = j_p$leads)
cat(sprintf(paste0("`alloc_event_study_prio.pdf`: exact RI joint lead test $p = %.2f$; ",
                   "Wald pre-test $W = %.1f$, $p = %.2f$.\n\n"), rp_p, j_p$W, j_p$p))
stopifnot("every gpa lead must stay under 0.03 SD" = max(abs(j_g$leads), na.rm = TRUE) < 0.03)
```

## Interpretation guards

Two checks that the headline is a change in who enters, with the measurement
itself standing still: the number of entrants per cell (seats are set by the
school and were full before and after SAE) and the share of a cell's entrants who
carry a prior-year GPA, computed for grades 2--12 where a prior GPA can exist.
Same estimator, standard errors, and randomization test as the headline; these are
the design's negative controls and are meant to stay unstarred, so the table
reports the randomization p directly.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
CP[, sh_gpa := fifelse(n_ent > 0L, n_gpa / n_ent, NA_real_)]
CPg <- CP[cod_nivel >= 2L]
set.seed(seed_head)
g_ent <- fit_cs(CP,  "n_ent",  "region")
g_cov <- fit_cs(CPg, "sh_gpa", "region")
z_ent <- score_stream(prep_ri(CP,  "n_ent"),  wave_map(CP),  seed_head, B_rob)
z_cov <- score_stream(prep_ri(CPg, "sh_gpa"), wave_map(CPg), seed_head, B_rob)

att_g  <- c(g_ent$att, g_cov$att)
se_g   <- c(g_ent$se,  g_cov$se)
pri_g  <- c(z_ent$p, z_cov$p)
prem_g <- c(pre_mean(CP, "n_ent"), pre_mean(CPg, "sh_gpa"))
nobs_g <- c(g_ent$n_cellyears, g_cov$n_cellyears)
tab_g <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Interpretation Guards: Entrant Counts and GPA Coverage}",
  "\\label{tab: alloc guards}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lcc}\n      \\toprule\n",
  "                   & Entrants per cell & Share with prior GPA\\\\\n",
  "                   & (1) & (2)\\\\\n      \\midrule\n",
  row_tex("Treated", fmt(att_g)),
  row_tex("", paste0("(", fmt(se_g), ")")),
  row_tex("RI two-sided p", sprintf("%.2f", pri_g)),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(prem_g)),
  row_tex("Observations", format(nobs_g, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: Callaway--Sant'Anna ATT, not-yet-treated ",
  "controls, as in the headline. Column (2) covers grades 2--12, where a prior-year GPA ",
  "can exist. Region-clustered multiplier-bootstrap standard errors in parentheses. ",
  note_gua, "}\n\\end{table}\n")
writeLines(tab_g, "tables/alloc_guards.tex")
cat(tab_g)
stopCluster(cl)
```
