---
title: "Piece 4 -- Where the Seats Are Lost: Decomposing the Assignment Effect"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r machinery, echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())
library(data.table)
library(fixest)
library(tools)

# Specs: DECOMP-GATES-2026-08-26, DECOMP-ROUTES-2026-08-26,
#        DECOMP-COUNT-2026-08-26, DECOMP-PACE-2026-08-26.
# RNG contract: every results chunk re-seeds 20260919 at its own top. Chunks
# share frame objects built upstream; run the document top to bottom.
set.seed(20260919)
B_RI <- 9999L

home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
ROOT <- file.path(home, "Dropbox", "Effect of Centralization")
# package-local data/ wins; certified panel from the Piece 3 package; scoring
# world from the mechanism exploration record (MD5s printed to the run log)
data_here <- function(f) {
  for (p in c(file.path("data", f),
              file.path(ROOT, "Updated Paper Spine", "Piece 3", f),
              file.path(ROOT, "Updated Paper Spine", "Piece 2", f),
              file.path(ROOT, "explorations", "2026-08-17_piece3_mechanism", "output", f),
              file.path(home, "Dropbox", "Effect SAE on SUA", "data", f)))
    if (file.exists(p)) return(p)
  stop(sprintf("input not found: %s", f))
}
dir.create("tables", showWarnings = FALSE)
say <- function(...) { cat(sprintf(...), "\n"); flush.console() }
fmt <- function(x) sprintf("%.4f", x)
star  <- function(p) ifelse(p < .01, "$^{***}$", ifelse(p < .05, "$^{**}$", ifelse(p < .1, "$^{*}$", "")))
starp <- function(p) ifelse(p < .01, "***", ifelse(p < .05, "**", ifelse(p < .1, "*", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# the certified frame machinery, verbatim from 1_effect_lottery.Rmd ----------
make_frame <- function(P, W, pur, cap, fc) {
  M <- copy(P); if (fc) M <- M[PREF == 1L]
  CU <- M[won == 1L, .(cutoff = max(loteria), n_win = .N), by = key]; M <- merge(M, CU, by = "key")[n_win >= W]
  M[, viol := won == 0L & loteria < cutoff]
  PU <- M[won == 0L, .(n_los = .N, n_viol = sum(viol)), by = key]
  M <- merge(M, PU, by = "key")[n_los >= 1L][n_viol / n_los <= pur]
  L <- M[won == 0L & loteria >= cutoff]
  if (cap) { setorder(L, key, loteria); L[, rk := seq_len(.N), by = key]; L <- L[rk <= n_win] }
  rbind(M[won == 1L], L, fill = TRUE)
}
core <- function(DT, y) {
  DT <- DT[!is.na(get(y))]; setorder(DT, key)
  KX <- DT[, .(i0 = .I[1], n = .N, m = sum(won)), by = key][m > 0L & m < n]
  DT <- DT[key %in% KX$key]; setorder(DT, key)
  KX <- DT[, .(i0 = .I[1], n = .N, m = sum(won)), by = key]
  denom <- KX[, sum(m * (n - m) / n)]; yv <- as.numeric(DT[[y]])
  base <- DT[, .(b = mean(get(y)) * sum(won)), by = key]$b
  obs <- (sum(yv * DT$won) - sum(base)) / denom
  pn <- numeric(B_RI)
  for (r in seq_len(nrow(KX))) { i0 <- KX$i0[r]; n <- KX$n[r]; m <- KX$m[r]
    yk <- yv[i0:(i0 + n - 1L)]; pn <- pn + .colSums(yk[replicate(B_RI, sample.int(n, m))], m, B_RI) }
  perm <- (pn - sum(base)) / denom
  fit <- feols(as.formula(paste0(y, " ~ won | key")), data = DT, cluster = ~sch_year, fixef.rm = "none")
  ct <- summary(fit)$coeftable
  stopifnot("closed form must equal the regression point" = abs(unname(ct["won", 1]) - obs) < 1e-8)
  data.table(itt = obs, se = ct["won", 2],
             ri_p = (1 + sum(abs(perm) >= abs(obs) - 1e-12)) / (B_RI + 1),
             n = nrow(DT), k = uniqueN(DT$key),
             lm = DT[won == 0L, mean(get(y))], win_mean = DT[won == 1L, mean(get(y))])
}
# nested-gate substitution split with RI; G = the ordered gate columns
chain_run <- function(A, G) {
  A <- copy(A); setorder(A, key)
  KX <- A[, .(i0 = .I[1], n = .N, m = sum(won)), by = key]
  stopifnot("all strata mixed" = KX[, all(m > 0L & m < n)])
  om <- KX[, m * (n - m) / n]; Som <- sum(om)
  YM <- as.matrix(A[, ..G]) * 1.0
  grp <- rep.int(seq_len(nrow(KX)), KX$n)
  TOT <- rowsum(YM, group = grp, reorder = FALSE)
  Sob <- rowsum(YM * A$won, group = grp, reorder = FALSE)
  cw <- om / KX$m; cl <- om / (KX$n - KX$m)
  lev <- function(S) list(w = colSums(cw * S) / Som, l = colSums(cl * (TOT - S)) / Som)
  splitk <- function(w, l) {
    rw <- w / c(1, head(w, -1)); rl <- l / c(1, head(l, -1))
    list(rw = rw, rl = rl, d = vapply(seq_along(rw), function(k)
      prod(rw[seq_len(k - 1L)]) * (rw[k] - rl[k]) * prod(rl[-seq_len(k)]), numeric(1)))
  }
  LO <- lev(Sob); DO <- splitk(LO$w, LO$l); d_obs <- DO$d
  set.seed(20260919)
  WA <- matrix(0, length(G), B_RI); LA <- matrix(0, length(G), B_RI)
  for (r in seq_len(nrow(KX))) {
    i0 <- KX$i0[r]; n <- KX$n[r]; m <- KX$m[r]
    idx <- replicate(B_RI, sample.int(n, m))
    blk <- YM[i0:(i0 + n - 1L), , drop = FALSE]
    for (g in seq_along(G)) {
      s <- .colSums(blk[idx, g], m, B_RI)
      WA[g, ] <- WA[g, ] + cw[r] * s
      LA[g, ] <- LA[g, ] + cl[r] * (TOT[r, g] - s)
    }
  }
  WA <- WA / Som; LA <- LA / Som
  DPERM <- vapply(seq_len(B_RI), function(b) splitk(WA[, b], LA[, b])$d, numeric(length(G)))
  ri_p <- vapply(seq_along(G), function(g)
    (1 + sum(abs(DPERM[g, ]) >= abs(d_obs[g]) - 1e-12)) / (B_RI + 1), numeric(1))
  list(d = d_obs, p = ri_p, DO = DO, LO = LO, tot = sum(d_obs))
}

# input fingerprints for the run log -----------------------------------------
for (f in c("lottery_panel_mech.rds", "score_students.rds", "score_prefs.rds",
            "pref_scores_final.rds", "program_weights_final.rds",
            "panel_programs_2007_2026.rds", "f1_selection_check.rds",
            "panel_performance_2002_2023.rds", "rendimiento_grades_2024_2025.rds"))
  say("input %-36s MD5 %s", f, unname(md5sum(data_here(f))))
for (f in list.files("data", full.names = TRUE))
  say("input %-36s MD5 %s", basename(f), unname(md5sum(f)))
say("input %-36s MD5 %s", "lottery_headline_values.csv",
    unname(md5sum(file.path(ROOT, "Updated Paper Spine", "Piece 2", "tables",
                            "lottery_headline_values.csv"))))
# the registered panel pin (all four specs): halt on any drift
stopifnot("panel of record must be MD5 6296a6de" =
  unname(md5sum(data_here("lottery_panel_mech.rds"))) == "6296a6de9feb98a18b5fc850e7b3a275")
```

## The frame and the gates

The certified 9th-door headline frame: first-choice applications to contested
classrooms, at least five winners per draw, the 10% violation screen, nearest
losers capped, one classroom by year by income tier per stratum -- asserted
equal to 45,123 applications in 1,215 strata. A seat requires four steps in
order; step 2 uses the definition of record (user ruling 2026-08-26): a usable
exam score means BOTH compulsory tests are present, because only then can any
program score be computed.

```{r frame, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(20260919)
Pm <- as.data.table(readRDS(data_here("lottery_panel_mech.rds")))
A  <- make_frame(copy(Pm)[, key := paste(key, prioritario)], 5L, 0.10, TRUE, TRUE)
stopifnot("certified frame" = nrow(A) == 45123L && uniqueN(A$key) == 1215L)
A[, cyear := proc + 5L]
stopifnot("child-year unique in the frame" = !anyDuplicated(A, by = c("MRUN", "cyear")))
ST <- as.data.table(readRDS(data_here("score_students.rds")))
stopifnot(!anyDuplicated(ST, by = c("MRUN", "YEAR")))
qi <- A[, .(MRUN, YEAR = cyear)]
A[, `:=`(raw_grad = fifelse(is.na(grad_ontime), 0L, as.integer(grad_ontime)), raw_reg = 0L, raw_both = 0L)]
A[!is.na(ST[qi, on = .(MRUN, YEAR), which = TRUE]), raw_reg := 1L]
A[which(ST[qi, on = .(MRUN, YEAR), !is.na(x.clec) & !is.na(x.mate1)]), raw_both := 1L]
A[, `:=`(raw_app = as.integer(applied), raw_asg = as.integer(assigned))]
A[, g1 := raw_grad]
A[, g2 := as.integer(g1 == 1L & raw_reg == 1L & raw_both == 1L)]
A[, g3 := as.integer(g2 == 1L & raw_app == 1L)]
A[, g4 := as.integer(g3 == 1L & raw_asg == 1L)]
say("gate counts: %s", paste(format(A[, sapply(.SD, sum), .SDcols = paste0("g", 1:4)],
                                    big.mark = ","), collapse = " | "))
gba <- A[, c(list(children = .N), lapply(.SD, sum)), .SDcols = paste0("g", 1:4), by = won][order(-won)]
say("gate counts by arm (children in study | pass gates 1-4):")
for (r in 1:2) say("  %s: %s children | %s | %s | %s | %s",
    ifelse(gba$won[r] == 1L, "winners", "losers"), format(gba$children[r], big.mark = ","),
    format(gba$g1[r], big.mark = ","), format(gba$g2[r], big.mark = ","),
    format(gba$g3[r], big.mark = ","), format(gba$g4[r], big.mark = ","))
stopifnot("both-tests applicants are the 15,906" = sum(A$g3) == 15906L,
          "seat holders are the 11,925" = sum(A$g4) == 11925L)

# external anchor: the assignment ITT on this frame must equal the certified
# Piece 2 headline row to machine precision
set.seed(20260919)
anchor <- core(A, "raw_asg")
CERT <- fread(file.path(ROOT, "Updated Paper Spine", "Piece 2", "tables", "lottery_headline_values.csv"))
stopifnot("assignment ITT reproduces the certified Piece 2 row" =
            abs(anchor$itt - CERT[outcome == "assigned", itt]) < 1e-10 &&
            abs(anchor$se - CERT[outcome == "assigned", se]) < 1e-10)
say("anchor: assignment ITT %+.4f (SE %.4f) = certified Piece 2 row, to machine precision",
    anchor$itt, anchor$se)
```

## Table 1 -- the four-gate split (DECOMP-GATES-2026-08-26)

```{r gates, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
G4 <- paste0("g", 1:4)
GATE <- c("Finished school on time", "Has a usable exam score",
          "Submitted an application", "Selected where she listed")
Rg <- chain_run(A, G4)
set.seed(20260919); chk <- core(A, "g4")
stopifnot("split reproduces the folded ITT" = abs(Rg$tot - chk$itt) < 1e-9)
# the pre-committed 81-child strict-order reconciliation to the certified row
say("assigned in the certified panel %s | folded final gate %s | difference %d",
    format(sum(A$raw_asg), big.mark = ","), format(sum(A$g4), big.mark = ","),
    sum(A$raw_asg) - sum(A$g4))
say("folded ITT %+.4f (clustered SE %.4f, RI p %.4f) vs certified assignment ITT %+.4f",
    100 * chk$itt, 100 * chk$se, chk$ri_p, 100 * anchor$itt)
stopifnot("the reconciliation is the 81 children" = sum(A$raw_asg) - sum(A$g4) == 81L)
# continuity: identical to the exploratory record (g12 both-tests column)
G12 <- fread(file.path(ROOT, "explorations", "2026-08-25_sheet_clarifications",
                       "output", "gate2_both_tests_values.csv"))[def == "both-tests"]
stopifnot("the exploratory record has exactly the four gate rows" = nrow(G12) == 4L,
          "pieces equal the exploratory record" = all(abs(100 * Rg$d - G12$per100) < 1e-9),
          "RI p equal the exploratory record"   = all(abs(Rg$p - G12$ri_p) < 1e-12))

tab <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Where the Fall in On-Time Assignment Happens}",
  "\\label{tab: gate decomposition}\n",
  "   \\centerline{\\scalebox{0.9}{\\begin{tabular}{lccc}\n      \\toprule\n",
  "                   & \\multicolumn{2}{c}{Pass rate (per 100 reaching)} & Contribution\\\\\n",
  "                   & Winners & Losers & (seats per 100)\\\\\n      \\midrule\n",
  paste0(vapply(1:4, function(k) row_tex(GATE[k],
    c(sprintf("%.2f", 100 * Rg$DO$rw[k]), sprintf("%.2f", 100 * Rg$DO$rl[k]),
      paste0(sprintf("%+.4f", 100 * Rg$d[k]), star(Rg$p[k])))), character(1)), collapse = ""),
  "      \\midrule\n",
  row_tex("Gets a seat (all four)",
          c(sprintf("%.2f", 100 * Rg$LO$w[4]), sprintf("%.2f", 100 * Rg$LO$l[4]),
            paste0(sprintf("%+.4f", 100 * Rg$tot), star(chk$ri_p)))),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: the chance of a seat is the product of the four ",
  "pass rates; the winner--loser difference in the product is divided into one additive piece ",
  "per step by substituting one rate at a time in time order (Kitagawa 1955; Das Gupta 1993; ",
  "ordering per Kim and Strobino 1984; funnel-decomposition precedent Chetty, Deming and ",
  "Friedman 2026). A usable exam score requires both compulsory tests. Rates are measured ",
  "within each draw and averaged with precision weights; the pieces sum to the difference ",
  sprintf("exactly (clustered SE of the total %.4f). Counted strictly step by step, 11,925 of the ", 100 * chk$se),
  "certified panel's 12,006 assigned children clear all four steps; the 81 remaining are ",
  "recorded with a seat despite failing an earlier step on paper, which is the whole gap ",
  "between this table's total and the certified assignment row ($-$1.8356 seats per 100). Stars from ",
  "randomization inference (9,999 within-draw reshuffles of who wins, the whole ",
  "split recomputed on each reshuffle): $^{***}$p$<$0.01, $^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab, "tables/decomp_gates.tex")
cat(tab)
fwrite(rbind(data.table(gate = GATE, w_pass = Rg$DO$rw, l_pass = Rg$DO$rl,
                        contribution_per100 = 100 * Rg$d, ri_p = Rg$p, stars = starp(Rg$p),
                        se = NA_real_),
             data.table(gate = "TOTAL (gets a seat)", w_pass = Rg$LO$w[4], l_pass = Rg$LO$l[4],
                        contribution_per100 = 100 * Rg$tot, ri_p = chk$ri_p,
                        stars = starp(chk$ri_p), se = 100 * chk$se)),
       "tables/decomp_gates_values.csv")
```

## Table 2 -- the selection split among applicants (DECOMP-ROUTES-2026-08-26)

Before 2024 DEMRE published the PACE and BEA selections in separate files; the
regular file alone is route-blind for 2021--2023. Folding the six separate
files in labels every year's routes and reconciles the assignment file with the
published statuses child for child, with nobody dropped.

```{r routes, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
AP <- A[g3 == 1L]
KXa <- AP[, .(n = .N, m = sum(won)), by = key][m > 0L & m < n]
EF <- AP[key %in% KXa$key]
say("applicant estimation frame: %s children, %s draws",
    format(nrow(EF), big.mark = ","), format(uniqueN(EF$key), big.mark = ","))

PALL <- as.data.table(readRDS(data_here("score_prefs.rds")))
kap <- EF[, .(MRUN, YEAR = cyear)]
PQ <- merge(PALL, kap, by = c("MRUN", "YEAR"))
PQ2 <- PQ[, .(l_bea = as.integer(any(tipo == "BEA")), l_pace = as.integer(any(tipo == "PACE")),
              l_gen = as.integer(any(tipo == "GENERO")),
              s_bea = as.integer(any(tipo == "BEA" & estado == 24L)),
              s_pace = as.integer(any(tipo == "PACE" & estado == 24L)),
              s_gen = as.integer(any(tipo == "GENERO" & estado == 24L)),
              pub_reg = as.integer(any(tipo == "REGULAR" & estado == 24L))), by = .(MRUN, YEAR)]
EF <- merge(EF, PQ2, by.x = c("MRUN", "cyear"), by.y = c("MRUN", "YEAR"), all.x = TRUE)
for (v in c("l_bea", "l_pace", "l_gen", "s_bea", "s_pace", "s_gen", "pub_reg"))
  EF[is.na(get(v)), (v) := 0L]

read_route <- function(f) {
  X <- fread(data_here(f), sep = ";", encoding = "UTF-8")
  ec <- grep("^ESTADO_PREF", names(X), value = TRUE)
  X[, .(MRUN = as.integer(MRUN), YEAR = as.integer(ANYO_PROCESO),
        sel = as.integer(rowSums(as.matrix(.SD) == 24L, na.rm = TRUE) > 0)), .SDcols = ec][
    , .(sel = max(sel), listed = 1L), by = .(MRUN, YEAR)]
}
PACEF <- rbind(read_route("C_POSTULANTES_SELECCION_PDT_CUPOS_PACE_2021_PUB_MRUN.csv"),
               read_route("C_POSTULANTES_SELECCION_PDT_CUPOS_PACE_2022_PUB_MRUN.csv"),
               read_route("C_POSTULACIONES_SELECCION_CUPOS_PACE_2023_PAES_MRUN.csv"))
BEAF  <- rbind(read_route("C_POSTULANTES_SELECCION_PDT_CUPOS_SUPERNUMERADOS_2021_PUB_MRUN.csv"),
               read_route("C_POSTULANTES_SELECCION_PDT_CUPOS_SUPERNUMERADOS_2022_PUB_MRUN.csv"),
               read_route("C_POSTULACIONES_SELECCION_CUPOS_SUPERNUMERARIOS_2023_PAES_PUB_MRUN.csv"))
stopifnot(!anyDuplicated(PACEF, by = c("MRUN", "YEAR")), !anyDuplicated(BEAF, by = c("MRUN", "YEAR")))
EF[, `:=`(fl_pace = 0L, fs_pace = 0L, fl_bea = 0L, fs_bea = 0L)]
EF[!is.na(PACEF[EF, on = .(MRUN, YEAR = cyear), which = TRUE]), fl_pace := 1L]
EF[PACEF[EF, on = .(MRUN, YEAR = cyear), x.sel] %in% 1L, fs_pace := 1L]
EF[!is.na(BEAF[EF, on = .(MRUN, YEAR = cyear), which = TRUE]), fl_bea := 1L]
EF[BEAF[EF, on = .(MRUN, YEAR = cyear), x.sel] %in% 1L, fs_bea := 1L]
stopifnot("separate files only carry 2021-2023" =
            EF[cyear >= 2024L, all(fl_pace + fs_pace + fl_bea + fs_bea == 0L)],
          "nobody selected in a special file is unplaced" =
            EF[(fs_pace == 1L | fs_bea == 1L) & raw_asg == 0L, .N] == 0L,
          "no child in two piles" =
            EF[fs_pace == 1L & fs_bea == 1L, .N] == 0L &&
            EF[(fs_pace == 1L | fs_bea == 1L) & pub_reg == 1L, .N] == 0L &&
            EF[(fs_pace == 1L | fs_bea == 1L) & (s_pace + s_bea + s_gen > 0L), .N] == 0L)
EF[, `:=`(S_PACE = as.integer(s_pace == 1L | fs_pace == 1L),
          S_BEA  = as.integer(s_bea  == 1L | fs_bea  == 1L),
          S_GEN  = s_gen,
          L_PACE = as.integer(l_pace == 1L | fl_pace == 1L),
          L_BEA  = as.integer(l_bea  == 1L | fl_bea  == 1L),
          L_GEN  = l_gen)]
EF[, NONREG := as.integer(S_PACE + S_BEA + S_GEN > 0L)]
stopifnot("routes exclusive" = EF[, all(S_PACE + S_BEA + S_GEN == NONREG)],
          "piles exclusive"  = EF[, all(pub_reg + NONREG <= 1L)])
EF[, ALLSEL := as.integer(pub_reg == 1L | NONREG == 1L)]
xt <- EF[, .N, keyby = .(raw_asg, ALLSEL)]
for (i in seq_len(nrow(xt)))
  say("assignment file %d, status codes %d: %s children", xt$raw_asg[i], xt$ALLSEL[i],
      format(xt$N[i], big.mark = ","))
stopifnot("both sources agree for every applicant, nobody dropped" = EF[, all(raw_asg == ALLSEL)])

LBL <- c(raw_asg = "Got a seat",
         pub_reg = "\\quad her score beat a program cutoff",
         NONREG  = "\\quad selected through a special route",
         S_PACE  = "\\qquad through PACE",
         S_BEA   = "\\qquad through BEA",
         S_GEN   = "\\qquad through the gender quota",
         L_PACE  = "Listed or held the PACE route",
         L_BEA   = "Listed or held the BEA route",
         L_GEN   = "Listed a gender-quota preference")
res <- list()
for (v in names(LBL)) { set.seed(20260919); res[[v]] <- core(EF, v) }
stopifnot("piles add"  = abs(res$pub_reg$itt + res$NONREG$itt - res$raw_asg$itt) < 1e-10,
          "routes add" = abs(res$S_PACE$itt + res$S_BEA$itt + res$S_GEN$itt - res$NONREG$itt) < 1e-10)
tab <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Why Applicants Are Selected Less: the Score Contest and the Special Routes}",
  "\\label{tab: route decomposition}\n",
  "   \\centerline{\\scalebox{0.9}{\\begin{tabular}{lccc}\n      \\toprule\n",
  "                   & Winner--loser diff. & (SE) & Loser mean\\\\\n      \\midrule\n",
  paste0(vapply(names(LBL), function(v) row_tex(LBL[[v]],
    c(paste0(sprintf("%+.4f", 100 * res[[v]]$itt), star(res[[v]]$ri_p)),
      paste0("(", sprintf("%.4f", 100 * res[[v]]$se), ")"),
      sprintf("%.2f", 100 * res[[v]]$lm))), character(1)), collapse = ""),
  "      \\midrule\n",
  row_tex("Observations / draws", c(format(res$raw_asg$n, big.mark = ","),
                                    format(res$raw_asg$k, big.mark = ","), "")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: per 100 applicants -- children who finished ",
  "school on time, hold both compulsory test scores, and submitted an application (15,906 of ",
  "the 45,123 frame children; the estimation frame keeps their two-sided draws), within-draw ",
  "comparisons. Special-route selections of 2021--2023 come from ",
  "DEMRE's separate PACE and BEA files, folded in so the published statuses agree with the ",
  "assignment file for every child; the two doors and the three routes are exact partitions, ",
  "so the sub-lines add to their parents to machine precision. School-by-year clustered SEs; ",
  "stars from randomization inference (9,999 within-draw reshuffles): $^{***}$p$<$0.01, ",
  "$^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab, "tables/decomp_routes.tex")
cat(tab)
fwrite(rbindlist(lapply(names(LBL), function(v) copy(res[[v]])[, outcome := v])),
       "tables/decomp_routes_values.csv")
```

## Table 3 -- the transcript swap (DECOMP-COUNT-2026-08-26)

```{r count, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
SC  <- as.data.table(readRDS(data_here("pref_scores_final.rds")))
CUT <- as.data.table(readRDS(data_here("panel_programs_2007_2026.rds")))[
  , .(carrera = as.integer(codigo_demre), YEAR, cutoff_reg)]
W   <- as.data.table(readRDS(data_here("program_weights_final.rds")))[
  , .(carrera, YEAR, w_nem, w_rank)]
NR  <- ST[, .(MRUN, YEAR, nem, rank)]
PP <- merge(merge(merge(SC, EF[, .(MRUN, YEAR = cyear)], by = c("MRUN", "YEAR")),
                  CUT, by = c("carrera", "YEAR"), all.x = TRUE),
            W, by = c("carrera", "YEAR"), all.x = TRUE)
PP[, usable := !is.na(score) & !is.na(cutoff_reg)]
UP <- PP[usable == TRUE]
stopifnot(UP[, !anyNA(w_nem) && !anyNA(w_rank)])
own <- UP[, .(own_clear = as.integer(any(score >= cutoff_reg)), n_usable = .N), by = .(MRUN, YEAR)]
EF <- merge(EF, own, by.x = c("MRUN", "cyear"), by.y = c("MRUN", "YEAR"), all.x = TRUE)
EF[is.na(own_clear), own_clear := 0L][is.na(n_usable), n_usable := 0L]
EF[, `:=`(nem_o = NR[EF[, .(MRUN, YEAR = cyear)], on = .(MRUN, YEAR), x.nem],
          rank_o = NR[EF[, .(MRUN, YEAR = cyear)], on = .(MRUN, YEAR), x.rank])]
F1 <- as.data.table(readRDS(data_here("f1_selection_check.rds")))
chk1 <- merge(EF[, .(MRUN, cyear, own_clear)], F1[, .(MRUN, cyear, sel_hat)], by = c("MRUN", "cyear"))
stopifnot("own-transcript clearing reproduces f1's sel_hat" =
            nrow(chk1) == nrow(EF) && chk1[, all(own_clear == sel_hat)])
EF[, has_nr := !is.na(nem_o) & !is.na(rank_o)]
EL <- EF[won == 0L & has_nr == TRUE]
pool <- EL[, .N, by = key]
EF[, pool_n := pool[EF, on = "key", x.N]][is.na(pool_n), pool_n := 0L]
CW <- EF[won == 1L & has_nr == TRUE & n_usable >= 1L & pool_n >= 1L]
say("funnel (by arm):                              winners   losers")
say("  estimation frame                            %7s  %7s",
    format(EF[won == 1L, .N], big.mark = ","), format(EF[won == 0L, .N], big.mark = ","))
say("  with official NEM and rank                  %7s  %7s",
    format(EF[won == 1L & has_nr, .N], big.mark = ","), format(EF[won == 0L & has_nr, .N], big.mark = ","))
say("  + at least one scoreable listed program     %7s  %7s",
    format(EF[won == 1L & has_nr & n_usable >= 1L, .N], big.mark = ","),
    format(EF[won == 0L & has_nr & n_usable >= 1L, .N], big.mark = ","))
say("  + eligible loser in the draw (COUNTED)      %7s", format(nrow(CW), big.mark = ","))
say("  excluded winners: no NEM/rank %d | no usable program %d | empty donor pool %d | missing gpa8 (companion only) %d",
    EF[won == 1L & has_nr == FALSE, .N], EF[won == 1L & has_nr & n_usable == 0L, .N],
    EF[won == 1L & has_nr & n_usable >= 1L & pool_n == 0L, .N], CW[is.na(gpa8_z), .N])
WPREF <- UP[CW[, .(MRUN, YEAR = cyear, key)], on = .(MRUN, YEAR), nomatch = NULL]
DON <- EL[, .(key, dMRUN = MRUN, dnem = nem_o, drank = rank_o)]
PAIRP <- merge(WPREF, DON, by = "key", allow.cartesian = TRUE)
stopifnot("no rescale branch" = PAIRP[, !anyNA(dnem) && !anyNA(drank)])
wnr <- CW[, .(MRUN, YEAR = cyear, wnem_o = nem_o, wrank_o = rank_o)]
PAIRP <- merge(PAIRP, wnr, by = c("MRUN", "YEAR"))
PAIRP[, cf_score := score + w_nem * (dnem - wnem_o) + w_rank * (drank - wrank_o)]
PW <- PAIRP[, .(cf_clear = as.integer(any(cf_score >= cutoff_reg))), by = .(MRUN, YEAR, key, dMRUN)]
CFB <- PW[, .(cfbar = mean(cf_clear), n_donors = .N), by = .(MRUN, YEAR, key)]
n0 <- nrow(CW)
CW <- merge(CW, CFB, by.x = c("MRUN", "cyear", "key"), by.y = c("MRUN", "YEAR", "key"))
stopifnot("every counted winner has a counterfactual" = nrow(CW) == n0 && !anyNA(CW$cfbar))
CS <- rbind(CW[, .(MRUN, cyear, key, sch_year, won, own_clear, cfbar)],
            EL[key %in% CW$key, .(MRUN, cyear, key, sch_year, won, own_clear, cfbar = NA_real_)])
KX2 <- CS[, .(n = .N, m = sum(won)), by = key]
stopifnot(KX2[, all(m > 0L & m < n)])
KX2[, om := m * (n - m) / n]; Som <- KX2[, sum(om)]
CS <- merge(CS, KX2[, .(key, n, m, om)], by = "key")
CS[, wgt := fifelse(won == 1L, om / m, om / (n - m))]
W_own <- CS[won == 1L, sum(wgt * own_clear)] / Som
W_cf  <- CS[won == 1L, sum(wgt * cfbar)] / Som
L_own <- CS[won == 0L, sum(wgt * own_clear)] / Som
set.seed(20260919); anch <- core(CS, "own_clear")
stopifnot("bridge equals the estimator" = abs((W_own - L_own) - anch$itt) < 1e-10)
# placebo and symmetric and companion
DONW <- CW[, .(key, dMRUN = MRUN, dnem = nem_o, drank = rank_o)]
PLP <- merge(WPREF, DONW, by = "key", allow.cartesian = TRUE)[MRUN != dMRUN]
PLP <- merge(PLP, wnr, by = c("MRUN", "YEAR"))
PLP[, cf_score := score + w_nem * (dnem - wnem_o) + w_rank * (drank - wrank_o)]
PLB <- PLP[, .(cf_clear = as.integer(any(cf_score >= cutoff_reg))), by = .(MRUN, YEAR, key, dMRUN)][
  , .(plbar = mean(cf_clear)), by = .(MRUN, YEAR)]
CWpl <- merge(CW, PLB, by.x = c("MRUN", "cyear"), by.y = c("MRUN", "YEAR"))
CL <- EF[won == 0L & has_nr == TRUE & n_usable >= 1L]
DON2 <- EF[won == 1L & has_nr == TRUE, .(key, dMRUN = MRUN, dnem = nem_o, drank = rank_o)]
pool2 <- DON2[, .N, by = key]
CL[, pool_n2 := pool2[CL, on = "key", x.N]][is.na(pool_n2), pool_n2 := 0L]
CL <- CL[pool_n2 >= 1L]
LPREF <- UP[CL[, .(MRUN, YEAR = cyear, key)], on = .(MRUN, YEAR), nomatch = NULL]
lnr <- CL[, .(MRUN, YEAR = cyear, wnem_o = nem_o, wrank_o = rank_o)]
SYP <- merge(merge(LPREF, DON2, by = "key", allow.cartesian = TRUE), lnr, by = c("MRUN", "YEAR"))
SYP[, cf_score := score + w_nem * (dnem - wnem_o) + w_rank * (drank - wrank_o)]
SYB <- SYP[, .(cf_clear = as.integer(any(cf_score >= cutoff_reg))), by = .(MRUN, YEAR, key, dMRUN)][
  , .(cfbar = mean(cf_clear)), by = .(MRUN, YEAR)]
CL <- merge(CL, SYB, by.x = c("MRUN", "cyear"), by.y = c("MRUN", "YEAR"))
# pre-committed companion: single nearest-gpa8 same-draw loser ---------------
CWnn <- CW[!is.na(gpa8_z)]
DNN <- EL[!is.na(gpa8_z), .(key, dMRUN = MRUN, dgpa8 = gpa8_z, dlot = loteria)]
MM <- merge(CWnn[, .(MRUN, cyear, key, gpa8_z)], DNN, by = "key", allow.cartesian = TRUE)
MM[, gap := abs(gpa8_z - dgpa8)]
setorder(MM, MRUN, cyear, gap, dlot)
NN <- MM[, .SD[1L], by = .(MRUN, cyear)]
NNC <- merge(NN[, .(MRUN, YEAR = cyear, key, dMRUN, gap)], PW,
             by = c("MRUN", "YEAR", "key", "dMRUN"))
CWnn <- merge(CWnn, NNC[, .(MRUN, cyear = YEAR, nn_clear = cf_clear, gap)], by = c("MRUN", "cyear"))

say("counted winners %s | own clears %.2f | with loser transcript %.2f | net %+.4f plain, %+.4f weighted",
    format(nrow(CW), big.mark = ","), 100 * CW[, mean(own_clear)], 100 * CW[, mean(cfbar)],
    100 * CW[, mean(own_clear - cfbar)], 100 * (W_own - W_cf))
say("anchor (ITT on own-transcript clearing, counted subsample): %+.4f (clustered SE %.4f, RI p %.4f) n %s k %s",
    100 * anch$itt, 100 * anch$se, anch$ri_p,
    format(anch$n, big.mark = ","), format(anch$k, big.mark = ","))
say("bridge: transcript piece %+.4f + exam and own-list remainder %+.4f = %+.4f (asserted)",
    100 * (W_own - W_cf), 100 * (W_cf - L_own), 100 * (W_own - L_own))
say("placebo (winners with a same-arm donor, %s of the counted): net %+.4f",
    format(nrow(CWpl), big.mark = ","), 100 * CWpl[, mean(own_clear - plbar)])
say("mirror image, per 100 counted LOSERS (%s): own %.2f | with winner transcript %.2f | net %+.4f",
    format(nrow(CL), big.mark = ","), 100 * CL[, mean(own_clear)], 100 * CL[, mean(cfbar)],
    100 * CL[, mean(own_clear - cfbar)])
say("companion (nearest-gpa8 single match, %s matched winners): net %+.4f | |gpa8 z-gap| median %.3f p90 %.3f",
    format(nrow(CWnn), big.mark = ","), 100 * CWnn[, mean(own_clear - nn_clear)],
    CWnn[, median(gap)], CWnn[, quantile(gap, .9)])
say("DISCLOSURES (pre-committed): (1) descriptive counts, no stars; the placebo is the noise floor.")
say("(2) transcript-swapped-first path: exam scores and the program list stay at treated values;")
say("    a different swap order gives a different attribution; pieces not claimed to add across channels.")
say("(3) cutoffs are the realized clearing scores held fixed (partial equilibrium).")
say("(4) removed by user order 2026-08-26: the registered has_nem disclosure was a full-frame")
say("    number restating the graduation gate and served no purpose here.")
say("(5) the counting lives in the computed-rule world (96.6%% agreement with the assignment file")
say("    over all applicants); counted-subsample agreement %.4f.", CS[, mean(own_clear == EF[CS, on = .(MRUN, cyear), x.raw_asg])])

fwrite(data.table(
  line = c("winners counted", "own clears per 100", "with loser transcript per 100",
           "net plain per 100 counted winners", "net weighted per 100",
           "anchor ITT own_clear per 100", "anchor clustered SE per 100", "anchor RI p",
           "anchor n", "anchor k", "bridge transcript piece", "bridge remainder",
           "placebo winners with same-arm donor", "placebo net per 100",
           "losers counted (mirror)", "mirror net per 100 counted losers",
           "companion matched winners", "companion net per 100", "companion gpa8 gap median"),
  value = c(nrow(CW), 100 * CW[, mean(own_clear)], 100 * CW[, mean(cfbar)],
            100 * CW[, mean(own_clear - cfbar)], 100 * (W_own - W_cf),
            100 * anch$itt, 100 * anch$se, anch$ri_p, anch$n, anch$k,
            100 * (W_own - W_cf), 100 * (W_cf - L_own),
            nrow(CWpl), 100 * CWpl[, mean(own_clear - plbar)],
            nrow(CL), 100 * CL[, mean(own_clear - cfbar)],
            nrow(CWnn), 100 * CWnn[, mean(own_clear - nn_clear)], CWnn[, median(gap)])),
  "tables/decomp_count_values.csv")

tab3 <- sprintf(paste0("\\begin{table}[htbp]\n   \\caption{The Transcript Swap: Seats at Real Weights and Real Cutoffs}",
  "\\label{tab: transcript swap}\n   \\centerline{\\scalebox{0.9}{\\begin{tabular}{lc}\n      \\toprule\n",
  "                   & Per 100 counted\\\\\n      \\midrule\n",
  "      Clears a listed cutoff, own transcript (winners) & %.2f\\\\\n",
  "      Same girl, a draw-loser's GPA and class-rank points (winners) & %.2f\\\\\n      \\midrule\n",
  "      Net transcript piece, per 100 counted winners & %+.2f$^{***}$\\\\\n",
  "      Noise floor (same-arm placebo), per 100 counted winners & %+.2f\\\\\n",
  "      Mirror image (winners' transcripts), per 100 counted losers & %+.2f$^{***}$\\\\\n",
  "      Companion (closest 8th-grade-GPA loser), per 100 matched winners & %+.2f\\\\\n",
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: the two level rows are counts. The net ",
  "transcript piece and the mirror carry permutation p-values from a registered addendum ",
  "test (spec DECOMP-COUNT-RI-2026-08-26, run log in the package): reshuffle who wins ",
  "within each draw exactly as the lottery did, rebuild the entire count, 9,999 times; ",
  "the observed nets exceed all 9,999 reshuffles, p $=$ .0001 each, as does the ",
  "estimator-weighted net ($-$2.34, p $=$ .0001). The re-randomization template is Cullen, ",
  "Jacob and Levitt (2006); the counterfactual-recomputation practice follows Ellison and ",
  "Pathak (2021) and Dur, Kominers, Pathak and S\\\"onmez (2018). ",
  "The same-arm placebo is the noise floor and sits inside the reshuffled distribution. ",
  "A counted winner has her official GPA score ",
  "(called NEM in the admission system) and class-rank points on file, at least one ",
  "scoreable listed program, and at least one eligible loser in her draw (a loser with her ",
  "own GPA and rank points on file). Her GPA and rank points are replaced jointly by each ",
  "eligible same-draw loser's, her own exam scores and program list kept, clearing ",
  "re-evaluated at real weights and real cutoffs and averaged over the draw's losers; exact ",
  "by linearity of the score in the GPA and rank points. ",
  "This is the transcript-swapped-first path: exam scores and the program list stay at their ",
  "treated values, a different swap order gives a different attribution, and the pieces are ",
  "not claimed to add across channels. Cutoffs are the realized clearing scores held fixed. ",
  "Clearing is evaluated in the reconstructed-score world, which agrees with the assignment ",
  "file for 96.6\\%% of applicants.}\n\\end{table}\n"),
  100 * CW[, mean(own_clear)], 100 * CW[, mean(cfbar)], 100 * CW[, mean(own_clear - cfbar)],
  100 * CWpl[, mean(own_clear - plbar)], 100 * CL[, mean(own_clear - cfbar)],
  100 * CWnn[, mean(own_clear - nn_clear)])
cat(tab3)
writeLines(tab3, "tables/decomp_count.tex")
```

## Table 4 -- the PACE door split (DECOMP-PACE-2026-08-26)

```{r pace, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
YM <- c(310L, 410L, 510L, 610L, 710L, 810L, 910L)
PF <- as.data.table(readRDS(data_here("panel_performance_2002_2023.rds")))
R12 <- PF[COD_ENSE %in% YM & COD_GRADO == 4L & AGNO %in% 2020:2023 & !is.na(MRUN) &
            !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7,
          .(MRUN = as.integer(MRUN), AGNO = as.integer(AGNO), RBD = as.integer(RBD))]
RS <- as.data.table(readRDS(data_here("rendimiento_grades_2024_2025.rds")))
R12 <- unique(rbind(R12, RS[COD_ENSE %in% YM & COD_GRADO == 4L & !is.na(MRUN) &
                              !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7,
                            .(MRUN = as.integer(MRUN), AGNO = as.integer(AGNO), RBD = as.integer(RBD))]))
S12 <- R12[, .(nsch = uniqueN(RBD), RBD = RBD[1L]), by = .(MRUN, AGNO)]
S12s <- S12[nsch == 1L]
LIST_PACE <- unique(rbind(
  read_route("C_POSTULANTES_SELECCION_PDT_CUPOS_PACE_2021_PUB_MRUN.csv")[, .(MRUN, YEAR)],
  read_route("C_POSTULANTES_SELECCION_PDT_CUPOS_PACE_2022_PUB_MRUN.csv")[, .(MRUN, YEAR)],
  read_route("C_POSTULACIONES_SELECCION_CUPOS_PACE_2023_PAES_MRUN.csv")[, .(MRUN, YEAR)],
  unique(PALL[tipo == "PACE", .(MRUN, YEAR)])))
LIST_PACE[, AGNO := YEAR - 1L]
FLAG <- unique(merge(LIST_PACE, S12s, by = c("MRUN", "AGNO"))[, .(RBD, YEAR)])
EF[, AGNO12 := cyear - 1L]
EF[, `:=`(sch12 = S12s[EF, on = .(MRUN, AGNO = AGNO12), x.RBD],
          nsch12 = S12[EF, on = .(MRUN, AGNO = AGNO12), x.nsch])]
EF[, S_raw := as.integer(!is.na(sch12) &
                           !is.na(FLAG[EF, on = .(RBD = sch12, YEAR = cyear), which = TRUE]))]
EF[, S := as.integer(S_raw == 1L | L_PACE == 1L)]
orf <- EF[L_PACE == 1L & S_raw == 0L]
say("route-holders whose school the join could not locate (counted into the school gate only through their own route-holding): %d of %s",
    nrow(orf), format(nrow(EF), big.mark = ","))
say("non-door children with no located single 12th-grade school (fold to not-PACE): %d (%.4f; lineage diagnostic, disclosed not halting)",
    EF[L_PACE == 0L & is.na(sch12), .N], EF[L_PACE == 0L & is.na(sch12), .N] / nrow(EF))
stopifnot("join failures below 1% (registered halt)" = nrow(orf) / nrow(EF) < 0.01,
          "door nests in the school gate" = EF[, all(L_PACE <= S)])
set.seed(20260919); door <- core(EF, "L_PACE")
Rp <- chain_run(EF, c("S", "L_PACE"))
stopifnot("pieces add to the door line" = abs(Rp$tot - door$itt) < 1e-9)
# sensitivities: official PDF lists substituted; any-school rule
PDF <- fread(data_here("pace_pdf_rbds.csv"))
cov_yrs <- c(2022L, 2023L, 2025L)
FLAG2 <- rbind(FLAG[!YEAR %in% cov_yrs], PDF[, .(RBD = rbd, YEAR = year)])
EF[, S2 := as.integer((!is.na(sch12) &
      !is.na(FLAG2[EF, on = .(RBD = sch12, YEAR = cyear), which = TRUE])) | L_PACE == 1L)]
R2 <- chain_run(EF, c("S2", "L_PACE"))
stopifnot(abs(R2$tot - door$itt) < 1e-9)
LJa <- merge(LIST_PACE, R12, by = c("MRUN", "AGNO"), allow.cartesian = TRUE)
FLAGa <- unique(LJa[, .(RBD, YEAR)])
EFR <- merge(EF[, .(MRUN, cyear, AGNO = AGNO12)], R12, by = c("MRUN", "AGNO"), allow.cartesian = TRUE)
EFR[, f := !is.na(FLAGa[EFR, on = .(RBD, YEAR = cyear), which = TRUE])]
ANY <- EFR[, .(anyf = as.integer(any(f))), by = .(MRUN, cyear)]
EF[, S_any := ANY[EF, on = .(MRUN, cyear), x.anyf]][is.na(S_any), S_any := 0L]
EF[, S3 := as.integer(S_any == 1L | L_PACE == 1L)]
R3 <- chain_run(EF, c("S3", "L_PACE"))
stopifnot(abs(R3$tot - door$itt) < 1e-9)
tab4 <- paste0(
  "\\begin{table}[htbp]\n   \\caption{The PACE Route Loss: the School Piece and the Standing Piece}",
  "\\label{tab: pace split}\n",
  "   \\centerline{\\scalebox{0.9}{\\begin{tabular}{lccc}\n      \\toprule\n",
  "                   & Derived flag & Official lists & Any-school rule\\\\\n      \\midrule\n",
  row_tex("Her school is not in the PACE program",
          c(paste0(sprintf("%+.4f", 100 * Rp$d[1]), star(Rp$p[1])),
            paste0(sprintf("%+.4f", 100 * R2$d[1]), star(R2$p[1])),
            paste0(sprintf("%+.4f", 100 * R3$d[1]), star(R3$p[1])))),
  row_tex("Standing inside a PACE school",
          c(paste0(sprintf("%+.4f", 100 * Rp$d[2]), star(Rp$p[2])),
            paste0(sprintf("%+.4f", 100 * R2$d[2]), star(R2$p[2])),
            paste0(sprintf("%+.4f", 100 * R3$d[2]), star(R3$p[2])))),
  "      \\midrule\n",
  row_tex("Adds to the route line", rep(sprintf("%+.4f", 100 * door$itt), 3)),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: per 100 applicants. The route line is the ",
  "winner--loser difference in listing or holding the PACE route. A school counts as ",
  "PACE-enrolled for admission year X when any of its year X$-$1 12th-graders appears in the ",
  "year-X PACE application data (derived flag); the second column substitutes the official ",
  "school lists for the three admission years such lists exist (2022, 2023, 2025) -- no ",
  "official list exists for admissions 2021, 2024, or 2026, where the derived flag stands ",
  "alone in every column; the third column locates ",
  "children graded at more than one school by any of their schools. Standing absorbs program ",
  "completion, school tenure, and listing behavior; the official rank rule (top 25\\% of her ",
  "school's ranking score, or 830 points nationally; Subsecretar\\'ia de Educaci\\'on Superior, ",
  "REX 7.447/2025 as consolidated by REX 4042/2026) gates the guaranteed seat itself, ",
  "downstream of this line. Winner and loser rates: school in PACE ",
  sprintf("%.2f vs %.2f per 100; holds the route among children at PACE schools %.2f vs %.2f. ",
          100 * Rp$LO$w[1], 100 * Rp$LO$l[1], 100 * Rp$DO$rw[2], 100 * Rp$DO$rl[2]),
  "The school piece prices the school gap at the losers' route-holding rate, the standing ",
  "piece prices the standing gap at the winners' PACE-school exposure. The 12th-grade school ",
  "is itself part of what the draw changed, so this split sorts children by an outcome of ",
  "the draw; it is an accounting of the route loss. Reading either piece as a separate ",
  sprintf("causal effect would require more than this table claims. Route line: %+.4f (clustered SE %.4f, RI p %.4f). Stars from ",
          100 * door$itt, 100 * door$se, door$ri_p),
  "randomization inference (9,999 within-draw reshuffles): $^{***}$p$<$0.01, ",
  "$^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab4, "tables/decomp_pace.tex")
cat(tab4)
say("split conditions on the post-lottery 12th-grade school; in-frame under the decomposition ruling of 2026-08-18")
fwrite(rbind(data.table(flag = rep(c("derived", "official-lists", "any-school"), each = 2),
                        piece = rep(c("school not in PACE", "standing"), 3),
                        per100 = c(100 * Rp$d, 100 * R2$d, 100 * R3$d),
                        ri_p = c(Rp$p, R2$p, R3$p), se = NA_real_, n = NA_integer_, k = NA_integer_),
             data.table(flag = "door line", piece = "listed/held PACE",
                        per100 = 100 * door$itt, ri_p = door$ri_p, se = 100 * door$se,
                        n = door$n, k = door$k),
             data.table(flag = c("arm rates", "arm rates"),
                        piece = c("school in PACE, winners/losers per 100",
                                  "holds door at PACE school, winners/losers per 100"),
                        per100 = c(100 * Rp$LO$w[1], 100 * Rp$DO$rw[2]),
                        ri_p = NA_real_, se = NA_real_,
                        n = NA_integer_, k = NA_integer_)[
                          , loser_rate := c(100 * Rp$LO$l[1], 100 * Rp$DO$rl[2])][],
             fill = TRUE),
       file = "tables/decomp_pace_values.csv")
fwrite(FLAG, "tables/pace_school_flag_derived.csv")
say("PIECE 4 ANALYSIS COMPLETE")
```
