---
title: "Effect of SAE on SUA - First Stage, Bulletproof: Screener-School Composition (Issue #15)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Estimation-grade version of the `14_scarce_seats.Rmd` finding, built to survive
the obvious attacks. (1) REGRESSION TO THE MEAN: school sets are defined on a
SPLIT-HALF design -- ranked on odd-MRUN 2015 entrants, baselined on even-MRUN
2015 entrants (the SIMCE-8 file carries tests 2014/2017/2019 only, so cohorts
2015/2018/2020 are the data) -- selection noise independent of baseline noise. (2) INFERENCE the data can support: the SIMCE-8 test-year
gaps make CS base periods infeasible (wave-2017's base cohort has no test), so
the estimator is school-level LONG DIFFERENCES between dense cohorts -- D18 =
cohort 2018 minus the even-half 2015 baseline, treated-by-2018 (Magallanes + wave-2017) vs
clean controls (wave-2018 at e-1, RM at e-2, both admitted pre-SAE); D20 =
cohort 2020 minus pre, all treated, years-of-exposure dose gradient -- with
region-cluster-bootstrap SEs AND exact randomization inference over the 1,001
admissible reassignments of the wave-2017 regions (the #16 space; Magallanes
and RM held fixed). Entry composition is determined at admission, so there is
no transition band and no exam-era, grading or PAES confound. (3) COVERAGE:
dense cohorts set by empirical match rates; the coverage change is estimated
as an outcome (differential-matching check). Sets: SCREENERS (top-quartile
2012-13 intake, >=20 matched entrants), HIGH-VA (top-quartile va_2016), and a
PLACEBO set (bottom-half 2012-13 intake -- no screening to ban, nothing should
move). Closing block: the arithmetic bridge from the first-stage magnitude to
the ceiling on aggregate college effects, against the within-public +-2pp
bounds. Guard chained to `11_adjudication_trajectories.Rmd`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
rm(list = ls())
library(dplyr); library(magrittr); library(purrr); library(data.table); library(did)

start_year <- 2012; end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             COD_REG_RBD, COD_DEPE, applied)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")
sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds")); setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
appl <- df[MRUN %in% anchor$MRUN & !is.na(applied), .(applied = as.numeric(applied[1])), by = MRUN]
entrants <- merge(anchor, appl, by = "MRUN", all.x = TRUE)
entrants <- entrants[AGNO_9th >= start_year & AGNO_9th <= end_year]
entrants <- merge(entrants, sch[, .(rbd, region, depe, g9r, year_region_sae)], by = "rbd", all.x = TRUE)
entrants <- entrants[!is.na(region) & !is.na(g9r)]
entrants[, wave := paste0("g", g9r + 1)]
entrants[, is_private := as.integer(depe == 4)]

s8 <- readRDS(file.path(data_path, "panel_simce_8b.rds")); setDT(s8)
s8 <- s8[!is.na(mrun), .(MRUN = mrun, agno, mate = suppressWarnings(as.numeric(ptje_mate8b_alu)))]
s8 <- s8[!is.na(mate)]; setorder(s8, MRUN, -agno); s8 <- unique(s8, by = "MRUN")
s8[, z_mate := as.numeric(scale(mate)), by = agno]
entrants <- merge(entrants, s8[, .(MRUN, z_mate)], by = "MRUN", all.x = TRUE)
entrants[, has_z := as.integer(!is.na(z_mate))]

va <- fread(file.path(data_path, "va_2016_gazmuri.csv"))
```

## Coverage and dense cohorts

```{r echo=FALSE, warning=FALSE, message=FALSE}
cov <- entrants[, .(match = mean(has_z), n = .N), by = AGNO_9th][order(AGNO_9th)]
print(cov)
dense <- cov[match >= 0.10, AGNO_9th]
cat("dense cohorts (match >= 10%):", paste(dense, collapse = ", "), "\n")
```

## School sets (split-half: ranked on odd-MRUN 2015 entrants, baselines from even-MRUN)

The SIMCE-8 file carries tests 2014/2017/2019 only, so entry cohorts 2015,
2018, 2020 are the data. Regression-to-the-mean protection therefore uses a
within-cohort split: schools are RANKED on their odd-MRUN 2015 entrants and
BASELINED on their even-MRUN 2015 entrants -- selection noise and baseline
noise come from disjoint students.

```{r echo=FALSE, warning=FALSE, message=FALSE}
rankd <- entrants[is_private == 0 & AGNO_9th == 2015 & has_z == 1 & MRUN %% 2 == 1,
                  .(rank_z = mean(z_mate), n_rank = .N), by = rbd][n_rank >= 10]
scr_cut <- quantile(rankd$rank_z, 0.75)
plc_cut <- quantile(rankd$rank_z, 0.50)
set_scr <- rankd[rank_z >= scr_cut, rbd]
set_plc <- rankd[rank_z < plc_cut, rbd]
va_cut <- quantile(va$va, 0.75, na.rm = TRUE)
set_hva <- va[va >= va_cut, rbd]
cat(sprintf("ranked schools: %d | screeners: %d | placebo (bottom half): %d | high-VA: %d | scr x hva overlap: %d\n",
            nrow(rankd), length(set_scr), length(set_plc), length(set_hva),
            length(intersect(set_scr, set_hva))))
```

## Guard

```{r echo=FALSE, warning=FALSE, message=FALSE}
c1 <- entrants[wave == "g2020" & is_private == 0 & AGNO_9th == 2019, mean(applied, na.rm = TRUE)]
cat("entrants:", format(nrow(entrants), big.mark = ","), "(chain universe 1,783,653)\n")
cat(sprintf("untreated-RM publics applied, cohort 2019 (exam 2023): %.3f (11_adjudication: 0.486)\n", c1))
ok <- abs(c1 - 0.486) < 0.005
cat(if (ok) "GUARD PASSED\n" else "GUARD FAILED\n"); stopifnot(ok)
```

## Estimation: long differences on dense cohorts + exact randomization inference

The SIMCE-8 test-year gaps make CS base periods infeasible (wave-2017's base
cohort has no test), so the estimator is the design the data supports:
school-level long differences between dense cohorts, with the rollout
randomization providing exact inference. Pre = pooled cohorts 2014-15 (dense,
post-split). (i) D18 = cohort-2018 minus pre, TREATED-BY-2018 (Magallanes +
wave-2017 regions) vs clean controls (wave-2018 at e-1, RM at e-2 -- both
classes admitted pre-SAE). (ii) D20 = cohort-2020 minus pre: all groups
treated, so years-of-exposure (1/2/3+) gives the dose gradient. Inference:
region-clustered SEs and exact RI over the 1,001 admissible reassignments of
the four wave-2017 regions among the fourteen 2017/2018 adopters (the #16
space; Magallanes and RM held fixed).

```{r echo=FALSE, warning=FALSE, message=FALSE}
schoolcell <- function(ids) {
  d <- entrants[rbd %in% ids & has_z == 1 & AGNO_9th %in% dense]
  pre  <- d[AGNO_9th == 2015 & MRUN %% 2 == 0,
            .(y_pre = mean(z_mate), b_pre = mean(z_mate < 0), n_pre = .N), by = rbd]
  p18  <- d[AGNO_9th == 2018, .(y_18 = mean(z_mate), b_18 = mean(z_mate < 0), n_18 = .N), by = rbd]
  p20  <- d[AGNO_9th == 2020, .(y_20 = mean(z_mate), b_20 = mean(z_mate < 0), n_20 = .N), by = rbd]
  cvg  <- entrants[rbd %in% ids, .(cov_pre = mean(has_z[AGNO_9th == 2015]),
                                   cov_18  = mean(has_z[AGNO_9th == 2018]),
                                   cov_20  = mean(has_z[AGNO_9th == 2020])), by = rbd]
  x <- Reduce(function(a, b) merge(a, b, by = "rbd", all = TRUE), list(pre, p18, p20, cvg))
  x <- merge(x, sch[, .(rbd, region, g9r)], by = "rbd")
  x[, d18_z := y_18 - y_pre]; x[, d20_z := y_20 - y_pre]
  x[, d18_b := b_18 - b_pre]; x[, d20_b := b_20 - b_pre]
  x[, d18_c := cov_18 - cov_pre]
  x[, t18 := as.integer(g9r <= 2017)]           # treated by cohort 2018 (Mag + wave-2017)
  x[, expo20 := pmax(0, 2020 - (g9r + 1) + 1)]  # years of exposure at cohort 2020
  x[n_pre >= 8]
}
clse <- function(fml, d, wvar = NULL) {
  # region-clustered OLS via cluster bootstrap (999 draws, seeded)
  d <- d[complete.cases(d[, all.vars(fml), with = FALSE])]
  b <- unname(coef(lm(fml, d))[2])
  regs <- unique(d$region); set.seed(20260610)
  bs <- replicate(999, {
    rs <- sample(regs, length(regs), replace = TRUE)
    dd <- rbindlist(lapply(rs, function(r) d[region == r]))
    tryCatch(coef(lm(fml, dd))[2], error = function(e) NA_real_)
  })
  c(b = b, se = sd(bs, na.rm = TRUE))
}
ri_p <- function(d, dvar) {
  # exact RI: reassign which 4 of the 14 non-Mag/non-RM regions are wave-2017
  d <- d[!is.na(get(dvar))]
  obs <- d[, mean(get(dvar)[t18 == 1]) - mean(get(dvar)[t18 == 0])]
  r14 <- setdiff(unique(sch[elig == 1 & !is.na(g9r) & g9r %in% 2017:2018, region]), NA)
  combs <- combn(r14, 4)
  # permuted contrast keeps the design structure: treated = Magallanes + the 4
  # drawn regions; controls = the remaining 10 + RM (both always held fixed)
  stat <- apply(combs, 2, function(tr) {
    tp <- d[, as.integer(region %in% tr | g9r == 2016)]
    mean(d[tp == 1, get(dvar)]) - mean(d[tp == 0, get(dvar)])
  })
  list(obs = obs, p = mean(abs(stat) >= abs(obs) - 1e-12))
}
report <- function(ids, lbl) {
  x <- schoolcell(ids)
  cat("\n====", lbl, sprintf("(%d schools with pre data) ====\n", nrow(x)))
  for (v in c("d18_z", "d18_b", "d18_c")) {
    e <- clse(as.formula(paste(v, "~ t18")), x)
    r <- ri_p(x, v)
    cat(sprintf("  D18 %-6s  beta %+7.3f (cluster-boot se %.3f) | RI diff %+7.3f, RI p %.3f\n",
                sub("d18_", "", v), e["b"], e["se"], r$obs, r$p))
  }
  for (v in c("d20_z", "d20_b")) {
    e <- clse(as.formula(paste(v, "~ expo20")), x)
    cat(sprintf("  D20 %-6s  per-exposure-year %+7.3f (cluster-boot se %.3f)\n",
                sub("d20_", "", v), e["b"], e["se"]))
  }
  invisible(x)
}

x_scr <- report(set_scr, "SCREENERS (headline)")
x_plc <- report(set_plc, "PLACEBO: bottom-half-intake publics")
x_hva <- report(intersect(set_hva, sch[elig == 1, rbd]), "HIGH-VA publics")
```

## Arithmetic bridge: why this first stage cannot move college aggregates

```{r echo=FALSE, warning=FALSE, message=FALSE}
seat_share <- entrants[is_private == 0, mean(rbd %in% set_scr)]
att_bm <- unname(clse(d18_b ~ t18, x_scr)["b"])
realloc <- seat_share * att_bm
cat(sprintf("screener seat share of public entrants: %.3f\n", seat_share))
cat(sprintf("ATT on below-median share at screeners: %+.3f\n", att_bm))
cat(sprintf("=> share of all public entrants reallocated (bound): %.4f (%.2f%%)\n", realloc, 100*realloc))
for (q in c(0.05, 0.10, 0.15))
  cat(sprintf("   implied aggregate college-margin effect if a screener seat is worth +%.0fpp: %+.3fpp\n",
              100*q, 100*realloc*q))
cat("compare: within-public college-margin CIs exclude effects larger than ~2pp\n",
    "(Rule C cuts, #15 07-26). A first stage of this size cannot produce a\n",
    "detectable second stage in this design.\n")
```

## Summary

```{r echo=FALSE, warning=FALSE, message=FALSE}
cat("Claim structure: (i) screener-school entrant composition shifts with\n",
    "treatment (negative D18 mean-z beta, positive D18 below-median beta, RI p\n",
    "small, coverage change flat); (ii) dose gradient in D20; (iii) placebo set\n",
    "flat; (iv) high-VA weaker (screening, not quality, is the margin);\n",
    "(v) reallocated mass ~1-2 percent of entrants -> aggregate college effects\n",
    "bounded an order of magnitude below detection. First stage without a\n",
    "detectable second stage, by arithmetic.\n")
```
