---
title: "Effect of SAE on SUA - Main Results"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

outcomes <- c(
  gpa_secondary  = "GPA",
  scored_lm      = "P(Take Exam)",
  prom_lm_actual = "Average LM",
  applied        = "Applied",
  assigned       = "Assigned",
  enrolled       = "Enrolled"
)

# helpers: table formatting
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# helper: collapse students to school x cohort classes (class means over the
# students passed in; num_students = their count)
class_panel <- function(students, cols = names(outcomes)) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

# helper: simple aggregated ATT table, coefficients with region-clustered
# bootstrap SEs, pre-treatment means over treated classes, one column per outcome
did_table <- function(res, means_pre, nobs, caption, label, file) {
  att <- map_dbl(res, "overall.att")
  se  <- map_dbl(res, "overall.se")
  p   <- 2 * pnorm(-abs(att / se))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", length(res)), "}\n      \\toprule\n",
    "                   & ", paste(names(res), collapse = " & "), "\\\\\n",
    "                   & ", paste0("(", seq_along(res), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", paste0(fmt(att), star(p))),
    row_tex("", paste0("(", fmt(se), ")")),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", fmt(means_pre)),
    row_tex("Observations", rep(format(nobs, big.mark = ","), length(res))),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

# helper: same table with two columns per outcome (e.g. female / male)
did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Student panel. Treatment: a school joins SAE the first year 9th grade in its
# region is assigned through the system; never-treated = private (DEPE 4).
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU,
             gpa_secondary, scored_lm, prom_lm_actual, applied, assigned, enrolled,
             assigned_full, assigned_not_full, assigned_top10, assigned_top25,
             assigned_top50, assigned_top75, assigned_top90)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## Headline table

Graduated high school on time, having spent all four years (grades 9 to 12) in
the same school. One observation per school x 9th-grade cohort; each outcome is
the average over the entire graduating class.

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
sel_flags <- c("assigned_full", "assigned_not_full", "assigned_top10", "assigned_top25",
               "assigned_top50", "assigned_top75", "assigned_top90")
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c(names(outcomes), sel_flags, "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]

dr <- class_panel(students)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  data = dr,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
models <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models) <- outcomes

means_pre <- dr[year_region_sae > 0 & year < year_region_sae,
                sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table(models, means_pre, nrow(dr), " Regression Results", "tab: main results",
              "tables/staggered_did_sae_on_outcomes_AM.tex"))
```

## Robustness: relaxed sample

Same design, relaxing the headline sample in two ways. First, students no
longer have to graduate on time: anyone who graduates counts, keyed to their
graduating class. Second, the four-years-in-the-same-school requirement is
relaxed to at most one publicly funded school in high school (so international
students and private-to-public movers enter; public-to-public switchers drop,
since there is no single class to attribute them to). Students entering 9th
grade before their school joined SAE do not count as treated.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
hs2 <- merge(hs, sch[, .(RBD = rbd, g_r = year_region_sae, elig_r = elig)], by = "RBD", all.x = TRUE)
hs2[is.na(g_r), g_r := 0]
hs2[is.na(elig_r), elig_r := 0]
setorder(hs2, MRUN, AGNO, COD_GRADO)
g12 <- hs2[COD_GRADO == 4, .(grad_year = AGNO[.N], grad_school = RBD[.N]), by = MRUN]
nine <- hs2[COD_GRADO == 1, .(anchor_rbd = RBD[1], AGNO_9th_actual = AGNO[1]), by = MRUN]
per2 <- hs2[, .(n_elig = uniqueN(RBD[elig_r == 1]),
                sae_touched = as.integer(any(g_r > 0 & (AGNO - COD_GRADO >= g_r)))), by = MRUN]
per2 <- merge(per2, g12, by = "MRUN")
per2 <- merge(per2, nine, by = "MRUN", all.x = TRUE)
per2 <- per2[n_elig <= 1]
vals2 <- df[MRUN %in% per2$MRUN,
            lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
            by = MRUN, .SDcols = c(names(outcomes), "graduated_secondary")]
students2 <- merge(per2, vals2, by = "MRUN")
students2 <- students2[graduated_secondary == 1]
students2 <- merge(students2, sch[, .(grad_school = rbd, g_g = year_region_sae)], by = "grad_school", all.x = TRUE)
students2 <- merge(students2, sch[, .(anchor_rbd = rbd, g_a = year_region_sae)], by = "anchor_rbd", all.x = TRUE)
students2[is.na(g_g), g_g := 0]
students2[is.na(g_a), g_a := 0]
students2[, grad_treated := as.integer(g_g > 0 & (grad_year - 4) >= g_g)]
students2 <- students2[!(grad_treated == 1 & g_a > 0 & !is.na(AGNO_9th_actual) &
                           (AGNO_9th_actual - 1) < g_a)]
students2 <- students2[!(grad_treated == 0 & sae_touched == 1)]
students2[, cohort := fifelse(!is.na(AGNO_9th_actual), AGNO_9th_actual, grad_year - 3)]
students2 <- students2[cohort >= start_year & cohort <= end_year]
students2[, rbd := grad_school]
students2[, AGNO_9th := grad_year - 3]

dr2 <- class_panel(students2)

common_args$data <- dr2
models2 <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models2) <- outcomes

means_pre2 <- dr2[year_region_sae > 0 & year < year_region_sae,
                  sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table(models2, means_pre2, nrow(dr2), "Regression Results: Relaxed Sample",
              "tab: robustness relaxed sample",
              "tables/staggered_did_sae_on_outcomes_filter2_AM.tex"))
```

## Heterogeneity: gender

Headline sample, class means computed separately over the female and male
students of each class (students with unknown sex are dropped); the control is
the subgroup class size.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female <- students[GEN_ALU == 2]
male   <- students[GEN_ALU == 1]
dr_f <- class_panel(female)
dr_m <- class_panel(male)

common_args$data <- dr_f
models_f <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_f) <- outcomes

common_args$data <- dr_m
models_m <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_m) <- outcomes

means_pre_f <- dr_f[year_region_sae > 0 & year < year_region_sae,
                    sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]
means_pre_m <- dr_m[year_region_sae > 0 & year < year_region_sae,
                    sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table_pair(models_f, models_m, "Female", "Male", means_pre_f, means_pre_m,
                   nrow(dr_f), nrow(dr_m), "Regression Results by Gender",
                   "tab: results by gender", "tables/did_by_gender_AM.tex"))
```

## Heterogeneity: school vulnerability

Treated schools split at the within-region median of the 2015 IVE-SINAE
vulnerability index, taken over treated schools (treated schools with no index
are dropped); never-treated private schools serve as the comparison group in
both arms.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
ee <- readRDS(file.path(data_path, "panel_ee_2004_2025_homogenized.rds"))
setDT(ee)
ive <- unique(ee[year == 2015, .(rbd = as.numeric(rbd), ive = as.numeric(ive_sinae_media))])
ive <- ive[!is.na(rbd)]
rm(ee)

schv <- unique(dr[, .(rbd = as.numeric(rbd), region, year_region_sae)])
schv <- merge(schv, ive, by = "rbd", all.x = TRUE)
med <- schv[year_region_sae > 0 & !is.na(ive), .(region_median = median(ive)), by = region]
schv <- merge(schv, med, by = "region", all.x = TRUE)
schv[, low  := as.integer(year_region_sae > 0 & !is.na(ive) & !is.na(region_median) & ive <  region_median)]
schv[, high := as.integer(year_region_sae > 0 & !is.na(ive) & !is.na(region_median) & ive >= region_median)]

dr_v <- merge(dr, schv[, .(rbd, low, high)], by = "rbd", all.x = TRUE)
dr_v[is.na(low), low := 0L]
dr_v[is.na(high), high := 0L]
dr_low  <- dr_v[year_region_sae == 0 | low  == 1]
dr_high <- dr_v[year_region_sae == 0 | high == 1]

common_args$data <- dr_low
models_low <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_low) <- outcomes

common_args$data <- dr_high
models_high <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_high) <- outcomes

means_pre_low <- dr_low[year_region_sae > 0 & year < year_region_sae,
                        sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]
means_pre_high <- dr_high[year_region_sae > 0 & year < year_region_sae,
                          sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table_pair(models_low, models_high, "Low", "High", means_pre_low, means_pre_high,
                   nrow(dr_low), nrow(dr_high), "Regression Results by School Vulnerability",
                   "tab: results by vulnerability", "tables/did_by_vulnerability_AM.tex"))
```

## Heterogeneity: program selectivity

Share of the graduating class assigned to a program, by the selectivity of the
program (programs that fill their seats, and cutoff-decile bins); each flag is
zero for students not assigned, so the shares are over the entire class.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
outcomes_sel <- c(
  assigned          = "All",
  assigned_full     = "Full",
  assigned_not_full = "Not Full",
  assigned_top10    = "Top 10",
  assigned_top25    = "Top 25",
  assigned_top50    = "Top 50",
  assigned_top75    = "Top 75",
  assigned_top90    = "Top 90"
)

dr_sel <- class_panel(students, names(outcomes_sel))

common_args$data <- dr_sel
models_sel <- map(names(outcomes_sel), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_sel) <- outcomes_sel

means_pre_sel <- dr_sel[year_region_sae > 0 & year < year_region_sae,
                        sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes_sel)]

cat(did_table(models_sel, means_pre_sel, nrow(dr_sel), "Regression Results by Program Selectivity",
              "tab: results by program selectivity",
              "tables/did_by_program_selectivity_AM.tex"))
```
