---
title: "Effect of SAE on SUA - Main Results"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)
library(fixest)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)
dir.create("figures", showWarnings = FALSE)

outcomes <- c(
  gpa_secondary  = "GPA",
  scored_lm      = "P(Take Exam)",
  prom_lm_actual = "Average LM",
  applied        = "Applied",
  assigned       = "Assigned",
  enrolled       = "Enrolled"
)

# helpers: table formatting
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# helper: collapse students to school x cohort classes (class means over the
# students passed in; num_students = their count)
class_panel <- function(students, cols = names(outcomes)) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

# helper: simple aggregated ATT table, coefficients with region-clustered
# bootstrap SEs, pre-treatment means over treated classes, one column per outcome
did_table <- function(res, means_pre, nobs, caption, label, file,
                      pre_label = "Mean (pre-treatment)") {
  att <- map_dbl(res, "overall.att")
  se  <- map_dbl(res, "overall.se")
  p   <- 2 * pnorm(-abs(att / se))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", length(res)), "}\n      \\toprule\n",
    "                   & ", paste(names(res), collapse = " & "), "\\\\\n",
    "                   & ", paste0("(", seq_along(res), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", paste0(fmt(att), star(p))),
    row_tex("", paste0("(", fmt(se), ")")),
    "      \\midrule\n",
    row_tex(pre_label, fmt(means_pre)),
    row_tex("Observations", rep(format(nobs, big.mark = ","), length(res))),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

# helper: same table with two columns per outcome (e.g. female / male)
did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

# helper: paired table in the circulated draft's layout ("Mean pre-treatment"
# row; htp! float; one multicolumn header line per outcome) -- used by the two
# value-added tables; generic in the number of outcomes
did_table_pair_draft <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                                 caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htp!]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\n   \\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n",
    "      ", paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean pre-treatment", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}\n   }}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

# helper: fit the CS-DiD once per outcome on a class panel and return the
# simple aggregates. `cols` is a named vector (panel column -> table label);
# the same seed inside every fit keeps results independent of run order.
fit_simple <- function(panel, cols) {
  common_args$data <- panel
  out <- map(names(cols), function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), common_args))
    suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
  })
  names(out) <- cols
  out
}

# helper: treated classes' own pre-period means over the given columns
pre_means <- function(panel, cols) {
  panel[year_region_sae > 0 & year < year_region_sae,
        sapply(.SD, mean, na.rm = TRUE), .SDcols = cols]
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Student panel. Treatment: a school joins SAE the first year 9th grade in its
# region is assigned through the system; never-treated = private (DEPE 4).
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             year_region_first_sae, graduated_secondary, COD_REG_RBD, COD_DEPE,
             COD_COM_RBD, COD_COM_ALU, SIT_FIN_R, GEN_ALU,
             gpa_secondary, scored_lm, prom_lm_actual, applied, assigned, enrolled,
             assigned_full, assigned_not_full, assigned_top10, assigned_top25,
             assigned_top50, assigned_top75, assigned_top90,
             assigned_selective_q10, assigned_selective_q20, assigned_selective_q30,
             assigned_selective_q40, assigned_selective_q50, assigned_selective_q60,
             assigned_selective_q70, assigned_selective_q80, assigned_selective_q90,
             assigned_selective_q100,
             num_apps_reg, num_valid_apps_reg, num_apps_above_prev_cutoff_reg)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## Headline table

Graduated high school on time, having spent all four years (grades 9 to 12) in
the same school. One observation per school x 9th-grade cohort; each outcome is
the average over the entire graduating class.

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
sel_flags <- c("assigned_full", "assigned_not_full", "assigned_top10", "assigned_top25",
               "assigned_top50", "assigned_top75", "assigned_top90")
dec_flags <- paste0("assigned_selective_q", seq(10, 100, 10))
app_counts <- c("num_apps_reg", "num_valid_apps_reg", "num_apps_above_prev_cutoff_reg")
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c(names(outcomes), sel_flags, dec_flags, app_counts,
                                  "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]

dr <- class_panel(students)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  data = dr,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
models <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models) <- outcomes

means_pre <- dr[year_region_sae > 0 & year < year_region_sae,
                sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table(models, means_pre, nrow(dr), " Regression Results", "tab: main results",
              "tables/staggered_did_sae_on_outcomes_AM.tex"))
```

## Robustness: relaxed sample

Same design, relaxing the headline sample in two ways. First, students no
longer have to graduate on time: anyone who graduates counts, keyed to their
graduating class. Second, the four-years-in-the-same-school requirement is
relaxed to at most one publicly funded school in high school (so international
students and private-to-public movers enter; public-to-public switchers drop,
since there is no single class to attribute them to). Students entering 9th
grade before their school joined SAE do not count as treated.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
hs2 <- merge(hs, sch[, .(RBD = rbd, g_r = year_region_sae, elig_r = elig)], by = "RBD", all.x = TRUE)
hs2[is.na(g_r), g_r := 0]
hs2[is.na(elig_r), elig_r := 0]
setorder(hs2, MRUN, AGNO, COD_GRADO)
g12 <- hs2[COD_GRADO == 4, .(grad_year = AGNO[.N], grad_school = RBD[.N]), by = MRUN]
nine <- hs2[COD_GRADO == 1, .(anchor_rbd = RBD[1], AGNO_9th_actual = AGNO[1]), by = MRUN]
per2 <- hs2[, .(n_elig = uniqueN(RBD[elig_r == 1]),
                sae_touched = as.integer(any(g_r > 0 & (AGNO - COD_GRADO >= g_r)))), by = MRUN]
per2 <- merge(per2, g12, by = "MRUN")
per2 <- merge(per2, nine, by = "MRUN", all.x = TRUE)
per2 <- per2[n_elig <= 1]
vals2 <- df[MRUN %in% per2$MRUN,
            lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
            by = MRUN, .SDcols = c(names(outcomes), "graduated_secondary")]
students2 <- merge(per2, vals2, by = "MRUN")
students2 <- students2[graduated_secondary == 1]
students2 <- merge(students2, sch[, .(grad_school = rbd, g_g = year_region_sae)], by = "grad_school", all.x = TRUE)
students2 <- merge(students2, sch[, .(anchor_rbd = rbd, g_a = year_region_sae)], by = "anchor_rbd", all.x = TRUE)
students2[is.na(g_g), g_g := 0]
students2[is.na(g_a), g_a := 0]
students2[, grad_treated := as.integer(g_g > 0 & (grad_year - 4) >= g_g)]
students2 <- students2[!(grad_treated == 1 & g_a > 0 & !is.na(AGNO_9th_actual) &
                           (AGNO_9th_actual - 1) < g_a)]
students2 <- students2[!(grad_treated == 0 & sae_touched == 1)]
students2[, cohort := fifelse(!is.na(AGNO_9th_actual), AGNO_9th_actual, grad_year - 3)]
students2 <- students2[cohort >= start_year & cohort <= end_year]
students2[, rbd := grad_school]
students2[, AGNO_9th := grad_year - 3]

dr2 <- class_panel(students2)

common_args$data <- dr2
models2 <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models2) <- outcomes

means_pre2 <- dr2[year_region_sae > 0 & year < year_region_sae,
                  sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table(models2, means_pre2, nrow(dr2), "Regression Results: Relaxed Sample",
              "tab: robustness relaxed sample",
              "tables/staggered_did_sae_on_outcomes_filter2_AM.tex"))
```

## Heterogeneity: gender

Headline sample, class means computed separately over the female and male
students of each class (students with unknown sex are dropped); the control is
the subgroup class size.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female <- students[GEN_ALU == 2]
male   <- students[GEN_ALU == 1]
dr_f <- class_panel(female)
dr_m <- class_panel(male)

common_args$data <- dr_f
models_f <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_f) <- outcomes

common_args$data <- dr_m
models_m <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_m) <- outcomes

means_pre_f <- dr_f[year_region_sae > 0 & year < year_region_sae,
                    sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]
means_pre_m <- dr_m[year_region_sae > 0 & year < year_region_sae,
                    sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table_pair(models_f, models_m, "Female", "Male", means_pre_f, means_pre_m,
                   nrow(dr_f), nrow(dr_m), "Regression Results by Gender",
                   "tab: results by gender", "tables/did_by_gender_AM.tex"))
```

## Heterogeneity: school vulnerability

Treated schools split at the within-region median of the 2015 IVE-SINAE
vulnerability index, taken over treated schools (treated schools with no index
are dropped); never-treated private schools serve as the comparison group in
both arms.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
ee <- readRDS(file.path(data_path, "panel_ee_2004_2025_homogenized.rds"))
setDT(ee)
ive <- unique(ee[year == 2015, .(rbd = as.numeric(rbd), ive = as.numeric(ive_sinae_media))])
ive <- ive[!is.na(rbd)]
rm(ee)

schv <- unique(dr[, .(rbd = as.numeric(rbd), region, year_region_sae)])
schv <- merge(schv, ive, by = "rbd", all.x = TRUE)
med <- schv[year_region_sae > 0 & !is.na(ive), .(region_median = median(ive)), by = region]
schv <- merge(schv, med, by = "region", all.x = TRUE)
schv[, low  := as.integer(year_region_sae > 0 & !is.na(ive) & !is.na(region_median) & ive <  region_median)]
schv[, high := as.integer(year_region_sae > 0 & !is.na(ive) & !is.na(region_median) & ive >= region_median)]

dr_v <- merge(dr, schv[, .(rbd, low, high)], by = "rbd", all.x = TRUE)
dr_v[is.na(low), low := 0L]
dr_v[is.na(high), high := 0L]
dr_low  <- dr_v[year_region_sae == 0 | low  == 1]
dr_high <- dr_v[year_region_sae == 0 | high == 1]

common_args$data <- dr_low
models_low <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_low) <- outcomes

common_args$data <- dr_high
models_high <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_high) <- outcomes

means_pre_low <- dr_low[year_region_sae > 0 & year < year_region_sae,
                        sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]
means_pre_high <- dr_high[year_region_sae > 0 & year < year_region_sae,
                          sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table_pair(models_low, models_high, "Low", "High", means_pre_low, means_pre_high,
                   nrow(dr_low), nrow(dr_high), "Regression Results by School Vulnerability",
                   "tab: results by vulnerability", "tables/did_by_vulnerability_AM.tex"))
```

## Heterogeneity: program selectivity

Share of the graduating class assigned to a program, by the selectivity of the
program (programs that fill their seats, and cutoff-decile bins); each flag is
zero for students not assigned, so the shares are over the entire class.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
outcomes_sel <- c(
  assigned          = "All",
  assigned_full     = "Full",
  assigned_not_full = "Not Full",
  assigned_top10    = "Top 10",
  assigned_top25    = "Top 25",
  assigned_top50    = "Top 50",
  assigned_top75    = "Top 75",
  assigned_top90    = "Top 90"
)

dr_sel <- class_panel(students, names(outcomes_sel))

common_args$data <- dr_sel
models_sel <- map(names(outcomes_sel), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_sel) <- outcomes_sel

means_pre_sel <- dr_sel[year_region_sae > 0 & year < year_region_sae,
                        sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes_sel)]

cat(did_table(models_sel, means_pre_sel, nrow(dr_sel), "Regression Results by Program Selectivity",
              "tab: results by program selectivity",
              "tables/did_by_program_selectivity_AM.tex"))
```

## Heterogeneity: school value added

Treated schools split at the within-region median of school value added, taken
over treated schools (treated schools with no measure are dropped);
never-treated private schools are the common control in both arms. The main
measure is the SES-adjusted value added on the 2016 10th-grade cohort: the
school fixed effect from a regression of standardized 10th-grade mathematics on
standardized 8th-grade mathematics and reading, sex, mother's education and
family income (both from the parent questionnaire), grade repetition, and
residential commune. The robustness measure pools the 2010--2016 cohorts and
keeps only prior achievement and sex.

```{r echo=FALSE, warning=FALSE, message=FALSE}
zscore_by <- function(x, g) {
  x <- as.numeric(x)
  (x - ave(x, g, FUN = function(v) mean(v, na.rm = TRUE))) /
    ave(x, g, FUN = function(v) sd(v, na.rm = TRUE))
}
to_na099 <- function(x) { x <- suppressWarnings(as.numeric(x)); x[x %in% c(0, 99, 999)] <- NA; x }

m2s <- readRDS(file.path(data_path, "panel_simce_2m.rds"))
setDT(m2s)
m2s <- m2s[agno >= 2010 & agno <= 2016 & !is.na(mrun) & !is.na(rbd),
           .(mrun = as.numeric(mrun), idalumno = as.numeric(idalumno),
             rbd = as.numeric(rbd), agno, gen_alu, mate2m = ptje_mate2m_alu)]
b8 <- readRDS(file.path(data_path, "panel_simce_8b.rds"))
setDT(b8)
b8 <- b8[!is.na(mrun), .(mrun = as.numeric(mrun), agno8b = agno,
                         mate8b = ptje_mate8b_alu, lect8b = ptje_lect8b_alu)]
setorder(b8, mrun, agno8b)
b8 <- b8[, .SD[1L], by = mrun]

# multi-cohort measure: earliest 10th-grade record per student, 2010-2016
m2m <- copy(m2s)
setorder(m2m, mrun, agno)
m2m <- m2m[, .SD[1L], by = mrun]
m2m[, female := as.integer(gen_alu %in% c("2", "Mujeres"))]
dva <- merge(m2m, b8, by = "mrun", all.x = TRUE)
dva[, `:=`(mate2m_z = zscore_by(mate2m, agno), mate8b_z = zscore_by(mate8b, agno8b),
           lect8b_z = zscore_by(lect8b, agno8b))]
# fixef.rm = "none": keep singleton-FE schools (one estimable student) so the
# value-added school set matches the certified split; newer fixest drops them
fe_multi <- fixef(feols(mate2m_z ~ mate8b_z + lect8b_z + female | rbd,
                        data = dva, fixef.rm = "none", notes = FALSE, warn = FALSE))$rbd
va_multi <- data.table(rbd = as.numeric(names(fe_multi)), va = as.numeric(fe_multi))

# SES-adjusted measure: the 2016 cohort, with mother's education and income
# from the parent questionnaire, ever-repeated and residential commune from the
# student panel; complete case on all controls
m2g <- m2s[agno == 2016]
setorder(m2g, mrun)
m2g <- m2g[, .SD[1L], by = mrun]
m2g[, female := as.integer(gen_alu %in% c("2", "Mujeres"))]
sub <- df[as.numeric(MRUN) %in% m2g$mrun]
rep_ever <- sub[, .(repeated = as.integer(any(SIT_FIN_R == "R", na.rm = TRUE))),
                by = .(mrun = as.numeric(MRUN))]
com16 <- sub[AGNO == 2016 & !is.na(COD_COM_ALU),
             .(comuna_stu = as.integer(COD_COM_ALU[1L])), by = .(mrun = as.numeric(MRUN))]
comfb <- sub[!is.na(COD_COM_ALU),
             .(comuna_fb = as.integer(COD_COM_ALU[1L])), by = .(mrun = as.numeric(MRUN))]
covs <- merge(comfb, com16, by = "mrun", all = TRUE)
covs[, comuna_stu := fifelse(is.na(comuna_stu), comuna_fb, comuna_stu)]
covs <- merge(covs[, .(mrun, comuna_stu)], rep_ever, by = "mrun", all = TRUE)
cp <- fread(file.path(data_path, "simce2m2016_cpad_privada_final_SEG.csv"), sep = "|",
            select = c("idalumno", "cpad_p08", "cpad_p10"), showProgress = FALSE)
cp[, `:=`(idalumno = as.numeric(idalumno),
          educ_madre = to_na099(cpad_p08), income = to_na099(cpad_p10))]
stopifnot(!anyDuplicated(cp$idalumno))   # a duplicate would row-explode the join
dG <- merge(m2g, b8, by = "mrun", all.x = TRUE)
dG <- merge(dG, covs, by = "mrun", all.x = TRUE)
dG <- merge(dG, cp[, .(idalumno, educ_madre, income)], by = "idalumno", all.x = TRUE)
dG[, `:=`(mate2m_z = zscore_by(mate2m, agno), mate8b_z = zscore_by(mate8b, agno8b),
          lect8b_z = zscore_by(lect8b, agno8b))]
dG <- dG[!is.na(mate2m_z) & !is.na(mate8b_z) & !is.na(lect8b_z) & !is.na(female) &
           !is.na(educ_madre) & !is.na(income) & !is.na(repeated) & !is.na(comuna_stu)]
fe_gaz <- fixef(feols(mate2m_z ~ mate8b_z + lect8b_z + female + i(educ_madre) +
                        i(income) + repeated + i(comuna_stu) | rbd,
                      data = dG, fixef.rm = "none", notes = FALSE, warn = FALSE))$rbd
va_gaz <- data.table(rbd = as.numeric(names(fe_gaz)), va = as.numeric(fe_gaz))
rm(m2s, m2m, m2g, b8, dva, dG, sub, rep_ever, com16, comfb, covs, cp)

# split treated schools at the within-region median over treated schools
va_split <- function(vadt) {
  s <- unique(dr[, .(rbd = as.numeric(rbd), region, year_region_sae)])
  s <- merge(s, vadt, by = "rbd", all.x = TRUE)
  med <- s[year_region_sae > 0 & !is.na(va), .(region_median = median(va)), by = region]
  s <- merge(s, med, by = "region", all.x = TRUE)
  s[, low  := as.integer(year_region_sae > 0 & !is.na(va) & !is.na(region_median) & va <  region_median)]
  s[, high := as.integer(year_region_sae > 0 & !is.na(va) & !is.na(region_median) & va >= region_median)]
  s[, .(rbd, low, high)]
}
va_arms <- function(vadt) {
  dv <- merge(dr, va_split(vadt), by = "rbd", all.x = TRUE)
  dv[is.na(low), low := 0L]
  dv[is.na(high), high := 0L]
  list(low = dv[year_region_sae == 0 | low == 1],
       high = dv[year_region_sae == 0 | high == 1])
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
arms_g <- va_arms(va_gaz)
models_vg_low  <- fit_simple(arms_g$low,  outcomes)
models_vg_high <- fit_simple(arms_g$high, outcomes)
cat(did_table_pair_draft(models_vg_low, models_vg_high, "Low", "High",
                    pre_means(arms_g$low, names(outcomes)),
                    pre_means(arms_g$high, names(outcomes)),
                    nrow(arms_g$low), nrow(arms_g$high),
                    "Effect on Outcomes by School Value Added",
                    "tab: did by value added", "tables/did_by_value_added_AM.tex"))
```

## Robustness: school value added, multi-cohort measure

Same split with the multi-cohort (2010--2016) value added that keeps only prior
achievement and sex.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
arms_m <- va_arms(va_multi)
models_vm_low  <- fit_simple(arms_m$low,  outcomes)
models_vm_high <- fit_simple(arms_m$high, outcomes)
cat(did_table_pair_draft(models_vm_low, models_vm_high, "Low", "High",
                    pre_means(arms_m$low, names(outcomes)),
                    pre_means(arms_m$high, names(outcomes)),
                    nrow(arms_m$low), nrow(arms_m$high),
                    "Effect on Outcomes by School Value Added (robustness: multi-cohort measure with prior achievement only)",
                    "tab: did by value added robust", "tables/did_by_value_added_robust_AM.tex"))
```

## Parallel trends: event study

Dynamic ATT off the headline panel, same fits and configuration as the headline
table. Calendar event time; the reference period is event time $-3$; the
anticipation band ($-2$, $-1$) is set aside in estimation but displayed. Bold
marks cells outside the 95% uniform (simultaneous) confidence band from the
region-clustered multiplier bootstrap. The joint pre-trend $p$-values are the
randomization-inference $p$ over the clean leads, pinned from the 1,001
exact rollout permutations (explorations/2026-07-10_randomization_inference,
`ri_full_pvalues.csv`); they are not recomputed at knit time.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# deliberate refit of the headline models (same seed -> identical fits): the
# headline chunk keeps only the simple aggregates, and we leave it untouched
pt_args <- common_args
pt_args$data <- dr
es <- map(names(outcomes), function(y) {
  set.seed(20260610)
  m  <- do.call(att_gt, c(list(yname = y), pt_args))
  ed <- suppressWarnings(aggte(m, type = "dynamic", na.rm = TRUE))
  data.table(e = ed$egt, att = ed$att.egt, se = ed$se.egt,
             bold = abs(ed$att.egt) > ed$crit.val.egt * ed$se.egt)
})
names(es) <- names(outcomes)

# randomization-inference joint pre-trend p over the clean leads (1,001 exact
# rollout permutations; explorations/2026-07-10_randomization_inference,
# ri_full_pvalues.csv). The RI run predates P(Take Exam)'s promotion to the
# outcome set, so that cell is blank until ri_full.R is extended to scored_lm.
ri_pretrend <- c(gpa_secondary = 0.464, scored_lm = NA_real_,
                 prom_lm_actual = 0.746, applied = 0.836,
                 assigned = 0.680, enrolled = 0.558)

es_cell <- function(y, k) {
  r <- es[[y]][e == k]
  if (nrow(r) == 0) return("")
  x <- fmt(r$att)
  if (isTRUE(r$bold)) x <- paste0("\\textbf{", x, "}")
  x
}
es_se <- function(y, k) {
  r <- es[[y]][e == k]
  if (nrow(r) == 0) "" else paste0("(", fmt(r$se), ")")
}
es_rows <- function(ks) paste0(vapply(ks, function(k) paste0(
  row_tex(paste("Event time", k), vapply(names(outcomes), es_cell, "", k = k)),
  row_tex("", vapply(names(outcomes), es_se, "", k = k))), ""), collapse = "")
span_row <- function(txt) paste0("      \\multicolumn{", length(outcomes) + 1,
                                 "}{l}{\\textit{", txt, "}}\\\\\n")

tab_es <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Parallel Trends Assumption: Event-Study Average ATT}",
  "\\label{tab: parallel trends assumption}\n",
  "   \\centerline{\\scalebox{0.85}{\n   \\begin{tabular}{l", strrep("c", length(outcomes)), "}\n      \\toprule\n",
  "                   & ", paste(outcomes, collapse = " & "), "\\\\\n",
  "                   & ", paste0("(", seq_along(outcomes), ")", collapse = " & "), "\\\\\n      \\midrule\n",
  span_row("Clean pre-treatment leads"),
  es_rows(-8:-4),
  span_row("Reference period: event time $-3$ (normalized to zero)"),
  span_row("Transition band (set aside in estimation; partial exposure by construction)"),
  es_rows(-2:-1),
  "      \\midrule\n",
  span_row("Post-treatment"),
  es_rows(0:4),
  "      \\midrule\n",
  row_tex("Joint pre-trend $p$-value",
          vapply(names(outcomes), function(y) {
            v <- ri_pretrend[[y]]
            if (is.na(v)) "" else fmt(v)
          }, "")),
  "      \\bottomrule\n   \\end{tabular}\n   }}\n\\end{table}\n")
writeLines(tab_es, "tables/parallel_trends_sae_on_outcomes_AM.tex")
cat(tab_es)
```

## Placebo: only private schools

The headline design run on the never-treated private classes alone,
pseudo-treated at their region's 9th-grade door. Private classes in the
Metropolitan region keep no door and serve as the never-treated comparison, so
late pseudo-cohorts still have valid comparisons under the two-year
anticipation band.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
dr_pl <- dr[year_region_sae == 0]
doors <- unique(dr[year_region_sae > 0, .(region, door = year_region_sae)])
stopifnot(!anyDuplicated(doors$region))
dr_pl <- merge(dr_pl, doors, by = "region", all.x = TRUE)
dr_pl[, year_region_sae := fifelse(region == 13, 0, door)]
dr_pl[, door := NULL]
stopifnot(!anyNA(dr_pl$year_region_sae))   # every private class needs a door or the Metro exemption

models_pl <- fit_simple(dr_pl, outcomes)
cat(did_table(models_pl, pre_means(dr_pl, names(outcomes)), nrow(dr_pl),
              "Placebo Test: Only Private Schools", "tab: did only private",
              "tables/staggered_did_sae_on_outcomes_only_private_AM.tex"))
```

## Mechanism: segregation by academic ability

Segregation of students by 4th-grade SIMCE ability across the schools of a
commune, on the full 9th-grade entrant population (not the stayer sample): the
Duncan dissimilarity index, low-to-high exposure, and low-to-low isolation, per
subject. The ability tag is the student's earliest 4th-grade high/low mark,
fixed before high school. Commune-cohort cells need both subjects defined and
at least 10 tagged students in each; the unit is the commune and the class-size
control is the commune's tagged count in the outcome's subject.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
com_lk <- df[!is.na(COD_COM_RBD), .(commune = as.integer(COD_COM_RBD[1L])), by = .(rbd = RBD)]
entr <- anchor[AGNO_9th >= start_year & AGNO_9th <= end_year]
entr <- merge(entr, sch[, .(rbd, region, year_region_sae)], by = "rbd", all.x = TRUE)
entr <- merge(entr, com_lk, by = "rbd", all.x = TRUE)

s4b <- readRDS(file.path(data_path, "panel_simce_4b.rds"))
setDT(s4b)
setorder(s4b, mrun, year)
tag <- s4b[, .(high_lect = { v <- high_lect4b[!is.na(high_lect4b)]; if (length(v)) as.numeric(v[1L]) else NA_real_ },
               high_mate = { v <- high_mate4b[!is.na(high_mate4b)]; if (length(v)) as.numeric(v[1L]) else NA_real_ }),
           by = .(MRUN = as.numeric(mrun))]
rm(s4b)
entr[, MRUN := as.numeric(MRUN)]
entr <- merge(entr, tag, by = "MRUN", all.x = TRUE)
entr[, `:=`(low_lect = 1 - high_lect, low_mate = 1 - high_mate)]

# across-school indices within (commune, cohort); untagged students are
# dropped, never counted as low; cells with no high/low split are dropped
seg_index <- function(tg, hi, lo) {
  schl <- tg[, .(h_s = sum(get(hi)), l_s = sum(get(lo))), by = .(commune, AGNO_9th, rbd)]
  schl[, n_s := h_s + l_s]
  out <- schl[, {
    H <- sum(h_s); L <- sum(l_s)
    .(duncan   = if (H > 0 && L > 0) 0.5 * sum(abs(h_s / H - l_s / L)) else NA_real_,
      p_lh     = if (L > 0) sum(fifelse(n_s > 0, (l_s / L) * (h_s / n_s), 0)) else NA_real_,
      p_ll     = if (L > 0) sum(fifelse(n_s > 0, (l_s / L) * (l_s / n_s), 0)) else NA_real_,
      n_tagged = sum(n_s))
  }, by = .(commune, AGNO_9th)]
  out[!is.na(duncan)]
}
seg_m <- seg_index(entr[!is.na(high_mate)], "high_mate", "low_mate")
seg_l <- seg_index(entr[!is.na(high_lect)], "high_lect", "low_lect")
setnames(seg_m, c("duncan", "p_lh", "p_ll", "n_tagged"),
         c("duncan_mate", "p_lh_mate", "p_ll_mate", "nt_mate"))
setnames(seg_l, c("duncan", "p_lh", "p_ll", "n_tagged"),
         c("duncan_lect", "p_lh_lect", "p_ll_lect", "nt_lect"))
seg <- merge(seg_m, seg_l, by = c("commune", "AGNO_9th"), all = TRUE)

modal <- function(x) { x <- x[!is.na(x)]; if (!length(x)) return(NA_real_)
  as.numeric(names(which.max(table(x)))) }
attr_c <- entr[, .(region = modal(region), year_region_sae = modal(year_region_sae)),
               by = commune]
seg <- merge(seg, attr_c, by = "commune", all.x = TRUE)
seg[, year := AGNO_9th - 1]
seg[year_region_sae > max(year, na.rm = TRUE), year_region_sae := 0]
seg <- seg[is.finite(duncan_mate) & is.finite(duncan_lect) &
             nt_mate >= 10 & nt_lect >= 10]

seg_args <- common_args
seg_args$idname <- "commune"
seg_cols <- c("duncan_mate", "duncan_lect", "p_lh_mate", "p_lh_lect",
              "p_ll_mate", "p_ll_lect")
models_seg <- map(seg_cols, function(y) {
  nt <- if (grepl("mate$", y)) "nt_mate" else "nt_lect"
  d <- seg[, c("commune", "year", "year_region_sae", "region", nt, y), with = FALSE]
  setnames(d, nt, "num_students")
  seg_args$data <- d
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), seg_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_seg) <- seg_cols

att_seg <- map_dbl(models_seg, "overall.att")
se_seg  <- map_dbl(models_seg, "overall.se")
p_seg   <- 2 * pnorm(-abs(att_seg / se_seg))
pre_seg <- pre_means(seg, seg_cols)
tab_seg <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect on Segregation by Ability}\\label{tab: did segregation}\n",
  "   \\centerline{\\scalebox{0.85}{\n\\begin{tabular}{lcccccc}\n\\toprule\n",
  "      & \\multicolumn{2}{c}{Duncan Index} & \\multicolumn{2}{c}{Exposure Low-High ($P_{LH}$)}",
  " & \\multicolumn{2}{c}{Isolation Low-Low ($P_{LL}$)} \\\\\n",
  "      \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7}\n",
  "                   & Math & Verbal & Math & Verbal & Math & Verbal\\\\\n",
  "                   & (1) & (2) & (3) & (4) & (5) & (6)\\\\\n      \\midrule\n",
  row_tex("Treated", paste0(fmt(att_seg), star(p_seg))),
  row_tex("", paste0("(", fmt(se_seg), ")")),
  "      \\midrule\n",
  row_tex("Mean pre-treatment", fmt(pre_seg)),
  row_tex("Observations", rep(format(nrow(seg), big.mark = ","), length(seg_cols))),
  "      \\bottomrule\n\\end{tabular}\n   }}\n\\end{table}\n")
writeLines(tab_seg, "tables/did_segregation_AM.tex")
cat(tab_seg)
```

## Mechanism: truthful reporting

OLS on the 2022 DEMRE admission-survey cross-section (not a DiD): survey
completers with fewer than 10 final applications, outside PACE, and outside the
RCT's information arms. Treated means graduating from a publicly funded school
in a region whose system started by the student's 9th-grade year (2018);
controls are Metropolitan-region public students and private-school students.
Strata and region/dependency fixed effects; standard errors clustered at the
school, significance from the normal reference.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
onset <- df[!is.na(year_region_first_sae) & year_region_first_sae > 0,
            .(onset = as.integer(median(year_region_first_sae))), by = .(region = COD_REG_RBD)]
d22 <- as.data.table(readRDS(file.path(data_path, "data_2022.rds")))
d22[, apps_fin_n := suppressWarnings(as.numeric(apps_fin))]
keep22 <- d22[in_survey == 1 & !is.na(apps_fin) & Progress == 100 &
                responded_survey == 1 & apps_fin_n < 10 & pace != "PACE"]
dov <- rbind(keep22[in_rct == 0], keep22[in_rct == 1 & treatment %in% c(1, 4)])
dov <- merge(dov, onset, by.x = "CODIGO_REGION_EGRESO", by.y = "region", all.x = TRUE)
dov[, treated := as.integer(DEPENDENCIA != 4 & onset <= 2018)]
dov_df <- as.data.frame(dov)

# cluster-robust (CR1) standard error for one coefficient, school clusters
cr1 <- function(m, cl, ti) {
  X <- model.matrix(m)[, !is.na(coef(m)), drop = FALSE]
  u <- residuals(m)
  n <- nrow(X); k <- ncol(X); G <- length(unique(cl))
  stopifnot(length(cl) == n, ti >= 1, ti <= k, G > 1, n > k)
  S <- rowsum(X * u, cl)
  a <- S %*% solve(crossprod(X))[, ti]
  sqrt(sum(a^2) * (G / (G - 1)) * ((n - 1) / (n - k)))
}
m2_fe   <- "+ factor(strata_score) + factor(strata_sexo) + factor(CODIGO_REGION_EGRESO) + factor(DEPENDENCIA)"
m2_outs <- c("applied_to_top_true_fin", "applied_to_bottom_true_fin",
             "prob_ratex_1_fin", "overall_ratex_fin",
             "bias_prob_top_reported", "bias_prob_overall")
m2 <- map(m2_outs, function(y) {
  m   <- lm(as.formula(paste(y, "~ treated", m2_fe)), data = dov_df)
  okc <- colnames(model.matrix(m))[!is.na(coef(m))]
  cl  <- as.character(dov_df$RBD[as.integer(rownames(m$model))])
  na_i <- which(is.na(cl))
  if (length(na_i)) cl[na_i] <- paste0("na", na_i)   # unknown school = own singleton cluster
  b  <- coef(m)[["treated"]]; b0 <- coef(m)[["(Intercept)"]]
  se_b <- cr1(m, cl, which(okc == "treated"))
  se_0 <- cr1(m, cl, which(okc == "(Intercept)"))
  list(att = b, se = se_b, p = 2 * pnorm(-abs(b / se_b)),
       const = b0, const_se = se_0, const_p = 2 * pnorm(-abs(b0 / se_0)), n = nobs(m))
})

m2row <- function(f) sapply(m2, `[[`, f)
tab_m2 <- paste0(
  "\\begin{table}[!htbp]\n\\caption{Effect on Truthful Reporting}\\label{tab: effect on truthful reporting}\n",
  "\\centerline{\\scalebox{0.85}{\n\\begin{tabular}{lcccccc}\n\\toprule\n",
  " & \\multicolumn{2}{c}{Truthful Reporting} & \\multicolumn{2}{c}{Adm. Probability} & \\multicolumn{2}{c}{Bias} \\\\\n",
  " \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7}\n",
  " & Top & Bottom & Top Reported & Overall & Top Reported & Overall \\\\\n",
  " & (1) & (2) & (3) & (4) & (5) & (6)\\\\\n\\midrule\n",
  " Treated & ",  paste0(fmt(m2row("att")), star(m2row("p")), collapse = " & "), " \\\\\n",
  " & ",         paste0("(", fmt(m2row("se")), ")",          collapse = " & "), " \\\\\n",
  " Constant & ", paste0(fmt(m2row("const")), star(m2row("const_p")), collapse = " & "), " \\\\\n",
  " & ",         paste0("(", fmt(m2row("const_se")), ")",    collapse = " & "), " \\\\\n\\midrule\n",
  " Observations & ", paste(format(m2row("n"), big.mark = ","), collapse = " & "), " \\\\\n",
  "\\bottomrule\n\\end{tabular}\n}}\n\\end{table}\n")
writeLines(tab_m2, "tables/effect_truthful_reporting_AM.tex")
cat(tab_m2)
```

## Mechanism: application breadth and quality

Unconditional application outcomes over the whole graduating class. Breadth:
the class share of graduates submitting at least $k$ applications (all
thresholds below 10, so they mean the same act in every year despite the list
cap change). Quality: class-level ratio-of-sums rates, valid applications and
applications above the previous-year cutoff over total applications; classes
with zero application volume have no defined rate and are dropped. The figure
plots the effect for every threshold $k = 1, \dots, 10$ with the 95% uniform
(sup-$t$) band from a region-clustered Mammen multiplier bootstrap with weights
shared across thresholds.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
thr_cols <- paste0("app_ge", 1:10)
for (k in 1:10) students[, (paste0("app_ge", k)) := as.integer(num_apps_reg >= k)]
dr_br <- class_panel(students, thr_cols)
models_br <- fit_simple(dr_br, setNames(thr_cols, thr_cols))

dr_rate <- students[, .(num_students = .N, s_apps = sum(num_apps_reg),
                        s_valid = sum(num_valid_apps_reg),
                        s_above = sum(num_apps_above_prev_cutoff_reg)),
                    by = .(rbd, AGNO_9th)]
dr_rate[, `:=`(valid_rate = s_valid / s_apps, above_rate = s_above / s_apps,
               year = AGNO_9th - 1)]
dr_rate <- merge(dr_rate, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
dr_rate[year_region_sae > max(year), year_region_sae := 0]
dr_rate <- dr_rate[s_apps > 0]
models_rate <- fit_simple(dr_rate, c(valid_rate = "Valid Apps", above_rate = "Above Cutoff"))

br_show <- c(models_br[c("app_ge1", "app_ge3", "app_ge5")], models_rate)
names(br_show) <- c("$\\geq 1$", "$\\geq 3$", "$\\geq 5$", "Valid Apps", "Above Cutoff")
att_br <- map_dbl(br_show, "overall.att")
se_br  <- map_dbl(br_show, "overall.se")
p_br   <- 2 * pnorm(-abs(att_br / se_br))
pre_br <- c(pre_means(dr_br, c("app_ge1", "app_ge3", "app_ge5")),
            pre_means(dr_rate, c("valid_rate", "above_rate")))
n_br   <- c(rep(nrow(dr_br), 3), rep(nrow(dr_rate), 2))
tab_br <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect on College Application Quality}\\label{tab: application quality}\n",
  "   \\centerline{\\scalebox{0.85}{\n\\begin{tabular}{lccccc}\n\\toprule\n",
  "      & \\multicolumn{3}{c}{Application Breadth} & \\multicolumn{2}{c}{Application Quality} \\\\\n",
  "      \\cmidrule(lr){2-4}\\cmidrule(lr){5-6}\n",
  "      & ", paste(names(br_show), collapse = " & "), " \\\\\n",
  "      & ", paste0("(", seq_along(br_show), ")", collapse = " & "), " \\\\\n\\midrule\n",
  row_tex("Treated", paste0(fmt(att_br), star(p_br))),
  row_tex("", paste0("(", fmt(se_br), ")")),
  "      \\midrule\n",
  row_tex("Mean (pre-treatment)", fmt(pre_br)),
  row_tex("Observations", format(n_br, big.mark = ",")),
  "\\bottomrule\n\\end{tabular}\n   }}\n\\end{table}\n")
writeLines(tab_br, "tables/did_app_breadth_quality_AM.tex")
cat(tab_br)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# uniform (sup-t) band across the 10 thresholds: cluster-mean influence
# functions from each fit, one Mammen weight vector per draw shared across
# thresholds (this is what makes the band simultaneous), critical value =
# 95th percentile of the max |t*|
psi_extract <- function(s) {
  inf <- as.numeric(s$inf.function$simple.att)
  dp  <- s$DIDparams
  dta <- as.data.frame(dp$data)
  dta <- dta[dta[[dp$tname]] == sort(unique(dta[[dp$tname]]))[1], ]
  cl  <- unique(dta[, c(dp$idname, dp$clustervars)])[, 2]
  stopifnot(length(cl) == length(inf))
  n_cl <- aggregate(cl, by = list(cl), length)[, 2]
  list(cmif = as.numeric(rowsum(inf, group = cl, reorder = TRUE) / n_cl),
       labs = as.integer(sort(unique(cl))))
}
psis <- map(models_br[thr_cols], psi_extract)
stopifnot(all(map_lgl(psis, function(p) identical(p$labs, psis[[1]]$labs))))
Psi  <- do.call(cbind, map(psis, "cmif"))
G_cl <- length(psis[[1]]$labs)

att_k <- map_dbl(models_br[thr_cols], "overall.att")
se_k  <- map_dbl(models_br[thr_cols], "overall.se")
mammen <- function(n) {
  k1 <- 0.5 * (1 - sqrt(5)); k2 <- 0.5 * (1 + sqrt(5))
  pk <- (sqrt(5) + 1) / (2 * sqrt(5))
  ifelse(runif(n) < pk, k1, k2)
}
set.seed(20260610)
dev_mat <- matrix(NA_real_, 10000, length(thr_cols))
for (b in seq_len(10000)) {
  V <- mammen(G_cl)
  dev_mat[b, ] <- as.numeric(crossprod(V, Psi)) / G_cl
}
# scaling self-check: the robust scale of the bootstrap deviations must
# reproduce each threshold's own bootstrap SE (within 10%); a failure means
# the influence-function extraction no longer matches the did internals and
# the band would be wrong -- stop before drawing anything
robust_scale <- function(b)
  (quantile(b, 0.75, type = 1, na.rm = TRUE) - quantile(b, 0.25, type = 1, na.rm = TRUE)) /
    (qnorm(0.75) - qnorm(0.25))
stopifnot(all(abs(apply(dev_mat, 2, robust_scale) / se_k - 1) < 0.10))
c_unif <- as.numeric(quantile(apply(abs(sweep(dev_mat, 2, se_k, "/")), 1, max),
                              0.95, type = 1, na.rm = TRUE))

lo <- att_k - c_unif * se_k
hi <- att_k + c_unif * se_k
pdf("figures/effect_uniform_errorbars.pdf", width = 7.2, height = 4.8)
par(mar = c(4.4, 4.8, 1, 1), family = "serif")
plot(1:10, att_k, type = "n", ylim = range(c(lo, hi)),
     xlab = expression(paste("Number of programs applied to (", italic(k), ")")),
     ylab = expression(paste("Effect on share applying to at least ", italic(k))),
     xaxt = "n")
axis(1, at = 1:10)
abline(h = 0, col = "grey55", lty = 2)
arrows(1:10, lo, 1:10, hi, length = 0.04, angle = 90, code = 3, col = "#1f77b4", lwd = 1.6)
points(1:10, att_k, pch = 19, col = "#1f77b4", cex = 1.1)
invisible(dev.off())
```

## Appendix: assignment to selective programs by quintile

Same unconditional class shares as the program-selectivity table, with the
selective (seat-filling) assignments partitioned into five selectivity
quintiles, least selective (Q1) to most selective (Q5); the quintiles sum to
the Full column.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
students[, `:=`(
  Q1 = assigned_selective_q10 + assigned_selective_q20,
  Q2 = assigned_selective_q30 + assigned_selective_q40,
  Q3 = assigned_selective_q50 + assigned_selective_q60,
  Q4 = assigned_selective_q70 + assigned_selective_q80,
  Q5 = assigned_selective_q90 + assigned_selective_q100)]
q_cols <- c(assigned = "All", assigned_full = "Full", assigned_not_full = "Not Full",
            Q1 = "Q1", Q2 = "Q2", Q3 = "Q3", Q4 = "Q4", Q5 = "Q5")
dr_q <- class_panel(students, names(q_cols))
models_q <- fit_simple(dr_q, q_cols)
cat(did_table(models_q, pre_means(dr_q, names(q_cols)), nrow(dr_q),
              "Effect on Assignment to Selective Programs by Quintile",
              "tab: effect admissions to selective programs by quintile",
              "tables/did_by_quantile_selectivity_AM.tex",
              pre_label = "Mean pre-treatment"))
```
