---
title: "Effect of SAE on SUA - PACE Robustness (Issue #11)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Checks whether the female-skewed SAE effects are confounded by PACE (Programa
de Acompañamiento y Acceso Efectivo), following issue #11. PACE guarantees
seats in the centralized system to top-15% students of covered
high-vulnerability schools; eligibility requires taking the entrance exam and
applying, so PACE moves the applied/assigned/enrolled outcomes mechanically,
and its benefits skew female (Carlana, Chiuri, Miglino and Tincani, 2026).
Three tests: (i) does PACE exposure itself load on the SAE rollout within our
design; (ii) does the gender effect survive when PACE-track admissions are
removed from the outcomes; (iii) does it survive when ever-PACE schools are
dropped. Sample, estimator, seed and table helpers are identical to
`3_effect_sae_am.Rmd`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

outcomes <- c(
  applied  = "Applied",
  assigned = "Assigned",
  enrolled = "Enrolled"
)
outcomes_xpace <- c(
  applied_xpace  = "Applied",
  assigned_xpace = "Assigned",
  enrolled_xpace = "Enrolled"
)
outcomes_pace <- c(
  pace_flag    = "Share PACE",
  pace_any     = "Any PACE",
  pace_any_imp = "Any PACE (imp.)"
)
pace_student_cols <- c("pace_flag", "applied_pace", "assigned_pace", "enrolled_pace",
                       names(outcomes_xpace))

# helpers: table formatting (verbatim from 3_effect_sae_am.Rmd)
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# helper: collapse students to school x cohort classes (class means over the
# students passed in; num_students = their count)
class_panel <- function(students, cols = names(outcomes)) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

# helper: simple aggregated ATT table, coefficients with region-clustered
# bootstrap SEs, pre-treatment means over treated classes, one column per outcome
did_table <- function(res, means_pre, nobs, caption, label, file) {
  att <- map_dbl(res, "overall.att")
  se  <- map_dbl(res, "overall.se")
  p   <- 2 * pnorm(-abs(att / se))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", length(res)), "}\n      \\toprule\n",
    "                   & ", paste(names(res), collapse = " & "), "\\\\\n",
    "                   & ", paste0("(", seq_along(res), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", paste0(fmt(att), star(p))),
    row_tex("", paste0("(", fmt(se), ")")),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", fmt(means_pre)),
    row_tex("Observations", rep(format(nobs, big.mark = ","), length(res))),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

# helper: same table with two columns per outcome (e.g. female / male)
did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Student panel. Treatment: a school joins SAE the first year 9th grade in its
# region is assigned through the system; never-treated = private (DEPE 4).
# Identical to 3_effect_sae_am.Rmd.
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU,
             gpa_secondary, scored_lm, prom_lm_actual, applied, assigned, enrolled)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## PACE variables

PACE flags come straight from the DEMRE student records, keyed to each
student's first admission process (same convention as the panel's
applied/assigned/enrolled in `1a_data_processing_primary_secondary.Rmd`):
`is_pace` marks students registered as PACE; `applied_pace`/`assigned_pace`
mark use of the PACE application track; enrollment through PACE is an
enrollment in the program the student was selected into via the PACE track.
The `x PACE` outcomes remove the PACE channel from the corresponding baseline
outcome: applications only on the PACE list, assignments through the PACE
track, and enrollments following a PACE-track assignment do not count. The
DEMRE PACE flag exists from the 2017 process onward, so the 69-school pilot
admissions of 2016 are not flagged; the diagnostics below print the counts by
year to check coverage against the program's expansion (69 schools in 2014,
456 by 2016; Carlana et al., 2026) and the 2023 change in the flag's coding.

```{r echo=FALSE, warning=FALSE, message=FALSE}
demre <- readRDS(file.path(data_path, "data_students_2004_2025.rds"))
setDT(demre)
demre <- demre[mrun %in% df$MRUN,
               .(mrun, year, pace, is_pace, applied, applied_reg, applied_pace,
                 assigned, assigned_reg, assigned_pace, enrolled, pref_seleccion_pace,
                 codigo_carrera_matricula, codigo_carrera_pace)]
setorder(demre, mrun, year)
demre <- unique(demre, by = "mrun")   # first admission process, as in 1a

demre[, pace_flag     := as.integer(coalesce(as.numeric(is_pace), 0) == 1)]
demre[, applied_reg   := as.integer(coalesce(as.numeric(applied_reg), 0) == 1)]
demre[, applied_pace  := as.integer(coalesce(as.numeric(applied_pace), 0) == 1)]
demre[, assigned_pace := as.integer(coalesce(as.numeric(assigned_pace), 0) == 1)]
demre[, enrolled_pace := as.integer(coalesce(as.numeric(enrolled), 0) == 1 & assigned_pace == 1 &
                                    !is.na(codigo_carrera_matricula) & !is.na(codigo_carrera_pace) &
                                    codigo_carrera_matricula == codigo_carrera_pace)]

# diagnostics: verify flag semantics before trusting the definitions below
cat("PACE-flagged students by admission process (first appearance):\n")
print(demre[pace_flag == 1, .N, by = year][order(year)])
cat("\nis_pace vs. raw pace label:\n")
print(demre[, table(pace_flag, pace_lbl = trimws(as.character(pace)) == "PACE", useNA = "ifany")])
cat("\nPACE-track use among flagged students:\n")
print(demre[pace_flag == 1, .(applied_pace = mean(applied_pace), assigned_pace = mean(assigned_pace),
                              enrolled_pace = mean(enrolled_pace))])
cat("\nassigned_pace outside the flagged group (should be ~0):",
    demre[pace_flag == 0, sum(assigned_pace)], "\n")
cat("\nassigned_pace vs. selected at some preference on the PACE track:\n")
print(demre[, table(assigned_pace, sel_pace = coalesce(as.numeric(pref_seleccion_pace), 0) > 0)])
cat("\nare PACE-track selections inside the baseline outcomes? (flagged students):\n")
print(demre[pace_flag == 1, table(assigned, assigned_pace, useNA = "ifany")])
print(demre[pace_flag == 1, table(applied, applied_pace, useNA = "ifany")])
cat("\nPACE-only applicants (applied_pace & !applied_reg) by year:\n")
print(demre[applied_pace == 1 & applied_reg == 0, .N, by = year][order(year)])
cat("\namong enrolled with a PACE-track assignment, share enrolled in that program:\n")
print(demre[assigned_pace == 1 & coalesce(as.numeric(enrolled), 0) == 1,
            mean(!is.na(codigo_carrera_matricula) & !is.na(codigo_carrera_pace) &
                 codigo_carrera_matricula == codigo_carrera_pace)])

df <- merge(df, demre[, .(MRUN = mrun, pace_flag, applied_reg, applied_pace,
                          assigned_pace, enrolled_pace)], by = "MRUN", all.x = TRUE)
rm(demre)
invisible(gc())

# xPACE outcomes: NA-ness mirrors the baseline outcome so class means compare
# like with like
df[, pace_flag := coalesce(pace_flag, 0L)]
df[, applied_xpace  := applied  * as.integer(!(coalesce(applied_pace, 0L) == 1 &
                                               coalesce(applied_reg, 0L) == 0))]
df[, assigned_xpace := assigned * (1L - coalesce(assigned_pace, 0L))]
df[, enrolled_xpace := enrolled * (1L - coalesce(enrolled_pace, 0L))]
df[, applied_pace   := coalesce(applied_pace, 0L)]
df[, assigned_pace  := coalesce(assigned_pace, 0L)]
df[, enrolled_pace  := coalesce(enrolled_pace, 0L)]
```

## Headline sample

Same sample as the headline table: graduated on time, all four grades in the
same school, cohorts entering 9th grade `r start_year`--`r end_year`.

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c(names(outcomes), pace_student_cols, "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]

dr <- class_panel(students, c(names(outcomes), pace_student_cols))
dr[, pace_any := as.integer(pace_flag > 0)]
fp <- dr[pace_any == 1, .(first_pace = min(year)), by = rbd]
dr <- merge(dr, fp, by = "rbd", all.x = TRUE)
# imputed exposure: schools stay PACE once entered (bridges the 2023 change in
# the DEMRE flag's coding)
dr[, pace_any_imp := as.integer(!is.na(first_pace) & year >= first_pace)]
```

## Does PACE exposure load on the SAE rollout?

The direct confounding test: same design and estimator as the headline table,
with the class's PACE exposure as the outcome. If PACE arrival is orthogonal
to SAE adoption, these ATTs are zero.

```{r echo=FALSE, warning=FALSE, message=FALSE}
cat("classes with any PACE-flagged student, by admission process (AGNO_9th + 4):\n")
print(dr[pace_any == 1, .(schools = uniqueN(rbd), classes = .N), by = .(adm_year = AGNO_9th + 4)][order(adm_year)])
cat("\nPACE exposure of treated classes by event time (year - year_region_sae):\n")
print(dr[year_region_sae > 0,
         .(share_any = mean(pace_any), share_any_imp = mean(pace_any_imp),
           share_students = mean(pace_flag), n = .N),
         by = .(event_time = year - year_region_sae)][order(event_time)])
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  data = dr,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
models_pace <- map(names(outcomes_pace), function(y) {
  set.seed(20260610)
  m <- do.call(att_gt, c(list(yname = y), common_args))
  suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
})
names(models_pace) <- outcomes_pace

means_pre_pace <- dr[year_region_sae > 0 & year < year_region_sae,
                     sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes_pace)]

cat(did_table(models_pace, means_pre_pace, nrow(dr), "Effect of SAE on PACE Exposure",
              "tab: pace exposure did", "tables/did_pace_exposure_AM.tex"))
```

## Gender balance of the PACE channel

```{r echo=FALSE, warning=FALSE, message=FALSE}
fem <- function(x) students[x == 1, mean(GEN_ALU == 2, na.rm = TRUE)]
cat("female share: sample =", fmt(students[, mean(GEN_ALU == 2, na.rm = TRUE)]),
    "| PACE-flagged =", fmt(fem(students$pace_flag)),
    "| applied via PACE =", fmt(fem(students$applied_pace)),
    "| assigned via PACE =", fmt(fem(students$assigned_pace)),
    "| enrolled via PACE =", fmt(fem(students$enrolled_pace)), "\n")
cat("\nshare of each student outcome accounted for by the PACE track\n",
    "(students in treated classes, post-treatment):\n")
tp <- dr[year_region_sae > 0 & year >= year_region_sae, .(rbd, AGNO_9th)]
stp <- students[tp, on = c("rbd", "AGNO_9th"), nomatch = 0]
print(stp[!is.na(GEN_ALU),
          .(applied = mean(applied, na.rm = TRUE), applied_pace_only = mean(applied - applied_xpace, na.rm = TRUE),
            assigned = mean(assigned, na.rm = TRUE), assigned_pace = mean(assigned - assigned_xpace, na.rm = TRUE),
            enrolled = mean(enrolled, na.rm = TRUE), enrolled_pace = mean(enrolled - enrolled_xpace, na.rm = TRUE)),
          by = .(female = as.integer(GEN_ALU == 2))])
```

## Heterogeneity by gender, removing the PACE channel

Class means computed separately over the female and male students of each
class, as in the gender table of `3_effect_sae_am.Rmd`; students with unknown
sex are dropped and the control is the subgroup class size. Baseline outcomes
replicate the draft's gender table (a check that this file reproduces it
exactly); the `x PACE` outcomes remove PACE-track applications, assignments
and enrollments. If the female effect is PACE, it shrinks toward zero here.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female <- students[GEN_ALU == 2]
male   <- students[GEN_ALU == 1]
dr_f <- class_panel(female, c(names(outcomes), names(outcomes_xpace)))
dr_m <- class_panel(male,   c(names(outcomes), names(outcomes_xpace)))

run_gender <- function(dat, ycols) {
  common_args$data <- dat
  res <- map(ycols, function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), common_args))
    suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
  })
  names(res) <- ycols
  res
}

models_f_base <- run_gender(dr_f, names(outcomes))
models_m_base <- run_gender(dr_m, names(outcomes))
models_f_x    <- run_gender(dr_f, names(outcomes_xpace))
models_m_x    <- run_gender(dr_m, names(outcomes_xpace))
names(models_f_x) <- names(models_m_x) <- outcomes_xpace

means_pre_f <- dr_f[year_region_sae > 0 & year < year_region_sae,
                    sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes_xpace)]
means_pre_m <- dr_m[year_region_sae > 0 & year < year_region_sae,
                    sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes_xpace)]

cat(did_table_pair(models_f_x, models_m_x, "Female", "Male", means_pre_f, means_pre_m,
                   nrow(dr_f), nrow(dr_m),
                   "Regression Results by Gender, Excluding the PACE Channel",
                   "tab: results by gender excl pace", "tables/did_by_gender_xpace_AM.tex"))
```

## Heterogeneity by gender, excluding ever-PACE schools

Belt and braces: drop every school with a PACE-flagged class at any point in
the sample (this removes the most vulnerable treated schools, so the estimand
changes; the comparison of interest is the female/male gap, not the level).

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
pace_schools <- dr[pace_any == 1, unique(rbd)]
cat(length(pace_schools), "ever-PACE schools dropped;",
    fmt(dr[year_region_sae > 0, mean(rbd %in% pace_schools)]),
    "of treated classes belong to them.\n\n")

dr_f_np <- class_panel(female[!rbd %in% pace_schools], names(outcomes))
dr_m_np <- class_panel(male[!rbd %in% pace_schools],   names(outcomes))

models_f_np <- run_gender(dr_f_np, names(outcomes))
models_m_np <- run_gender(dr_m_np, names(outcomes))
names(models_f_np) <- names(models_m_np) <- outcomes

means_pre_f_np <- dr_f_np[year_region_sae > 0 & year < year_region_sae,
                          sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]
means_pre_m_np <- dr_m_np[year_region_sae > 0 & year < year_region_sae,
                          sapply(.SD, mean, na.rm = TRUE), .SDcols = names(outcomes)]

cat(did_table_pair(models_f_np, models_m_np, "Female", "Male", means_pre_f_np, means_pre_m_np,
                   nrow(dr_f_np), nrow(dr_m_np),
                   "Regression Results by Gender, Excluding Ever-PACE Schools",
                   "tab: results by gender excl pace schools",
                   "tables/did_by_gender_excl_pace_schools_AM.tex"))
```

## Summary

```{r echo=FALSE, warning=FALSE, message=FALSE}
line <- function(res_f, res_m, y, lbl) {
  a <- function(r) sprintf("%7.3f (%.3f)", r[[y]]$overall.att, r[[y]]$overall.se)
  cat(sprintf("%-28s female %s | male %s\n", lbl, a(res_f), a(res_m)))
}
for (y in names(outcomes)) {
  cat("\n--", outcomes[[y]], "--\n")
  line(models_f_base, models_m_base, y, "baseline (replication)")
  line(models_f_x,    models_m_x,    outcomes_xpace[paste0(y, "_xpace")], "excl. PACE channel")
  line(models_f_np,   models_m_np,   outcomes[y], "excl. ever-PACE schools")
}
cat("\n-- SAE effect on PACE exposure --\n")
for (y in outcomes_pace) cat(sprintf("%-28s %7.3f (%.3f)\n", y,
                                     models_pace[[y]]$overall.att, models_pace[[y]]$overall.se))
cat("\nBaseline rows must match tables/did_by_gender_AM.tex; if they do not,\n",
    "the sample replication is off and the other rows are not comparable.\n")
```
