---
title: "Effect of SAE on SUA - Gender Mechanisms, Phase 1 (Issue #14)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Phase 1 of issue #14: where in the pipeline does the female effect originate,
and which of two channels moves — application behavior (frictions/uncertainty:
portfolio size, application mistakes, targeting of attainable programs) or
admission scores (NEM/ranking percentiles vs. LM). Three gender tables: (A)
application behavior, (B) admission-score percentiles, (C) upstream margins on
entrant classes (retention, on-time graduation), the only section whose sample
is not conditioned on graduating. Sample, estimator, seed and table helpers
are identical to `3_effect_sae_am.Rmd`; the `Applied` column of Table A must
match the draft's gender table (replication guard).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

outcomes_app <- c(
  scored_lm                      = "P(Take Exam)",
  applied                        = "Applied",
  num_apps_reg                   = "Num. Apps",
  pct_valid_reg                  = "Share Valid",
  pct_apps_above_prev_cutoff_reg = "Share Above Cutoff"
)
outcomes_sco <- c(
  nem_pct          = "NEM Pct.",
  ranking_pct      = "Ranking Pct.",
  nem_pct_cond     = "NEM Pct. (takers)",
  ranking_pct_cond = "Ranking Pct. (takers)",
  lm_pct_cond      = "LM Pct. (takers)"
)
outcomes_grad <- c(
  ret4         = "4 Years Same School",
  on_time_grad = "Graduated On Time",
  grad_ever    = "Graduated"
)
outcomes <- c(outcomes_app, outcomes_sco)

# helpers: table formatting (verbatim from 3_effect_sae_am.Rmd)
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# helper: collapse students to school x cohort classes (class means over the
# students passed in; num_students = their count)
class_panel <- function(students, cols = names(outcomes)) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

# helper: two columns per outcome (female / male), verbatim from 3_effect_sae_am.Rmd
did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Student panel. Treatment: a school joins SAE the first year 9th grade in its
# region is assigned through the system; never-treated = private (DEPE 4).
# Identical to 3_effect_sae_am.Rmd, keeping the application-behavior and
# score-percentile columns.
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU,
             scored_lm, applied, num_apps_reg, pct_valid_reg,
             pct_apps_above_prev_cutoff_reg, nem_pct, ranking_pct, lm_pct)]

# conditional versions: percentile among students with a computed score, and
# non-registrants excluded; the unconditional versions carry zeros
df[, nem_pct_cond     := fifelse(nem_pct > 0, nem_pct, NA_real_)]
df[, ranking_pct_cond := fifelse(ranking_pct > 0, ranking_pct, NA_real_)]
df[, lm_pct_cond      := fifelse(lm_pct > 0, lm_pct, NA_real_)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## Headline sample

Same sample as the headline table: graduated on time, all four grades in the
same school, cohorts entering 9th grade `r start_year`--`r end_year`; class
means computed separately over the female and male students of each class.

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c(names(outcomes), "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]

female <- students[GEN_ALU == 2]
male   <- students[GEN_ALU == 1]
dr_f <- class_panel(female, names(outcomes))
dr_m <- class_panel(male,   names(outcomes))

common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
run_gender <- function(dat, ycols) {
  common_args$data <- dat
  res <- map(ycols, function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), common_args))
    suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
  })
  names(res) <- ycols
  res
}

means_pre <- function(dat, ycols) {
  dat[year_region_sae > 0 & year < year_region_sae,
      sapply(.SD, mean, na.rm = TRUE), .SDcols = ycols]
}
```

## A. Application behavior

`Num. Apps` counts regular-track applications over all class members (zero for
non-applicants); `Share Valid` and `Share Above Cutoff` are conditional on
applying (share of the student's listed programs that are valid, and whose
previous-year cutoff the student's score clears), so they capture portfolio
quality among appliers. If application frictions bind more on women, the
female effect starts at the take-exam/apply margin and new applicants list
attainable programs.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
models_f_app <- run_gender(dr_f, names(outcomes_app))
models_m_app <- run_gender(dr_m, names(outcomes_app))
names(models_f_app) <- names(models_m_app) <- outcomes_app

cat(did_table_pair(models_f_app, models_m_app, "Female", "Male",
                   means_pre(dr_f, names(outcomes_app)), means_pre(dr_m, names(outcomes_app)),
                   nrow(dr_f), nrow(dr_m),
                   "Application Behavior by Gender",
                   "tab: gender application behavior", "tables/did_by_gender_app_behavior_AM.tex"))
```

## B. Admission-score percentiles

NEM and ranking percentiles are computed within admission year; the
unconditional versions assign zero to students without a computed score, so
they mix the registration margin with rank position. The `(takers)` versions
condition on having a score. If the rank-based channel operates, girls'
NEM/ranking percentiles rise while LM percentiles among takers do not.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
models_f_sco <- run_gender(dr_f, names(outcomes_sco))
models_m_sco <- run_gender(dr_m, names(outcomes_sco))
names(models_f_sco) <- names(models_m_sco) <- outcomes_sco

cat(did_table_pair(models_f_sco, models_m_sco, "Female", "Male",
                   means_pre(dr_f, names(outcomes_sco)), means_pre(dr_m, names(outcomes_sco)),
                   nrow(dr_f), nrow(dr_m),
                   "Admission-Score Percentiles by Gender",
                   "tab: gender score percentiles", "tables/did_by_gender_scores_AM.tex"))
```

## C. Upstream margins: entrant classes

One observation per school x 9th-grade entering cohort, all entrants counted
(the only sample in the paper not conditioned on graduating). `4 Years Same
School` is the headline sample's retention condition; `Graduated On Time`
requires reaching 12th grade at entry year + 3 and graduating; `Graduated`
drops the timing requirement. If SAE keeps girls on track through secondary
school, more of them reach the college margin.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
g12 <- hs[COD_GRADO == 4, .(g12_year = min(AGNO)), by = MRUN]
vals_e <- df[MRUN %in% anchor$MRUN,
             lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
             by = MRUN, .SDcols = c("graduated_secondary", "GEN_ALU")]
entrants <- merge(anchor, per[, .(MRUN, ng, nsch)], by = "MRUN")
entrants <- merge(entrants, vals_e, by = "MRUN", all.x = TRUE)
entrants <- merge(entrants, g12, by = "MRUN", all.x = TRUE)
entrants[, grad_ever    := as.integer(coalesce(graduated_secondary, 0) == 1)]
entrants[, on_time_grad := as.integer(grad_ever == 1 & !is.na(g12_year) & g12_year == AGNO_9th + 3)]
entrants[, ret4         := as.integer(ng == 4 & nsch == 1)]
entrants <- entrants[AGNO_9th >= start_year & AGNO_9th <= end_year]

female_e <- entrants[GEN_ALU == 2]
male_e   <- entrants[GEN_ALU == 1]
dr_f_e <- class_panel(female_e, names(outcomes_grad))
dr_m_e <- class_panel(male_e,   names(outcomes_grad))

models_f_grad <- run_gender(dr_f_e, names(outcomes_grad))
models_m_grad <- run_gender(dr_m_e, names(outcomes_grad))
names(models_f_grad) <- names(models_m_grad) <- outcomes_grad

cat(did_table_pair(models_f_grad, models_m_grad, "Female", "Male",
                   means_pre(dr_f_e, names(outcomes_grad)), means_pre(dr_m_e, names(outcomes_grad)),
                   nrow(dr_f_e), nrow(dr_m_e),
                   "Secondary-School Progression by Gender (Entrant Classes)",
                   "tab: gender progression entrants", "tables/did_by_gender_graduation_AM.tex"))
```

## Summary

```{r echo=FALSE, warning=FALSE, message=FALSE}
show <- function(res_f, res_m, labs) {
  for (y in labs) cat(sprintf("%-24s female %7.3f (%.3f) | male %7.3f (%.3f)\n", y,
                              res_f[[y]]$overall.att, res_f[[y]]$overall.se,
                              res_m[[y]]$overall.att, res_m[[y]]$overall.se))
}
cat("-- A. Application behavior --\n"); show(models_f_app, models_m_app, outcomes_app)
cat("\n-- B. Score percentiles --\n");  show(models_f_sco, models_m_sco, outcomes_sco)
cat("\n-- C. Entrant progression --\n"); show(models_f_grad, models_m_grad, outcomes_grad)
cat("\nGuard: the Applied column must match the draft gender table",
    "(female 0.078, male 0.020); if it does not, the sample replication is off.\n")
```
