---
title: "Effect of SAE on SUA - Gender Mechanisms, Phase 2: Sorting (Issue #14)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Phase 2 of issue #14: does SAE re-sort girls across schools? Three tests. (A)
Class gender composition: effect of SAE on the share of female students, with
treated schools split by academic vs vocational type and by pre-reform
college-pipeline intensity — if centralized choice moves girls toward
college-pipeline schools, their share rises there and falls elsewhere. (B)
Commuting: share of students enrolled outside their commune of residence in
9th grade, by gender — a feasible-set test. (C) Composition: average 8th-grade
SIMCE of the class's female and male students — measured before high school,
so it moves only through who sorts where. Sample, estimator, seed and helpers
identical to `3_effect_sae_am.Rmd`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

# helpers: table formatting (verbatim from 3_effect_sae_am.Rmd)
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

class_panel <- function(students, cols) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

did_table <- function(res, means_pre, nobs, caption, label, file) {
  att <- map_dbl(res, "overall.att")
  se  <- map_dbl(res, "overall.se")
  p   <- 2 * pnorm(-abs(att / se))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", length(res)), "}\n      \\toprule\n",
    "                   & ", paste(names(res), collapse = " & "), "\\\\\n",
    "                   & ", paste0("(", seq_along(res), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", paste0(fmt(att), star(p))),
    row_tex("", paste0("(", fmt(se), ")")),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", fmt(means_pre)),
    row_tex("Observations", format(nobs, big.mark = ",")),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Identical to 3_effect_sae_am.Rmd, keeping communes and school type (COD_ENSE).
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU,
             COD_COM_ALU, COD_COM_RBD, applied)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# school type: share of pre-2016 secondary enrollment in the academic track
# (COD_ENSE 310); majority rule
stype <- df[COD_ENSE %in% media & AGNO <= 2015, .(hc_share = mean(COD_ENSE == 310)), by = RBD]
setnames(stype, "RBD", "rbd")
stype[, hc := as.integer(hc_share >= 0.5)]

# commune of residence vs commune of the school at the 9th-grade record
com9 <- df[COD_ENSE %in% media & COD_GRADO == 1 & !is.na(COD_COM_ALU) & !is.na(COD_COM_RBD),
           .(MRUN, AGNO, diff_com = as.integer(COD_COM_ALU != COD_COM_RBD))]
setorder(com9, MRUN, AGNO)
com9 <- unique(com9, by = "MRUN")[, .(MRUN, diff_com)]

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## Headline sample

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c("applied", "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]
students <- merge(students, com9, by = "MRUN", all.x = TRUE)
students[, female := fifelse(GEN_ALU %in% c(1, 2), as.numeric(GEN_ALU == 2), NA_real_)]

common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
run_set <- function(dat, ycols) {
  common_args$data <- dat
  res <- map(ycols, function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), common_args))
    suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
  })
  names(res) <- ycols
  res
}

means_pre <- function(dat, ycols) {
  dat[year_region_sae > 0 & year < year_region_sae,
      sapply(.SD, mean, na.rm = TRUE), .SDcols = ycols]
}
```

## A. Class gender composition, by school type and pipeline intensity

One observation per school x cohort over the whole class; the outcome is the
share of female students. Treated schools are split (i) by academic (HC) vs
vocational (TP) type, classified by the majority track of their pre-2016
enrollment, and (ii) at the within-region median of pre-reform college-pipeline
intensity (mean application share of the school's 2012--2014 classes), taken
over treated schools; never-treated private schools serve as the comparison
group in every arm.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
cls <- students[!is.na(female)]
dr_all <- class_panel(cls, c("female", "applied"))

pipe <- dr_all[AGNO_9th <= 2014, .(pipe = mean(applied, na.rm = TRUE)), by = rbd]
schx <- unique(dr_all[, .(rbd, region, year_region_sae)])
schx <- merge(schx, pipe, by = "rbd", all.x = TRUE)
schx <- merge(schx, stype[, .(rbd, hc)], by = "rbd", all.x = TRUE)
medp <- schx[year_region_sae > 0 & !is.na(pipe), .(pmed = median(pipe)), by = region]
schx <- merge(schx, medp, by = "region", all.x = TRUE)
schx[, hi_pipe := as.integer(year_region_sae > 0 & !is.na(pipe) & !is.na(pmed) & pipe >= pmed)]
schx[, lo_pipe := as.integer(year_region_sae > 0 & !is.na(pipe) & !is.na(pmed) & pipe <  pmed)]

dr_x <- merge(dr_all, schx[, .(rbd, hc, hi_pipe, lo_pipe)], by = "rbd", all.x = TRUE)
arms <- list(
  "HC Schools"    = dr_x[year_region_sae == 0 | (year_region_sae > 0 & hc == 1)],
  "TP Schools"    = dr_x[year_region_sae == 0 | (year_region_sae > 0 & hc == 0)],
  "High Pipeline" = dr_x[year_region_sae == 0 | hi_pipe == 1],
  "Low Pipeline"  = dr_x[year_region_sae == 0 | lo_pipe == 1]
)
models_sf <- imap(arms, function(dat, nm) run_set(dat, "female")[[1]])
means_sf  <- map_dbl(arms, ~means_pre(.x, "female"))

cat(did_table(models_sf, means_sf, map_int(arms, nrow), "Effect on Share of Female Students",
              "tab: share female by school type", "tables/did_share_female_by_type_AM.tex"))
```

## B. Enrollment outside the commune of residence

Share of the class's female (male) students whose 9th-grade school is in a
different commune than their residence. If centralized choice relaxes
constraints that bind more for daughters, girls' out-of-commune enrollment
rises relative to boys'.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female_s <- students[GEN_ALU == 2]
male_s   <- students[GEN_ALU == 1]
dr_f <- class_panel(female_s, "diff_com")
dr_m <- class_panel(male_s,   "diff_com")

models_f_dc <- run_set(dr_f, "diff_com"); names(models_f_dc) <- "Outside Own Commune"
models_m_dc <- run_set(dr_m, "diff_com"); names(models_m_dc) <- "Outside Own Commune"

cat(did_table_pair(models_f_dc, models_m_dc, "Female", "Male",
                   means_pre(dr_f, "diff_com"), means_pre(dr_m, "diff_com"),
                   nrow(dr_f), nrow(dr_m),
                   "Enrollment Outside the Commune of Residence by Gender",
                   "tab: gender out of commune", "tables/did_by_gender_commune_AM.tex"))
```

## C. Baseline composition: 8th-grade SIMCE

Average 8th-grade SIMCE of the class's female (male) students, standardized
within test year. The score is measured before high school, so any effect
reflects who sorts into the school, not what the school does — separating
re-sorting from behavior. The 8th-grade test is not administered every year,
so only cohorts with a tested 8th-grade year contribute.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
s8 <- readRDS(file.path(data_path, "panel_simce_8b.rds"))
setDT(s8)
s8 <- s8[!is.na(mrun), .(MRUN = mrun, agno,
                         lect = suppressWarnings(as.numeric(ptje_lect8b_alu)),
                         mate = suppressWarnings(as.numeric(ptje_mate8b_alu)))]
s8 <- s8[!is.na(lect) | !is.na(mate)]
setorder(s8, MRUN, -agno)
s8 <- unique(s8, by = "MRUN")
s8[, z_lect := as.numeric(scale(lect)), by = agno]
s8[, z_mate := as.numeric(scale(mate)), by = agno]

students <- merge(students, s8[, .(MRUN, z_lect, z_mate)], by = "MRUN", all.x = TRUE)
female_s <- students[GEN_ALU == 2]
male_s   <- students[GEN_ALU == 1]
dr_f8 <- class_panel(female_s, c("z_mate", "z_lect"))
dr_m8 <- class_panel(male_s,   c("z_mate", "z_lect"))

outcomes_s8 <- c(z_mate = "SIMCE-8 Math", z_lect = "SIMCE-8 Reading")
models_f_s8 <- run_set(dr_f8, names(outcomes_s8)); names(models_f_s8) <- outcomes_s8
models_m_s8 <- run_set(dr_m8, names(outcomes_s8)); names(models_m_s8) <- outcomes_s8

cat(did_table_pair(models_f_s8, models_m_s8, "Female", "Male",
                   means_pre(dr_f8, names(outcomes_s8)), means_pre(dr_m8, names(outcomes_s8)),
                   nrow(dr_f8), nrow(dr_m8),
                   "Baseline 8th-Grade SIMCE Composition by Gender",
                   "tab: gender simce8 composition", "tables/did_by_gender_simce8_AM.tex"))
```

## Summary

```{r echo=FALSE, warning=FALSE, message=FALSE}
cat("-- A. Share female --\n")
for (nm in names(models_sf)) cat(sprintf("%-16s %7.3f (%.3f)\n", nm,
                                         models_sf[[nm]]$overall.att, models_sf[[nm]]$overall.se))
cat("\n-- B. Outside own commune --\n")
cat(sprintf("female %7.3f (%.3f) | male %7.3f (%.3f)\n",
            models_f_dc[[1]]$overall.att, models_f_dc[[1]]$overall.se,
            models_m_dc[[1]]$overall.att, models_m_dc[[1]]$overall.se))
cat("\n-- C. SIMCE-8 composition --\n")
for (y in outcomes_s8) cat(sprintf("%-16s female %7.3f (%.3f) | male %7.3f (%.3f)\n", y,
                                   models_f_s8[[y]]$overall.att, models_f_s8[[y]]$overall.se,
                                   models_m_s8[[y]]$overall.att, models_m_s8[[y]]$overall.se))
```
