---
title: "Effect of SAE on SUA - Gender Mechanisms, Phase 3: Destination Fields (Issue #14)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Phase 3 of issue #14: is the female effect a supply-composition story — new
admissions absorbed by female-typed programs? Programs are typed by their
female share among students selected in the 2014--2019 processes (before any
SAE-treated cohort reaches college admission): top tercile = female-typed,
bottom = male-typed, middle (and programs with fewer than 30 selected
students) = mixed. Assigned and enrolled are then partitioned by the
destination program's type, so the three columns add up to the baseline
outcome, and the decomposition reports how much of the female--male gap each
type accounts for. Sample, estimator, seed and helpers identical to
`3_effect_sae_am.Rmd`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

outcomes_asg <- c(asg_fem = "Female-typed", asg_mix = "Mixed", asg_mal = "Male-typed")
outcomes_enr <- c(enr_fem = "Female-typed", enr_mix = "Mixed", enr_mal = "Male-typed")

# helpers: table formatting (verbatim from 3_effect_sae_am.Rmd)
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

class_panel <- function(students, cols) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Identical to 3_effect_sae_am.Rmd, keeping the selected and enrolled program codes.
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU,
             applied, assigned, enrolled,
             codigo_carrera_seleccion, codigo_carrera_matricula)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## Program typing by pre-reform female share

```{r echo=FALSE, warning=FALSE, message=FALSE}
demre <- readRDS(file.path(data_path, "data_students_2004_2025.rds"))
setDT(demre)
fs <- demre[year %in% 2014:2019 & !is.na(codigo_carrera_seleccion) & codigo_carrera_seleccion > 0,
            .(fem_share = mean(is_female, na.rm = TRUE), n = .N), by = codigo_carrera_seleccion]
rm(demre)
invisible(gc())

fs_ok <- fs[n >= 30]
cuts <- quantile(fs_ok$fem_share, c(1/3, 2/3))
fs_ok[, ptype := fifelse(fem_share >= cuts[2], "fem", fifelse(fem_share < cuts[1], "mal", "mix"))]

cat("programs typed:", nrow(fs_ok), "of", nrow(fs),
    "(", fmt(sum(fs_ok$n) / sum(fs$n)), "of selections );",
    "tercile cutoffs:", fmt(cuts[1]), "/", fmt(cuts[2]), "\n")
cat("female share by type:\n")
print(fs_ok[, .(programs = .N, mean_fem_share = mean(fem_share)), by = ptype][order(ptype)])

type_of <- function(code) {
  t <- fs_ok$ptype[match(code, fs_ok$codigo_carrera_seleccion)]
  fifelse(is.na(t), "mix", t)   # untyped small programs count as mixed
}
df[, asg_type := fifelse(assigned == 1 & codigo_carrera_seleccion > 0,
                         type_of(codigo_carrera_seleccion), NA_character_)]
df[, enr_type := fifelse(enrolled == 1 & !is.na(codigo_carrera_matricula),
                         type_of(codigo_carrera_matricula), NA_character_)]
df[, `:=`(
  asg_fem = fifelse(is.na(assigned), NA_real_, as.numeric(assigned == 1 & asg_type == "fem" & !is.na(asg_type))),
  asg_mix = fifelse(is.na(assigned), NA_real_, as.numeric(assigned == 1 & asg_type == "mix" & !is.na(asg_type))),
  asg_mal = fifelse(is.na(assigned), NA_real_, as.numeric(assigned == 1 & asg_type == "mal" & !is.na(asg_type))),
  enr_fem = fifelse(is.na(enrolled), NA_real_, as.numeric(enrolled == 1 & enr_type == "fem" & !is.na(enr_type))),
  enr_mix = fifelse(is.na(enrolled), NA_real_, as.numeric(enrolled == 1 & enr_type == "mix" & !is.na(enr_type))),
  enr_mal = fifelse(is.na(enrolled), NA_real_, as.numeric(enrolled == 1 & enr_type == "mal" & !is.na(enr_type)))
)]
cat("\npartition check (assigned==1): ")
print(df[assigned == 1, table(asg_type, useNA = "ifany")])
```

## Headline sample

```{r echo=FALSE, warning=FALSE, message=FALSE}
part_cols <- c(names(outcomes_asg), names(outcomes_enr))
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c("assigned", "enrolled", part_cols, "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]

female <- students[GEN_ALU == 2]
male   <- students[GEN_ALU == 1]
dr_f <- class_panel(female, c("assigned", "enrolled", part_cols))
dr_m <- class_panel(male,   c("assigned", "enrolled", part_cols))

common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
run_gender <- function(dat, ycols) {
  common_args$data <- dat
  res <- map(ycols, function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), common_args))
    suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
  })
  names(res) <- ycols
  res
}

means_pre <- function(dat, ycols) {
  dat[year_region_sae > 0 & year < year_region_sae,
      sapply(.SD, mean, na.rm = TRUE), .SDcols = ycols]
}
```

## Assignment by destination-program type

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
models_f_asg <- run_gender(dr_f, names(outcomes_asg))
models_m_asg <- run_gender(dr_m, names(outcomes_asg))
names(models_f_asg) <- names(models_m_asg) <- outcomes_asg

cat(did_table_pair(models_f_asg, models_m_asg, "Female", "Male",
                   means_pre(dr_f, names(outcomes_asg)), means_pre(dr_m, names(outcomes_asg)),
                   nrow(dr_f), nrow(dr_m),
                   "Assignment by Destination-Program Type and Gender",
                   "tab: gender assigned by program type", "tables/did_by_gender_asg_ptype_AM.tex"))
```

## Enrollment by destination-program type

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
models_f_enr <- run_gender(dr_f, names(outcomes_enr))
models_m_enr <- run_gender(dr_m, names(outcomes_enr))
names(models_f_enr) <- names(models_m_enr) <- outcomes_enr

cat(did_table_pair(models_f_enr, models_m_enr, "Female", "Male",
                   means_pre(dr_f, names(outcomes_enr)), means_pre(dr_m, names(outcomes_enr)),
                   nrow(dr_f), nrow(dr_m),
                   "Enrollment by Destination-Program Type and Gender",
                   "tab: gender enrolled by program type", "tables/did_by_gender_enr_ptype_AM.tex"))
```

## Decomposition

```{r echo=FALSE, warning=FALSE, message=FALSE}
decomp <- function(res_f, res_m, labs, total_lbl) {
  af <- map_dbl(res_f, "overall.att"); am <- map_dbl(res_m, "overall.att")
  cat(sprintf("%-14s %10s %10s %10s\n", total_lbl, "female", "male", "gap F-M"))
  for (i in seq_along(labs)) cat(sprintf("%-14s %10.3f %10.3f %10.3f\n", labs[i], af[i], am[i], af[i] - am[i]))
  cat(sprintf("%-14s %10.3f %10.3f %10.3f\n", "sum", sum(af), sum(am), sum(af) - sum(am)))
  cat(sprintf("share of the F-M gap from each type: %s\n\n",
              paste(sprintf("%s %.0f%%", labs, 100 * (af - am) / (sum(af) - sum(am))), collapse = ", ")))
}
cat("-- Assigned --\n");  decomp(models_f_asg, models_m_asg, outcomes_asg, "type")
cat("-- Enrolled --\n");  decomp(models_f_enr, models_m_enr, outcomes_enr, "type")
cat("Component ATTs are estimated on the same classes, so they sum to the",
    "baseline gender ATTs up to estimation error (Assigned F 0.075 / M 0.030,",
    "Enrolled F 0.064 / M 0.024 in the draft table).\n")
```
