---
title: "Effect of SAE on SUA - Gender Gap Gate (Issue #14)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Minimum gate agreed in issue #14 after the adversarial review. Every gender
comparison is reported as Female, Male, and the within-class gap (female mean
minus male mean per school x cohort), with the gap estimated through the same
Callaway--Sant'Anna machinery, so the F$-$M column carries joint
region-clustered inference. (1) Pipeline and take-up gaps with event studies.
(2) Application validity as unconditional counts over all class members (valid
applications, invalid applications, any-invalid indicator), replacing the
applicant-conditional shares. (3) Entrant-ITT rerun of the pipeline: all
9th-grade entrants, outcomes zero-filled, no graduation filters. (4) The
pre-specified direct sorting test: probability that an entrant's 9th-grade
school is in the academic (HC) track, on a commune-of-residence x cohort
panel; every region is eventually treated, so this design has no never-treated
arm and leans on timing variation only. Observations rows count classes with a
usable outcome, column by column. Estimator, seed and sample rules as in
`3_effect_sae_am.Rmd`; `did` 2.5.1.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

outcomes_pipe <- c(scored_lm = "P(Take Exam)", applied = "Applied",
                   assigned = "Assigned", enrolled = "Enrolled")
outcomes_val  <- c(num_valid_reg = "Valid Apps", num_invalid_reg = "Invalid Apps",
                   any_invalid_reg = "Any Invalid")
outcomes_ent  <- c(applied = "Applied", assigned = "Assigned", enrolled = "Enrolled")
outcomes_hc   <- c(hc_att = "P(HC Track)")

# helpers: table formatting (fmt/star/row_tex verbatim from 3_effect_sae_am.Rmd)
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

class_panel <- function(students, cols) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

# female-minus-male panel: classes with both genders present; the control is
# total class size
gap_panel <- function(fem, mal, cols) {
  a <- class_panel(fem, cols)
  b <- class_panel(mal, cols)
  setnames(a, c(cols, "num_students"), c(paste0(cols, "_f"), "nf"))
  setnames(b, c(cols, "num_students"), c(paste0(cols, "_m"), "nm"))
  g <- merge(a, b[, c("rbd", "AGNO_9th", "nm", paste0(cols, "_m")), with = FALSE],
             by = c("rbd", "AGNO_9th"))
  for (cc in cols) set(g, j = cc, value = g[[paste0(cc, "_f")]] - g[[paste0(cc, "_m")]])
  g[, num_students := nf + nm]
  g
}

# helper: three columns per outcome (female / male / within-class gap), with
# column-specific usable-class counts
did_table_trio <- function(res_f, res_m, res_g, means_f, means_m, means_g,
                           n_f, n_m, n_g, caption, label, file) {
  ext <- function(res, fld) map_dbl(res, function(r) r$simple[[fld]])
  att <- list(ext(res_f, "overall.att"), ext(res_m, "overall.att"), ext(res_g, "overall.att"))
  se  <- list(ext(res_f, "overall.se"),  ext(res_m, "overall.se"),  ext(res_g, "overall.se"))
  p   <- map2(att, se, ~2 * pnorm(-abs(.x / .y)))
  k <- length(res_f)
  i3 <- function(a, b, c) as.vector(rbind(a, b, c))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.8}{\\begin{tabular}{l", strrep("c", 3 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{3}{c}{", names(res_f), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 3 * seq_len(k) - 1, "-", 3 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c("Female", "Male", "F$-$M"), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(3 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", i3(paste0(fmt(att[[1]]), star(p[[1]])),
                          paste0(fmt(att[[2]]), star(p[[2]])),
                          paste0(fmt(att[[3]]), star(p[[3]])))),
    row_tex("", i3(paste0("(", fmt(se[[1]]), ")"), paste0("(", fmt(se[[2]]), ")"),
                   paste0("(", fmt(se[[3]]), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", i3(fmt(means_f), fmt(means_m), fmt(means_g))),
    row_tex("Observations", i3(format(n_f, big.mark = ","), format(n_m, big.mark = ","),
                               format(n_g, big.mark = ","))),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Identical to 3_effect_sae_am.Rmd, keeping application counts, communes and track.
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU, COD_COM_ALU,
             scored_lm, applied, assigned, enrolled,
             num_apps_reg, num_valid_apps_reg)]
df[, num_valid_reg   := coalesce(num_valid_apps_reg, 0)]
df[, num_invalid_reg := pmax(coalesce(num_apps_reg, 0) - num_valid_reg, 0)]
df[, any_invalid_reg := as.integer(num_invalid_reg > 0)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)

common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap; keep the group-time fit so both the simple ATT
# and the event study come from one estimation
run_both <- function(dat, ycols, id = "rbd") {
  ca <- common_args
  ca$data <- dat
  ca$idname <- id
  res <- map(ycols, function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), ca))
    list(simple = suppressWarnings(aggte(m, type = "simple", na.rm = TRUE)),
         dyn    = suppressWarnings(aggte(m, type = "dynamic", min_e = -6, max_e = 5, na.rm = TRUE)))
  })
  names(res) <- ycols
  res
}

means_pre <- function(dat, ycols) {
  dat[year_region_sae > 0 & year < year_region_sae,
      sapply(.SD, mean, na.rm = TRUE), .SDcols = ycols]
}
n_usable <- function(dat, ycols) sapply(ycols, function(y) sum(!is.na(dat[[y]])))

print_es <- function(res, lbl) {
  d <- res$dyn
  cat("\nEvent study —", lbl, "\n")
  for (i in seq_along(d$egt)) cat(sprintf("  e=%+d  %7.3f (%.3f)\n", d$egt[i], d$att.egt[i], d$se.egt[i]))
}
```

## Headline sample

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
gate_cols <- c(names(outcomes_pipe), names(outcomes_val))
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c(gate_cols, "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]

female <- students[GEN_ALU == 2]
male   <- students[GEN_ALU == 1]
dr_f <- class_panel(female, gate_cols)
dr_m <- class_panel(male,   gate_cols)
dr_g <- gap_panel(female, male, gate_cols)
```

## 1. Pipeline and take-up, with joint F$-$M inference

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
res_f_pipe <- run_both(dr_f, names(outcomes_pipe))
res_m_pipe <- run_both(dr_m, names(outcomes_pipe))
res_g_pipe <- run_both(dr_g, names(outcomes_pipe))
names(res_f_pipe) <- names(res_m_pipe) <- names(res_g_pipe) <- outcomes_pipe

cat(did_table_trio(res_f_pipe, res_m_pipe, res_g_pipe,
                   means_pre(dr_f, names(outcomes_pipe)), means_pre(dr_m, names(outcomes_pipe)),
                   means_pre(dr_g, names(outcomes_pipe)),
                   n_usable(dr_f, names(outcomes_pipe)), n_usable(dr_m, names(outcomes_pipe)),
                   n_usable(dr_g, names(outcomes_pipe)),
                   "Pipeline Outcomes by Gender with Within-Class Gap",
                   "tab: gap pipeline", "tables/did_gap_pipeline_AM.tex"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
for (y in outcomes_pipe[c("applied", "assigned", "enrolled")])
  print_es(res_g_pipe[[y]], paste("F-M gap,", y))
```

## 2. Application validity, unconditional

Counts over all class members (zero for non-applicants): valid applications,
invalid applications, and an indicator for listing at least one invalid
program. These replace the applicant-conditional validity shares, so applier
composition cannot drive them.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
res_f_val <- run_both(dr_f, names(outcomes_val))
res_m_val <- run_both(dr_m, names(outcomes_val))
res_g_val <- run_both(dr_g, names(outcomes_val))
names(res_f_val) <- names(res_m_val) <- names(res_g_val) <- outcomes_val

cat(did_table_trio(res_f_val, res_m_val, res_g_val,
                   means_pre(dr_f, names(outcomes_val)), means_pre(dr_m, names(outcomes_val)),
                   means_pre(dr_g, names(outcomes_val)),
                   n_usable(dr_f, names(outcomes_val)), n_usable(dr_m, names(outcomes_val)),
                   n_usable(dr_g, names(outcomes_val)),
                   "Application Validity by Gender with Within-Class Gap (Unconditional)",
                   "tab: gap validity", "tables/did_gap_validity_AM.tex"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
print_es(res_g_val[["Any Invalid"]], "F-M gap, any invalid application")
```

## 3. Entrant ITT: pipeline without graduation filters

All 9th-grade entrants keyed to their entry school and cohort; outcomes are
zero unless the student later registers with the admission system, so no
post-treatment filter selects the sample.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
vals_e <- df[MRUN %in% anchor$MRUN,
             lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
             by = MRUN, .SDcols = c(names(outcomes_ent), "GEN_ALU")]
entrants <- merge(anchor, vals_e, by = "MRUN")
entrants <- entrants[AGNO_9th >= start_year & AGNO_9th <= end_year]
fem_e <- entrants[GEN_ALU == 2]
mal_e <- entrants[GEN_ALU == 1]
dr_f_e <- class_panel(fem_e, names(outcomes_ent))
dr_m_e <- class_panel(mal_e, names(outcomes_ent))
dr_g_e <- gap_panel(fem_e, mal_e, names(outcomes_ent))

res_f_ent <- run_both(dr_f_e, names(outcomes_ent))
res_m_ent <- run_both(dr_m_e, names(outcomes_ent))
res_g_ent <- run_both(dr_g_e, names(outcomes_ent))
names(res_f_ent) <- names(res_m_ent) <- names(res_g_ent) <- outcomes_ent

cat(did_table_trio(res_f_ent, res_m_ent, res_g_ent,
                   means_pre(dr_f_e, names(outcomes_ent)), means_pre(dr_m_e, names(outcomes_ent)),
                   means_pre(dr_g_e, names(outcomes_ent)),
                   n_usable(dr_f_e, names(outcomes_ent)), n_usable(dr_m_e, names(outcomes_ent)),
                   n_usable(dr_g_e, names(outcomes_ent)),
                   "Entrant-ITT Pipeline by Gender with Within-Class Gap",
                   "tab: gap entrant itt", "tables/did_gap_entrants_AM.tex"))
```

## 4. Direct sorting test: P(academic track), commune panel

Share of a commune's 9th-grade entrants whose school is in the academic (HC)
track, from each entrant's own 9th-grade record; commune of residence at 9th
grade, region from the commune code. Every region joins SAE, so there is no
never-treated arm: identification uses not-yet-treated timing variation only,
and effects for the last-treated regions are not identified.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
ent9 <- df[MRUN %in% keep & COD_ENSE %in% media & COD_GRADO == 1 & !is.na(COD_COM_ALU),
           .(MRUN, AGNO, com = as.numeric(COD_COM_ALU), hc_att = as.integer(COD_ENSE == 310))]
setorder(ent9, MRUN, AGNO)
ent9 <- unique(ent9, by = "MRUN")
ent9[, AGNO_9th := AGNO]
ent9 <- merge(ent9, vals_e[, .(MRUN, GEN_ALU)], by = "MRUN", all.x = TRUE)
ent9 <- ent9[AGNO_9th >= start_year & AGNO_9th <= end_year]

com_panel <- function(st, cols) {
  d <- st[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
          by = .(com, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d[, region := as.integer(floor(com / 1000))]
  d <- merge(d, reg9, by = "region", all.x = TRUE)
  # communes whose region has no SAE timing (e.g., code remaps) are dropped
  # rather than miscoded as never-treated
  d <- d[!is.na(g9r)]
  d[, year_region_sae := as.numeric(g9r)]
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}
gap_com <- function(fem, mal, cols) {
  a <- com_panel(fem, cols); b <- com_panel(mal, cols)
  setnames(a, c(cols, "num_students"), c(paste0(cols, "_f"), "nf"))
  setnames(b, c(cols, "num_students"), c(paste0(cols, "_m"), "nm"))
  g <- merge(a, b[, c("com", "AGNO_9th", "nm", paste0(cols, "_m")), with = FALSE],
             by = c("com", "AGNO_9th"))
  for (cc in cols) set(g, j = cc, value = g[[paste0(cc, "_f")]] - g[[paste0(cc, "_m")]])
  g[, num_students := nf + nm]
  g
}

dr_f_c <- com_panel(ent9[GEN_ALU == 2], "hc_att")
dr_m_c <- com_panel(ent9[GEN_ALU == 1], "hc_att")
dr_g_c <- gap_com(ent9[GEN_ALU == 2], ent9[GEN_ALU == 1], "hc_att")

res_f_hc <- run_both(dr_f_c, "hc_att", id = "com")
res_m_hc <- run_both(dr_m_c, "hc_att", id = "com")
res_g_hc <- run_both(dr_g_c, "hc_att", id = "com")
names(res_f_hc) <- names(res_m_hc) <- names(res_g_hc) <- outcomes_hc

cat(did_table_trio(res_f_hc, res_m_hc, res_g_hc,
                   means_pre(dr_f_c, "hc_att"), means_pre(dr_m_c, "hc_att"), means_pre(dr_g_c, "hc_att"),
                   n_usable(dr_f_c, "hc_att"), n_usable(dr_m_c, "hc_att"), n_usable(dr_g_c, "hc_att"),
                   "P(Academic Track) by Gender with Within-Commune Gap",
                   "tab: gap hc commune", "tables/did_gap_hc_commune_AM.tex"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
print_es(res_g_hc[["P(HC Track)"]], "F-M gap, P(HC track), commune panel")
```

## Summary

```{r echo=FALSE, warning=FALSE, message=FALSE}
gline <- function(res, y) sprintf("%-14s gap %7.3f (%.3f)", y, res[[y]]$simple$overall.att, res[[y]]$simple$overall.se)
cat("-- Within-class F-M gap ATTs --\n")
for (y in outcomes_pipe) cat(gline(res_g_pipe, y), "\n")
for (y in outcomes_val)  cat(gline(res_g_val, y), "\n")
cat("-- Entrant ITT gaps --\n")
for (y in outcomes_ent)  cat(gline(res_g_ent, y), "\n")
cat("-- Commune P(HC) gap --\n")
cat(gline(res_g_hc, outcomes_hc[[1]]), "\n")
cat("\nThis gate adds 33 estimations (11 outcomes x female/male/gap;",
    "event studies reuse the same fits). Cumulative for issue #14: 5",
    "pre-specified channels, 81 estimations. All under did 2.5.1;",
    "point estimates replicate the draft tables, SEs are 2.5.1-specific.\n")
```
