---
title: "Effect of SAE on SUA - Outcomes Upgrade (Issue #15)"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["booktabs"]
  html_document: default
---

Outcomes upgrade of issue #15, in three parts. (A1) Net-enrollment margin:
enrollment in any higher-education institution the year after on-time
graduation, from the SIES enrollment records (`panel_enrollment_2007_2024.rds`,
all institutions 2007--2024), split into the substitution margin (enrolled
outside the centralized system) and destination type (CRUCH universities,
private universities, professional institutes, technical centers), with
institutional accreditation as a quality proxy; overall and by gender. (A2)
Unconditional class ITT: the headline and gender tables re-estimated on all
9th-grade entrants keyed to their entry class, dropping the single-school,
graduation, and on-time filters (the planned robustness check of Appendix 7.4),
plus the net-enrollment margin on the same entrant classes. (A5) LM
decomposition: the headline LM outcome split into the taking margin
(P(take exam)) and the score conditional on taking, with the ranking-score
percentile as the GPA-based sanity outcome (issues #8/#10). Sample, estimator,
seed and table helpers are identical to `3_effect_sae_am.Rmd`; the headline
table is replicated first as a guard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(dplyr)
library(magrittr)
library(purrr)
library(data.table)
library(did)

start_year <- 2012
end_year <- 2021
# last entry cohort whose first admission year (entry + 4) falls inside the
# SIES enrollment records (2007-2024); one year earlier for the two-year window
hei_end <- 2020
hei_end2 <- 2019
media <- c(310, 410, 510, 610, 710, 810)
clustering_var <- "region"

# shared Dropbox data folder (resolves to each user's own Dropbox)
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
data_path <- file.path(home, "Dropbox", "Effect SAE on SUA", "data")
dir.create("tables", showWarnings = FALSE)

# headline outcomes (replication guard) -----------------------------------
outcomes <- c(
  gpa_secondary  = "GPA",
  scored_lm      = "P(Take Exam)",
  prom_lm_actual = "Average LM",
  applied        = "Applied",
  assigned       = "Assigned",
  enrolled       = "Enrolled"
)
# A5: LM decomposition on the headline sample
outcomes_lm <- c(
  scored_lm        = "P(Take Exam)",
  prom_lm_actual   = "Average LM",
  lm_pct_cond      = "LM Pct. (takers)",
  ranking_pct      = "Ranking Pct.",
  ranking_pct_cond = "Ranking Pct. (takers)"
)
# A1: net-enrollment margin (headline sample, entry cohorts <= hei_end)
outcomes_hei <- c(
  enrolled       = "Enrolled SUA",
  enrolled_any   = "Any HEI",
  enrolled_out   = "Outside SUA",
  enrolled_cruch = "CRUCH",
  enrolled_upriv = "Private Univ.",
  enrolled_ip    = "IP",
  enrolled_cft   = "CFT",
  enrolled_acc   = "Accredited"
)
outcomes_hei_g <- c(
  enrolled     = "Enrolled SUA",
  enrolled_any = "Any HEI",
  enrolled_out = "Outside SUA",
  enrolled_tec = "IP or CFT",
  enrolled_acc = "Accredited"
)
# A2: unconditional entrant classes
outcomes_ent <- c(
  on_time_grad   = "Grad. On Time",
  gpa_secondary  = "GPA",
  scored_lm      = "P(Take Exam)",
  prom_lm_actual = "Average LM",
  applied        = "Applied",
  assigned       = "Assigned",
  enrolled       = "Enrolled"
)
outcomes_ent_hei <- c(
  enrolled     = "Enrolled SUA",
  enrolled_any = "Any HEI",
  enrolled_out = "Outside SUA",
  enrolled_tec = "IP or CFT"
)
hei_cols <- c("enrolled_any", "enrolled_out", "enrolled_cruch", "enrolled_upriv",
              "enrolled_ip", "enrolled_cft", "enrolled_tec", "enrolled_acc",
              "enrolled_any2", "enrolled_out2")

# helpers: table formatting (verbatim from 3_effect_sae_am.Rmd)
fmt  <- function(x) sub("^-(0\\.000)$", "\\1", sprintf("%.3f", x))
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# helper: collapse students to school x cohort classes (class means over the
# students passed in; num_students = their count)
class_panel <- function(students, cols = names(outcomes)) {
  d <- students[, c(.(num_students = .N), lapply(.SD, mean, na.rm = TRUE)),
                by = .(rbd, AGNO_9th), .SDcols = cols]
  d[, year := AGNO_9th - 1]
  d <- merge(d, sch[, .(rbd, year_region_sae, region)], by = "rbd", all.x = TRUE)
  d[year_region_sae > max(year), year_region_sae := 0]
  d
}

# helper: simple aggregated ATT table, coefficients with region-clustered
# bootstrap SEs, pre-treatment means over treated classes, one column per outcome
did_table <- function(res, means_pre, nobs, caption, label, file) {
  att <- map_dbl(res, "overall.att")
  se  <- map_dbl(res, "overall.se")
  p   <- 2 * pnorm(-abs(att / se))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", length(res)), "}\n      \\toprule\n",
    "                   & ", paste(names(res), collapse = " & "), "\\\\\n",
    "                   & ", paste0("(", seq_along(res), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", paste0(fmt(att), star(p))),
    row_tex("", paste0("(", fmt(se), ")")),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", fmt(means_pre)),
    row_tex("Observations", rep(format(nobs, big.mark = ","), length(res))),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}

# helper: same table with two columns per outcome (e.g. female / male)
did_table_pair <- function(res_a, res_b, lab_a, lab_b, means_a, means_b, n_a, n_b,
                           caption, label, file) {
  att_a <- map_dbl(res_a, "overall.att"); se_a <- map_dbl(res_a, "overall.se")
  att_b <- map_dbl(res_b, "overall.att"); se_b <- map_dbl(res_b, "overall.se")
  p_a <- 2 * pnorm(-abs(att_a / se_a))
  p_b <- 2 * pnorm(-abs(att_b / se_b))
  k <- length(res_a)
  interleave <- function(a, b) as.vector(rbind(a, b))
  tab <- paste0(
    "\\begin{table}[htbp]\n   \\caption{", caption, "}\\label{", label, "}\n",
    "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{l", strrep("c", 2 * k), "}\n      \\toprule\n",
    paste0("      & \\multicolumn{2}{c}{", names(res_a), "}", collapse = "\n"), " \\\\\n      ",
    paste0("\\cmidrule(lr){", 2 * seq_len(k), "-", 2 * seq_len(k) + 1, "}", collapse = ""), "\n",
    "                   & ", paste(rep(c(lab_a, lab_b), k), collapse = " & "), " \\\\\n",
    "                   & ", paste0("(", seq_len(2 * k), ")", collapse = " & "), "\\\\\n      \\midrule\n",
    row_tex("Treated", interleave(paste0(fmt(att_a), star(p_a)), paste0(fmt(att_b), star(p_b)))),
    row_tex("", interleave(paste0("(", fmt(se_a), ")"), paste0("(", fmt(se_b), ")"))),
    "      \\midrule\n",
    row_tex("Mean (pre-treatment)", interleave(fmt(means_a), fmt(means_b))),
    row_tex("Observations", rep(c(format(n_a, big.mark = ","), format(n_b, big.mark = ",")), k)),
    "      \\bottomrule\n   \\end{tabular}}}\n\\end{table}\n")
  writeLines(tab, file)
  tab
}
```

## Data and treatment

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Student panel. Treatment: a school joins SAE the first year 9th grade in its
# region is assigned through the system; never-treated = private (DEPE 4).
# Identical to 3_effect_sae_am.Rmd, keeping the score-percentile columns.
df <- readRDS(file.path(data_path, "panel_9-12_2010_2021.rds"))
setDT(df)
df <- df[, .(MRUN, AGNO, RBD, COD_GRADO, COD_ENSE, year_rbd_first_sae,
             graduated_secondary, COD_REG_RBD, COD_DEPE, GEN_ALU,
             gpa_secondary, scored_lm, prom_lm_actual, applied, assigned, enrolled,
             lm_pct, ranking_pct)]

# conditional versions: percentile among students with a computed score
# (same convention as 5_gender_mechanisms.Rmd)
df[, lm_pct_cond      := fifelse(lm_pct > 0, lm_pct, NA_real_)]
df[, ranking_pct_cond := fifelse(ranking_pct > 0, ranking_pct, NA_real_)]

sch <- unique(df[, .(RBD, year_rbd_first_sae, COD_REG_RBD, COD_DEPE)])[, .(
  yfs    = { v <- year_rbd_first_sae[!is.na(year_rbd_first_sae)]; if (length(v)) as.numeric(v[1]) else 0 },
  region = { v <- COD_REG_RBD[!is.na(COD_REG_RBD)]; if (length(v)) as.integer(v[1]) else NA_integer_ },
  depe   = { v <- COD_DEPE[!is.na(COD_DEPE)]; if (length(v)) as.integer(names(which.max(table(v)))) else NA_integer_ }
), by = RBD]
setnames(sch, "RBD", "rbd")

sae <- readRDS(file.path(data_path, "sae_data_2016_2023.rds"))
setDT(sae)
nine <- sae[COD_ENSE %in% media & COD_GRADO == 1, .(g9 = as.numeric(min(AGNO))), by = RBD]
setnames(nine, "RBD", "rbd")
sch <- merge(sch, nine, by = "rbd", all.x = TRUE)
reg9 <- sch[!is.na(g9), .(g9r = min(g9)), by = region][!is.na(region)]
sch <- merge(sch, reg9, by = "region", all.x = TRUE)
sch[, elig := fifelse(!is.na(depe), as.integer(depe != 4), as.integer(yfs > 0 | !is.na(g9)))]
sch[, year_region_sae := fifelse(elig == 1 & !is.na(g9r), as.numeric(g9r), 0)]
rm(sae, nine, reg9)

# regular-track students; high school records for grades 9-12
keep <- df[, all(COD_ENSE == 110 | COD_ENSE %in% media), by = MRUN][V1 == TRUE, MRUN]
hs <- df[MRUN %in% keep & AGNO >= 2010 & AGNO <= 2024 &
           COD_ENSE %in% media & COD_GRADO %in% 1:4, .(MRUN, AGNO, RBD, COD_GRADO)]
setorder(hs, MRUN, AGNO, COD_GRADO)
```

## Higher-education enrollment records

SIES enrollment records, all institutions and years 2007--2024, one record per
student x program x year. A student counts as enrolled in a year if they hold
any undergraduate (pregrado) record in it; institution type comes from the
SIES classification (CRUCH universities, private universities, professional
institutes, technical centers), and the quality proxy is enrollment in an
institution accredited in that year. `hei_demre` marks enrollment in a program
carrying a DEMRE code (a program admitting through the centralized system),
printed below as a cross-check of the DEMRE-based enrollment flag.

```{r echo=FALSE, warning=FALSE, message=FALSE}
sies <- readRDS(file.path(data_path, "panel_enrollment_2007_2024.rds"))
setDT(sies)
sies <- sies[nivel_global == "Pregrado" & mrun %in% unique(df$MRUN),
             .(mrun, YEAR, tipo_inst_2, acreditada_inst, codigo_demre)]
invisible(gc())

hei <- sies[, .(
  hei_any   = 1L,
  hei_cruch = as.integer(any(tipo_inst_2 == "Universidades CRUCH")),
  hei_upriv = as.integer(any(tipo_inst_2 %chin% c("Universidades Privadas",
                                                  "Universidades (* Carrera en Convenio)"))),
  hei_ip    = as.integer(any(tipo_inst_2 == "Institutos Profesionales")),
  hei_cft   = as.integer(any(tipo_inst_2 == "Centros de Formación Técnica")),
  hei_acc   = as.integer(any(acreditada_inst == "ACREDITADA")),
  hei_demre = as.integer(any(!is.na(codigo_demre)))
), by = .(mrun, YEAR)]
rm(sies)
invisible(gc())

cat("panel students with any pregrado enrollment, by year:\n")
print(hei[, .(students = .N, cruch = sum(hei_cruch), upriv = sum(hei_upriv),
              ip = sum(hei_ip), cft = sum(hei_cft), demre_prog = sum(hei_demre)),
          by = YEAR][order(YEAR)])

# merge enrollment at entry + 4 (first admission year after on-time
# graduation) and entry + 5 onto a student table keyed by MRUN and AGNO_9th
hei2 <- copy(hei)
setnames(hei2, setdiff(names(hei2), c("mrun", "YEAR")),
         paste0(setdiff(names(hei2), c("mrun", "YEAR")), "_2"))
add_hei <- function(d) {
  d[, adm_year := AGNO_9th + 4]
  d[, adm_year2 := AGNO_9th + 5]
  d <- merge(d, hei, by.x = c("MRUN", "adm_year"), by.y = c("mrun", "YEAR"), all.x = TRUE)
  d <- merge(d, hei2, by.x = c("MRUN", "adm_year2"), by.y = c("mrun", "YEAR"), all.x = TRUE)
  d[, enrolled_any   := fcoalesce(hei_any, 0L)]
  d[, enrolled_out   := as.integer(enrolled_any == 1L & enrolled == 0)]
  d[, enrolled_cruch := fcoalesce(hei_cruch, 0L)]
  d[, enrolled_upriv := fcoalesce(hei_upriv, 0L)]
  d[, enrolled_ip    := fcoalesce(hei_ip, 0L)]
  d[, enrolled_cft   := fcoalesce(hei_cft, 0L)]
  d[, enrolled_tec   := pmax(enrolled_ip, enrolled_cft)]
  d[, enrolled_acc   := fcoalesce(hei_acc, 0L)]
  d[, enrolled_any2  := pmax(enrolled_any, fcoalesce(hei_any_2, 0L))]
  d[, enrolled_out2  := as.integer(enrolled_any2 == 1L & enrolled == 0)]
  d
}
```

## Headline sample and replication guard

Same sample as the headline table: graduated on time, all four grades in the
same school, cohorts entering 9th grade `r start_year`--`r end_year`. The six
headline outcomes replicate `3_effect_sae_am.Rmd` (a check that this file
reproduces it exactly) before any new outcome is estimated.

```{r echo=FALSE, warning=FALSE, message=FALSE}
anchor <- hs[COD_GRADO == 1, .(rbd = RBD[1], AGNO_9th = AGNO[1]), by = MRUN]
hs1 <- merge(hs, anchor[, .(MRUN, rbd)], by = "MRUN")
per <- hs1[, .(ng = uniqueN(COD_GRADO[RBD == rbd]), nsch = uniqueN(RBD)), by = MRUN]
per <- merge(per, anchor, by = "MRUN")
stayers <- per[ng == 4 & nsch == 1, .(MRUN, rbd, AGNO_9th)]
setkey(stayers, MRUN)
grad <- hs1[MRUN %in% stayers$MRUN, .(grad_year = max(AGNO)), by = MRUN]
vals <- df[MRUN %in% stayers$MRUN,
           lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
           by = MRUN, .SDcols = c(names(outcomes), "lm_pct_cond", "ranking_pct",
                                  "ranking_pct_cond", "graduated_secondary", "GEN_ALU")]
students <- merge(stayers, vals, by = "MRUN")
students <- merge(students, grad, by = "MRUN")
students <- students[graduated_secondary == 1 & grad_year == AGNO_9th + 3 &
                       AGNO_9th >= start_year & AGNO_9th <= end_year]
students <- add_hei(students)

cat("students found in any-HEI enrollment at entry + 4, by entry cohort\n",
    "(cohorts through", hei_end, "are covered by the SIES records):\n")
print(students[, .(n = .N, enrolled_sua = mean(enrolled), any_hei = mean(enrolled_any),
                   outside = mean(enrolled_out)), by = AGNO_9th][order(AGNO_9th)])
cat("\ncross-check: DEMRE enrollment flag vs. SIES enrollment at entry + 4\n",
    "(cohorts through", hei_end, "):\n")
print(students[AGNO_9th <= hei_end, table(enrolled, enrolled_any)])
cat("\nDEMRE-enrolled students found in a DEMRE-coded SIES program at entry + 4:",
    fmt(students[AGNO_9th <= hei_end & enrolled == 1, mean(fcoalesce(hei_demre, 0L))]), "\n")

dr <- class_panel(students, unique(c(names(outcomes), names(outcomes_lm))))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
common_args <- list(
  tname = "year",
  idname = "rbd",
  gname = "year_region_sae",
  xformla = ~num_students,
  data = dr,
  control_group = "notyettreated",
  anticipation = 2,
  base_period = "universal",
  est_method = "dr",
  allow_unbalanced_panel = TRUE,
  clustervars = clustering_var,
  panel = TRUE,
  bstrap = TRUE,
  biters = 10000,
  cband = TRUE
)

# same seed for each bootstrap so results do not depend on the order the
# models are run
run_set <- function(dat, ycols) {
  common_args$data <- dat
  res <- map(ycols, function(y) {
    set.seed(20260610)
    m <- do.call(att_gt, c(list(yname = y), common_args))
    suppressWarnings(aggte(m, type = "simple", na.rm = TRUE))
  })
  names(res) <- ycols
  res
}

means_pre <- function(dat, ycols) {
  dat[year_region_sae > 0 & year < year_region_sae,
      sapply(.SD, mean, na.rm = TRUE), .SDcols = ycols]
}

models_base <- run_set(dr, names(outcomes))
names(models_base) <- outcomes

cat(did_table(models_base, means_pre(dr, names(outcomes)), nrow(dr),
              "Regression Results (Replication)", "tab: headline replication",
              "tables/did_headline_replication_AM.tex"))
```

## A5. LM decomposition

The headline `Average LM` outcome is the within-admission-year percentile of
the average Language and Math score, with zero assigned to students who never
take the exam, so its treatment effect mixes the taking margin with
performance conditional on taking. `P(Take Exam)` isolates the first margin;
`LM Pct. (takers)` the second. The ranking-score percentile is a deterministic
function of GPA relative to the school's recent history, so its
taker-conditional effect is a composition gauge for the taker pool: it moves
only if treatment changes who registers, not how they perform (issues #8/#10).

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
models_lm <- run_set(dr, names(outcomes_lm))
names(models_lm) <- outcomes_lm

cat(did_table(models_lm, means_pre(dr, names(outcomes_lm)), nrow(dr),
              "Decomposition of the LM Effect: Taking vs. Performance",
              "tab: lm decomposition", "tables/did_lm_decomposition_AM.tex"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# arithmetic: ATT on the unconditional percentile ~ taking margin x pre-mean
# conditional score + pre-mean taking rate x conditional margin
att <- function(m, y) m[[y]]$overall.att
p_pre <- means_pre(dr, "scored_lm"); c_pre <- means_pre(dr, "lm_pct_cond")
a_take <- att(models_lm, "P(Take Exam)"); a_cond <- att(models_lm, "LM Pct. (takers)")
cat(sprintf(paste0("LM decomposition arithmetic: ATT(unconditional) = %.4f;\n",
                   "  taking margin  ATT(P take) x pre-mean cond. score = %.4f x %.3f = %.4f\n",
                   "  performance    pre-mean P take x ATT(cond. score) = %.3f x %.4f = %.4f\n",
                   "  sum of the two margins = %.4f\n"),
            att(models_lm, "Average LM"),
            a_take, c_pre, a_take * c_pre,
            p_pre, a_cond, p_pre * a_cond,
            a_take * c_pre + p_pre * a_cond))
```

## A1. Net-enrollment margin

Same headline sample and estimator, entry cohorts `r start_year`--`r hei_end`
(the last cohort whose first admission year is covered by the SIES records).
`Enrolled SUA` is the headline enrollment outcome re-estimated on these
cohorts; `Any HEI` is enrollment in any institution in the year after on-time
graduation; `Outside SUA` is enrollment in some institution without an
enrollment through the centralized system; the destination types and the
accreditation proxy are unconditional class shares, so they add up over the
`Any HEI` margin.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
students_hei <- students[AGNO_9th <= hei_end]
dr_hei <- class_panel(students_hei, c("enrolled", hei_cols))

models_hei <- run_set(dr_hei, names(outcomes_hei))
names(models_hei) <- outcomes_hei

cat(did_table(models_hei, means_pre(dr_hei, names(outcomes_hei)), nrow(dr_hei),
              "Effect on Enrollment in Any Higher-Education Institution",
              "tab: net enrollment", "tables/did_net_enrollment_AM.tex"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# two-year window (entry + 4 or entry + 5), one cohort shorter
dr_hei2 <- class_panel(students[AGNO_9th <= hei_end2], c("enrolled_any2", "enrolled_out2"))
models_hei2 <- run_set(dr_hei2, c("enrolled_any2", "enrolled_out2"))
cat(sprintf(paste0("two-year window (cohorts through %d, %s classes):\n",
                   "  Any HEI (2y)     %7.3f (%.3f)   pre-mean %.3f\n",
                   "  Outside SUA (2y) %7.3f (%.3f)   pre-mean %.3f\n"),
            hei_end2, format(nrow(dr_hei2), big.mark = ","),
            models_hei2$enrolled_any2$overall.att, models_hei2$enrolled_any2$overall.se,
            means_pre(dr_hei2, "enrolled_any2"),
            models_hei2$enrolled_out2$overall.att, models_hei2$enrolled_out2$overall.se,
            means_pre(dr_hei2, "enrolled_out2")))
```

### By gender

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female_hei <- students_hei[GEN_ALU == 2]
male_hei   <- students_hei[GEN_ALU == 1]
dr_f_hei <- class_panel(female_hei, c("enrolled", hei_cols))
dr_m_hei <- class_panel(male_hei,   c("enrolled", hei_cols))

models_f_hei <- run_set(dr_f_hei, names(outcomes_hei_g))
models_m_hei <- run_set(dr_m_hei, names(outcomes_hei_g))
names(models_f_hei) <- names(models_m_hei) <- outcomes_hei_g

cat(did_table_pair(models_f_hei, models_m_hei, "Female", "Male",
                   means_pre(dr_f_hei, names(outcomes_hei_g)),
                   means_pre(dr_m_hei, names(outcomes_hei_g)),
                   nrow(dr_f_hei), nrow(dr_m_hei),
                   "Effect on Enrollment in Any Higher-Education Institution by Gender",
                   "tab: net enrollment by gender",
                   "tables/did_net_enrollment_by_gender_AM.tex"))
```

## A2. Unconditional entrant classes

All 9th-grade entrants, keyed to their entry class (school x entry year), with
no single-school, graduation, or on-time requirement: the class ITT of
Appendix 7.4, on the same regular-track universe as the headline sample.
Outcomes carry zeros for students who never register with DEMRE, so class
means are over the entire entering class; GPA averages each student's recorded
year grades over the secondary years they attended, so partial careers enter
its class mean. `Grad. On Time` makes the margin between the entrant and
headline samples visible in the same exhibit.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
g12 <- hs[COD_GRADO == 4, .(g12_year = min(AGNO)), by = MRUN]
vals_e <- df[MRUN %in% anchor$MRUN,
             lapply(.SD, function(x) { v <- x[!is.na(x)]; if (length(v)) as.numeric(v[1]) else NA_real_ }),
             by = MRUN, .SDcols = c(names(outcomes), "graduated_secondary", "GEN_ALU")]
entrants <- merge(anchor, vals_e, by = "MRUN", all.x = TRUE)
entrants <- merge(entrants, g12, by = "MRUN", all.x = TRUE)
entrants[, grad_ever    := as.integer(coalesce(graduated_secondary, 0) == 1)]
entrants[, on_time_grad := as.integer(grad_ever == 1 & !is.na(g12_year) & g12_year == AGNO_9th + 3)]
entrants <- entrants[AGNO_9th >= start_year & AGNO_9th <= end_year]
entrants <- add_hei(entrants)

cat(format(nrow(entrants), big.mark = ","), "entrants;",
    format(uniqueN(entrants[on_time_grad == 1, MRUN]), big.mark = ","),
    "graduate on time;", format(nrow(students), big.mark = ","),
    "pass the headline filters.\n\n")

dr_e <- class_panel(entrants, names(outcomes_ent))

models_e <- run_set(dr_e, names(outcomes_ent))
names(models_e) <- outcomes_ent

cat(did_table(models_e, means_pre(dr_e, names(outcomes_ent)), nrow(dr_e),
              "Regression Results: All 9th-Grade Entrants",
              "tab: entrants", "tables/did_entrants_AM.tex"))
```

### By gender

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female_e <- entrants[GEN_ALU == 2]
male_e   <- entrants[GEN_ALU == 1]
dr_f_e <- class_panel(female_e, names(outcomes))
dr_m_e <- class_panel(male_e,   names(outcomes))

models_f_e <- run_set(dr_f_e, names(outcomes))
models_m_e <- run_set(dr_m_e, names(outcomes))
names(models_f_e) <- names(models_m_e) <- outcomes

cat(did_table_pair(models_f_e, models_m_e, "Female", "Male",
                   means_pre(dr_f_e, names(outcomes)),
                   means_pre(dr_m_e, names(outcomes)),
                   nrow(dr_f_e), nrow(dr_m_e),
                   "Regression Results by Gender: All 9th-Grade Entrants",
                   "tab: entrants by gender", "tables/did_entrants_by_gender_AM.tex"))
```

### Net enrollment on entrant classes

The access question on the unconditional sample: does enrollment in any
higher-education institution rise for the full entering class? Entry cohorts
`r start_year`--`r hei_end`; enrollment measured in the year after on-time
graduation (entry + 4), so late graduates enrolling later are picked up by the
two-year window printed below the table.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
entrants_hei <- entrants[AGNO_9th <= hei_end]
dr_e_hei <- class_panel(entrants_hei, c("enrolled", hei_cols))

models_e_hei <- run_set(dr_e_hei, names(outcomes_ent_hei))
names(models_e_hei) <- outcomes_ent_hei

cat(did_table(models_e_hei, means_pre(dr_e_hei, names(outcomes_ent_hei)), nrow(dr_e_hei),
              "Effect on Enrollment in Any Higher-Education Institution: All 9th-Grade Entrants",
              "tab: entrants net enrollment", "tables/did_entrants_net_enrollment_AM.tex"))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
dr_e_hei2 <- class_panel(entrants[AGNO_9th <= hei_end2], c("enrolled_any2", "enrolled_out2"))
models_e_hei2 <- run_set(dr_e_hei2, c("enrolled_any2", "enrolled_out2"))
cat(sprintf(paste0("two-year window (cohorts through %d, %s classes):\n",
                   "  Any HEI (2y)     %7.3f (%.3f)   pre-mean %.3f\n",
                   "  Outside SUA (2y) %7.3f (%.3f)   pre-mean %.3f\n"),
            hei_end2, format(nrow(dr_e_hei2), big.mark = ","),
            models_e_hei2$enrolled_any2$overall.att, models_e_hei2$enrolled_any2$overall.se,
            means_pre(dr_e_hei2, "enrolled_any2"),
            models_e_hei2$enrolled_out2$overall.att, models_e_hei2$enrolled_out2$overall.se,
            means_pre(dr_e_hei2, "enrolled_out2")))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
female_e_hei <- entrants_hei[GEN_ALU == 2]
male_e_hei   <- entrants_hei[GEN_ALU == 1]
dr_f_e_hei <- class_panel(female_e_hei, c("enrolled", hei_cols))
dr_m_e_hei <- class_panel(male_e_hei,   c("enrolled", hei_cols))

models_f_e_hei <- run_set(dr_f_e_hei, names(outcomes_ent_hei))
models_m_e_hei <- run_set(dr_m_e_hei, names(outcomes_ent_hei))
names(models_f_e_hei) <- names(models_m_e_hei) <- outcomes_ent_hei

cat(did_table_pair(models_f_e_hei, models_m_e_hei, "Female", "Male",
                   means_pre(dr_f_e_hei, names(outcomes_ent_hei)),
                   means_pre(dr_m_e_hei, names(outcomes_ent_hei)),
                   nrow(dr_f_e_hei), nrow(dr_m_e_hei),
                   "Effect on Enrollment in Any HEI by Gender: All 9th-Grade Entrants",
                   "tab: entrants net enrollment by gender",
                   "tables/did_entrants_net_enrollment_by_gender_AM.tex"))
```

## Summary

```{r echo=FALSE, warning=FALSE, message=FALSE}
cat("did version:", as.character(packageVersion("did")), "|", R.version.string, "\n\n")

line <- function(res, y, lbl) cat(sprintf("%-24s %7.3f (%.3f)\n", lbl,
                                          res[[y]]$overall.att, res[[y]]$overall.se))
cat("-- Replication guard: committed headline table (did_by 3_effect_sae_am.Rmd) --\n")
draft <- c(gpa_secondary = -0.005, prom_lm_actual = -0.016, applied = 0.045,
           assigned = 0.051, enrolled = 0.041)
for (y in names(draft)) {
  cat(sprintf("%-16s this run %7.3f | draft table %7.3f | diff %8.4f\n",
              outcomes[[y]], models_base[[outcomes[[y]]]]$overall.att, draft[[y]],
              models_base[[outcomes[[y]]]]$overall.att - draft[[y]]))
}
cat("(draft table is the committed 5-column vintage; P(Take Exam) is the HEAD\n",
    "6th outcome and has no committed counterpart. Bootstrap SEs under did\n",
    "2.5.1 differ from the draft tables even though point estimates match.)\n")

cat("\n-- A5. LM decomposition (headline sample) --\n")
for (y in outcomes_lm) line(models_lm, y, y)

cat("\n-- A1. Net enrollment (headline sample, cohorts through", hei_end, ") --\n")
for (y in outcomes_hei) line(models_hei, y, y)
cat("\n   by gender (female | male):\n")
for (y in outcomes_hei_g) cat(sprintf("%-24s female %7.3f (%.3f) | male %7.3f (%.3f)\n", y,
                                      models_f_hei[[y]]$overall.att, models_f_hei[[y]]$overall.se,
                                      models_m_hei[[y]]$overall.att, models_m_hei[[y]]$overall.se))

cat("\n-- A2. Entrant classes --\n")
for (y in outcomes_ent) line(models_e, y, y)
cat("\n   by gender (female | male):\n")
for (y in outcomes) cat(sprintf("%-24s female %7.3f (%.3f) | male %7.3f (%.3f)\n", y,
                                models_f_e[[y]]$overall.att, models_f_e[[y]]$overall.se,
                                models_m_e[[y]]$overall.att, models_m_e[[y]]$overall.se))
cat("\n   net enrollment on entrant classes (cohorts through", hei_end, "):\n")
for (y in outcomes_ent_hei) line(models_e_hei, y, y)
cat("\n   by gender (female | male):\n")
for (y in outcomes_ent_hei) cat(sprintf("%-24s female %7.3f (%.3f) | male %7.3f (%.3f)\n", y,
                                        models_f_e_hei[[y]]$overall.att, models_f_e_hei[[y]]$overall.se,
                                        models_m_e_hei[[y]]$overall.att, models_m_e_hei[[y]]$overall.se))

cat("\nGuard: the replication block must match the committed headline table;\n",
    "if it does not, the sample replication is off and the new rows are not\n",
    "comparable to the draft.\n")
```
