---
title: "Piece 3 -- Building the Mechanism Panel"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

This document builds `lottery_panel_mech.rds`: the STAMPED Piece 2 panel
(asserted by MD5, its columns re-asserted byte-identical at the end -- never
modified) plus the mechanism outcome columns, one row per child per lottery.
The analysis file `1_effect_mech.Rmd` reads only this panel.

**Inputs.** All resolve through one chain: the shared data folder
(`~/Dropbox/Effect SAE on SUA/data`), the project's `code/data` mirror, the
Piece 2 folder, the Piece 1 `data/` folder, then this folder.

- `lottery_panel.rds` -- Piece 2's stamped panel of record
  (MD5 `acb5b2fff844f8692f234442273ab4c8`, asserted).
- `college_outcomes_2018_2026.rds` -- Piece 2's stamped outcome table
  (MD5 `28dd1874092a0835bbcb11839ff2e57c`, asserted): exam scores and the
  official NEM/rank scores, processes 2018--2026.
- `panel_performance_2002_2023.rds` -- the canonical graded panel: grades,
  schools, promotion status through 2023.
- `rendimiento_grades_2024_2025.rds` -- Piece 3's grades supplement (built by
  `00a_pack_rendimiento_grades.R` from the two raw public releases): grades,
  school id, and promotion status for academic years 2024--2025.
- `matricula_2011_2023.rds` -- the packed enrolment census: who sits in which
  school (the peer measure's rooms).

**The new columns.** For each child: the strength of the room they enter
(leave-one-out mean 8th-grade GPA, national z, of the census 9th-grade cohort
at the school attended in the entry year); own 9th-grade GPA (raw and national
z) and midrank percentile in the school-grade graded cohort; the on-time 4-year
GPA (grades 9-12 in proc+1..proc+4, all four required; raw and national
entry-cohort z) and the 12th-grade percentile; rank vs school history (the
4-year GPA against the pooled media-GPA of the three promoted classes preceding
the child's own at the child's 12th-grade school; promoted = ANY record of the
student-year-school key says so, decided before dedup, so row order can never
decide membership; guards: 3 class years, n >= 5, sd > 0); the college exam
score at proc+5 (NA where no exam sat -- NEVER 0), its floor-imputed version
(no-show = the process's minimum observed score, which equals the scale
minimum, in taker-z units), and DEMRE's official NEM and rank scores with
national within-process z versions; and the coverage flags. Conventions carried
from the certified pieces: 8th grade = COD_ENSE 110 grade 8; media years =
COD_ENSE in {310..910} grades 1-4; valid grades PROM_GRAL in [1,7]; national z
within year; duplicate graded student-years collapse by mean. The 8th-grade z
is recomputed and asserted equal to the stamped panel's `gpa8_z`. No stochastic
element: no seed needed.

```{r build, echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({ library(data.table); library(tools) })
setDTthreads(4); options(warn = 1)
stopifnot("R 4.5.1 required (piece standard)" = getRversion() == "4.5.1")
t0 <- proc.time()
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
ROOT <- file.path(home, "Dropbox", "Effect of Centralization")
OUTF <- "lottery_panel_mech.rds"
data_path <- function(f) {
  for (p in c(file.path(home, "Dropbox", "Effect SAE on SUA", "data", f),
              file.path(ROOT, "code", "data", f),
              file.path(ROOT, "Updated Paper Spine", "Piece 2", f),
              file.path(ROOT, "Updated Paper Spine", "Piece 1", "data", f),
              f))
    if (file.exists(p)) return(p)
  stop(sprintf("input file not found in any data location: %s", f))
}
say <- function(...) { cat(sprintf(...), "\n"); flush.console() }

MD5_PANEL <- "acb5b2fff844f8692f234442273ab4c8"
MD5_CO    <- "28dd1874092a0835bbcb11839ff2e57c"
YM <- c(310L, 410L, 510L, 610L, 710L, 810L, 910L)   # media school types (shipped list)

## Stage A -- the stamped panel
fp <- data_path("lottery_panel.rds")
stopifnot("lottery_panel.rds must match the stamped MD5" =
            unname(md5sum(fp)) == MD5_PANEL)
PT <- as.data.table(readRDS(fp))
stopifnot("row count of record" = nrow(PT) == 395950L,
          "one row per child per lottery" = nrow(PT) == uniqueN(PT, by = c("key", "MRUN")))
ORIG <- copy(PT)
say("panel loaded: %s rows | MD5 ok", format(nrow(PT), big.mark = ","))

## Stage B -- graded records (canonical panel + 2024-25 grades supplement)
PF <- as.data.table(readRDS(data_path("panel_performance_2002_2023.rds")))
G8 <- PF[COD_GRADO == 8L & COD_ENSE == 110L & AGNO %in% 2016:2021 &
           !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7, .(MRUN, AGNO, PROM_GRAL)]
G8 <- G8[, .(gpa8 = mean(PROM_GRAL)), by = .(MRUN, AGNO)]
G8[, z := (gpa8 - mean(gpa8)) / sd(gpa8), by = AGNO]
qz <- data.table(MRUN = PT$MRUN, AGNO = PT$proc)
z_new <- G8[qz, on = .(MRUN, AGNO), x.z]
stopifnot("recomputed 8th-grade z must match the stamped panel (values)" =
            max(abs(PT$gpa8_z - z_new), na.rm = TRUE) < 1e-12,
          "recomputed 8th-grade z must match the stamped panel (NA pattern)" =
            sum(is.na(PT$gpa8_z) != is.na(z_new)) == 0L)
rm(qz, z_new); say("8th-grade z recomputation matches the stamped panel")

MED <- PF[COD_ENSE %in% YM & COD_GRADO %in% 1:4 & AGNO %in% 2014:2023 &
            !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7,
          .(MRUN, AGNO, RBD, COD_GRADO, PROM_GRAL, sit = SIT_FIN_R)]
rm(PF); invisible(gc())
RS <- as.data.table(readRDS(data_path("rendimiento_grades_2024_2025.rds")))
RS <- RS[COD_ENSE %in% YM & COD_GRADO %in% 1:4 &
           !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7,
         .(MRUN, AGNO, RBD, COD_GRADO, PROM_GRAL, sit = SIT_FIN_R)]
MED <- rbind(MED, RS); rm(RS); invisible(gc())
# promoted-12th membership is decided BEFORE the keep-first dedup: a student is
# promoted if ANY record of the key says so (the certified Piece 2 convention
# for promotion status), so row order can never decide history-class membership
P12A <- unique(MED[COD_GRADO == 4L & sit == "P", .(MRUN, AGNO, RBD)])
MED <- unique(MED, by = c("MRUN", "AGNO", "RBD", "COD_GRADO"))
stopifnot("media records must cover 2014-2025" = all(2014:2025 %in% unique(MED$AGNO)))
say("media graded records 2014-2025: %s rows", format(nrow(MED), big.mark = ","))
MC <- MED[, .(v = mean(PROM_GRAL)), by = .(MRUN, AGNO, COD_GRADO)]

## Stage C -- own 9th-grade GPA, national z, school percentile
R9 <- MED[COD_GRADO == 1L]
R9[, pct := (frank(PROM_GRAL, ties.method = "average") - 0.5) / .N, by = .(RBD, AGNO)]
P9 <- R9[, .(gpa9 = mean(PROM_GRAL), pct9 = mean(pct), nsch = uniqueN(RBD)), by = .(MRUN, AGNO)]
stopifnot("P9 unique per student-year" = !anyDuplicated(P9, by = c("MRUN", "AGNO")))
P9[, gpa9_z := (gpa9 - mean(gpa9)) / sd(gpa9), by = AGNO]
say("9th-grade student-years: %s | graded at >1 school %.4f",
    format(nrow(P9), big.mark = ","), P9[, mean(nsch > 1L)])
q1 <- data.table(MRUN = PT$MRUN, AGNO = PT$proc + 1L)
PT[, gpa9   := P9[q1, on = .(MRUN, AGNO), x.gpa9]]
PT[, gpa9_z := P9[q1, on = .(MRUN, AGNO), x.gpa9_z]]
PT[, pct9   := P9[q1, on = .(MRUN, AGNO), x.pct9]]
PT[, has_g9 := as.integer(!is.na(gpa9))]
rm(R9); invisible(gc())

## Stage D -- on-time 4-year GPA with national cohort z
G4L <- list()
for (e in 2017L:2022L) {
  X <- MC[AGNO == e & COD_GRADO == 1L, .(MRUN, v1 = v)]
  for (k in 2:4) {
    Xk <- MC[AGNO == e + k - 1L & COD_GRADO == k, .(MRUN, vk = v)]
    setnames(Xk, "vk", paste0("v", k))
    X <- merge(X, Xk, by = "MRUN", all.x = TRUE)
  }
  X[, gpa4 := (v1 + v2 + v3 + v4) / 4][, e := e]
  G4L[[as.character(e)]] <- X[, .(MRUN, e, gpa4)]
}
G4 <- rbindlist(G4L); rm(G4L)
stopifnot("one row per student per entry cohort" = !anyDuplicated(G4, by = c("MRUN", "e")))
G4[!is.na(gpa4), gpa4_z := (gpa4 - mean(gpa4)) / sd(gpa4), by = e]
say("national entry cohorts 2017-2022: %s students | all-four-years share %.4f",
    format(nrow(G4), big.mark = ","), G4[, mean(!is.na(gpa4))])
qe <- data.table(MRUN = PT$MRUN, e = PT$proc + 1L)
PT[, gpa4   := G4[qe, on = .(MRUN, e), x.gpa4]]
PT[, gpa4_z := G4[qe, on = .(MRUN, e), x.gpa4_z]]
PT[, has_4yr := as.integer(!is.na(gpa4))]

## Stage E -- 12th-grade school percentile, on-time year
R12 <- MED[COD_GRADO == 4L]
R12[, pct := (frank(PROM_GRAL, ties.method = "average") - 0.5) / .N, by = .(RBD, AGNO)]
P12 <- R12[, .(pct12 = mean(pct), nsch = uniqueN(RBD), RBD1 = RBD[1L]), by = .(MRUN, AGNO)]
stopifnot("P12 unique per student-year" = !anyDuplicated(P12, by = c("MRUN", "AGNO")))
q4 <- data.table(MRUN = PT$MRUN, AGNO = PT$proc + 4L)
PT[, pct12 := P12[q4, on = .(MRUN, AGNO), x.pct12]]
n_amb12 <- P12[q4, on = .(MRUN, AGNO), nomatch = 0L][nsch > 1L, .N]
say("12th-grade student-years: %s | panel children graded at two 12th-grade schools: %d",
    format(nrow(P12), big.mark = ","), n_amb12)

## Stage F -- rank vs the school's last three promoted classes
S12 <- P12[nsch == 1L, .(MRUN, AGNO, RBD = RBD1)]
CH <- data.table(MRUN = PT$MRUN, g = PT$proc + 4L)
CH[, RBD := S12[CH, on = .(MRUN, AGNO = g), x.RBD]]
say("children with an unambiguous 12th-grade school: %s of %s rows",
    format(CH[!is.na(RBD), .N], big.mark = ","), format(nrow(CH), big.mark = ","))
HS <- unique(CH[!is.na(RBD), .(RBD, g)])
ME <- unique(R12[P12A, on = .(MRUN, AGNO, RBD), nomatch = 0L][
  RBD %in% HS$RBD & AGNO %in% 2017:2024, .(MRUN, gy = AGNO, RBD)])
ME[, `:=`(lo = gy - 3L, hi = gy)]
ME[, m := MC[ME, on = .(MRUN, AGNO >= lo, AGNO <= hi), mean(x.v), by = .EACHI]$V1]
stopifnot("every promoted 12th grader has at least their own graded year" = !anyNA(ME$m))
HX <- HS[rep(seq_len(nrow(HS)), each = 3L)]
HX[, gy := g - rep(1:3, times = nrow(HS))]
HP <- merge(HX, ME[, .(RBD, gy, m)], by = c("RBD", "gy"), allow.cartesian = TRUE)[
  , .(hist_mean = mean(m), hist_sd = sd(m), hist_n = .N, hist_yrs = uniqueN(gy)),
  by = .(RBD, g)]
n_needed <- nrow(HS)
HP <- HP[hist_yrs == 3L & hist_n >= 5L & hist_sd > 0]
say("school-class history pools: %s usable of %s needed (3 class years, n>=5, sd>0)",
    format(nrow(HP), big.mark = ","), format(n_needed, big.mark = ","))
CH[, hist_mean := HP[CH, on = .(RBD, g), x.hist_mean]]
CH[, hist_sd   := HP[CH, on = .(RBD, g), x.hist_sd]]
CH[, hist_n    := HP[CH, on = .(RBD, g), x.hist_n]]
PT[, rank_hist_z := (gpa4 - CH$hist_mean) / CH$hist_sd]
PT[, hist_n := CH$hist_n]
rm(R12, ME, HX, HP, S12, MC, MED, CH, HS, P12, P12A, q4, qe, q1); invisible(gc())

## Stage G -- the room: peer 8th-grade GPA of the 9th-grade census cohort
MATC <- readRDS(data_path("matricula_2011_2023.rds"))
C9 <- unique(MATC[COD_ENSE %in% YM & COD_GRADO == 1L & agno %in% 2017:2022,
                  .(agno, MRUN = as.integer(MRUN), RBD = as.integer(RBD))])
rm(MATC); invisible(gc())
C9[, agno8 := agno - 1L]
C9[, z8 := G8[C9, on = .(MRUN, AGNO = agno8), x.z]]
CS <- C9[, .(S = sum(z8, na.rm = TRUE), K = sum(!is.na(z8)), N = .N), by = .(RBD, agno)]
CC <- C9[, .(nsch = uniqueN(RBD), RBD = RBD[1L], z8 = z8[1L]), by = .(MRUN, agno)]
say("census 9th-grade cohorts: %s | children at two schools in the entry census: %d",
    format(nrow(CS), big.mark = ","), CC[nsch > 1L, .N])
CC <- CC[nsch == 1L]
Q <- data.table(MRUN = PT$MRUN, agno = PT$proc + 1L)
Q[, RBD := CC[Q, on = .(MRUN, agno), x.RBD]]
Q[, z8  := CC[Q, on = .(MRUN, agno), x.z8]]
Q[, S := CS[Q, on = .(RBD, agno), x.S]]
Q[, K := CS[Q, on = .(RBD, agno), x.K]]
Q[, N := CS[Q, on = .(RBD, agno), x.N]]
Q[, own := as.integer(!is.na(z8))]
Q[, Kx := K - own][, Nx := N - 1L]
PT[, in_room9    := as.integer(!is.na(Q$RBD))]
PT[, peer_n      := fifelse(is.na(Q$RBD) | Q$Nx < 1L, NA_integer_, Q$Nx)]
PT[, peer_cov    := fifelse(is.na(Q$RBD) | Q$Nx < 1L, NA_real_, Q$Kx / Q$Nx)]
PT[, peer_gpa8_z := fifelse(is.na(Q$RBD) | Q$Kx < 1L, NA_real_,
                            (Q$S - fifelse(Q$own == 1L, Q$z8, 0)) / Q$Kx)]
rm(C9, CS, CC, Q, G8, P9, G4); invisible(gc())

## Stage H -- exam score and the official NEM/rank scores at proc+5
fc <- data_path("college_outcomes_2018_2026.rds")
stopifnot("college outcome table must match the stamped MD5" = unname(md5sum(fc)) == MD5_CO)
CO <- as.data.table(readRDS(fc))
CO[!is.na(exam_lm), exam_z    := (exam_lm - mean(exam_lm)) / sd(exam_lm), by = proc]
CO[!is.na(nem),     nem_z     := (nem     - mean(nem))     / sd(nem),     by = proc]
CO[!is.na(ranking), ranking_z := (ranking - mean(ranking)) / sd(ranking), by = proc]
q <- data.table(mrun = PT$MRUN, proc = PT$proc + 5L)
o <- CO[q, on = .(mrun, proc)]
PT[, exam_lm := o$exam_lm]
PT[, nem     := o$nem]
PT[, ranking := o$ranking]
PT[, exam_z    := o$exam_z]
PT[, nem_z     := o$nem_z]
PT[, ranking_z := o$ranking_z]
PT[, has_exam := as.integer(!is.na(exam_lm))]
stopifnot("z NA patterns must mirror the raw scores" =
            !any(is.na(PT$exam_z) != is.na(PT$exam_lm)) &&
            !any(is.na(PT$nem_z) != is.na(PT$nem)) &&
            !any(is.na(PT$ranking_z) != is.na(PT$ranking)))
# floor-imputed exam column (spec LOTTERY-MECH-FLOOR-2026-08-18): a no-show
# counts as the process's floor -- the minimum observed national score among
# scored students, which equals the scale minimum (150 raw in processes
# 2021-2022, 100 in 2023-2026) -- in the same taker-z units; every row covered
FL <- CO[!is.na(exam_lm), .(floor_z = (min(exam_lm) - mean(exam_lm)) / sd(exam_lm)), by = proc]
qf <- data.table(proc = PT$proc + 5L)
PT[, exam_z_floor := fifelse(!is.na(exam_z), exam_z, FL[qf, on = .(proc), x.floor_z])]
stopifnot("exam_z_floor must cover every row" = !anyNA(PT$exam_z_floor),
          "floor column equals exam_z wherever a score exists" =
            PT[has_exam == 1L & exam_z_floor != exam_z, .N] == 0L)
rm(FL, qf, CO, q, o); invisible(gc())

## Assemble, guard, save
same <- vapply(names(ORIG), function(cn) identical(ORIG[[cn]], PT[[cn]]), logical(1))
stopifnot("stamped columns must be identical to the stamped panel" = all(same),
          "row count unchanged" = nrow(PT) == 395950L)
saveRDS(PT, OUTF)
say("wrote %s: %s rows, %d columns (%d new)",
    OUTF, format(nrow(PT), big.mark = ","), ncol(PT), ncol(PT) - ncol(ORIG))
for (v in c("in_room9", "has_g9", "has_4yr", "has_exam"))
  say("coverage %-9s winners %.4f | losers %.4f", v,
      PT[won == 1L, mean(get(v))], PT[won == 0L, mean(get(v))])
say("peer measure present: winners %.4f | losers %.4f",
    PT[won == 1L, mean(!is.na(peer_gpa8_z))], PT[won == 0L, mean(!is.na(peer_gpa8_z))])
say("rank_hist_z present:  winners %.4f | losers %.4f",
    PT[won == 1L, mean(!is.na(rank_hist_z))], PT[won == 0L, mean(!is.na(rank_hist_z))])
say("official scores present (registrants): nem %.4f | ranking %.4f",
    PT[, mean(!is.na(nem))], PT[, mean(!is.na(ranking))])
say("total elapsed %.1f min", (proc.time() - t0)[["elapsed"]] / 60)
```
