---
title: "Piece 3 -- Why Winning Hurts: the Room, the Transcript, and the Exam"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Libraries ---------------------------------------------------------------
library(data.table)
library(fixest)
library(tools)

# RNG contract: every results chunk re-seeds 20260919 at its own top, so each
# exhibit's random draws do not depend on which chunks ran before it. Chunks
# share the frame objects built upstream, so run the document top to bottom.
set.seed(20260919)
B_RI <- 9999L

# The panel (built by 00_build_mech_panel.Rmd). A copy next to this file wins;
# otherwise the shared data folder's copy of record. Pinned by MD5 to the
# validated build of record.
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
panel_file <- "lottery_panel_mech.rds"
if (!file.exists(panel_file))
  panel_file <- file.path(home, "Dropbox", "Effect SAE on SUA", "data", "lottery_panel_mech.rds")
stopifnot("panel not found next to the code or in the shared data folder" =
            file.exists(panel_file),
          "mech panel must be the validated build of record" =
            unname(md5sum(panel_file)) == "6296a6de9feb98a18b5fc850e7b3a275")
dir.create("tables", showWarnings = FALSE)

fmt  <- function(x) sprintf("%.4f", x)
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# the primary frame and the estimator: copied VERBATIM from certified Piece 2
# (1_effect_lottery.Rmd). make_frame: first-choice applications, lotteries with
# >= W winners, loser violations <= pur, nearest losers capped at the winner
# count. core: within-lottery ITT, precision-weighted across lotteries; the
# closed form is asserted equal to the lottery-fixed-effects regression at every
# run; stars from reshuffling winners within each lottery (winner count fixed).
# For an NA-bearing outcome, core drops the NA rows and any stratum left
# single-armed, then reshuffles within what remains -- the same certified
# treatment the Piece 2 balance test applied to the continuous 8th-grade GPA z.
make_frame <- function(P, W, pur, cap, fc) {
  M <- copy(P); if (fc) M <- M[PREF == 1L]
  CU <- M[won == 1L, .(cutoff = max(loteria), n_win = .N), by = key]; M <- merge(M, CU, by = "key")[n_win >= W]
  M[, viol := won == 0L & loteria < cutoff]
  PU <- M[won == 0L, .(n_los = .N, n_viol = sum(viol)), by = key]
  M <- merge(M, PU, by = "key")[n_los >= 1L][n_viol / n_los <= pur]
  L <- M[won == 0L & loteria >= cutoff]
  if (cap) { setorder(L, key, loteria); L[, rk := seq_len(.N), by = key]; L <- L[rk <= n_win] }
  rbind(M[won == 1L], L, fill = TRUE)
}
core <- function(DT, y) {
  DT <- DT[!is.na(get(y))]; setorder(DT, key)
  KX <- DT[, .(i0 = .I[1], n = .N, m = sum(won)), by = key][m > 0L & m < n]
  DT <- DT[key %in% KX$key]; setorder(DT, key)
  KX <- DT[, .(i0 = .I[1], n = .N, m = sum(won)), by = key]
  denom <- KX[, sum(m * (n - m) / n)]; yv <- as.numeric(DT[[y]])
  base <- DT[, .(b = mean(get(y)) * sum(won)), by = key]$b
  obs <- (sum(yv * DT$won) - sum(base)) / denom
  pn <- numeric(B_RI)
  for (r in seq_len(nrow(KX))) { i0 <- KX$i0[r]; n <- KX$n[r]; m <- KX$m[r]
    yk <- yv[i0:(i0 + n - 1L)]; pn <- pn + .colSums(yk[replicate(B_RI, sample.int(n, m))], m, B_RI) }
  perm <- (pn - sum(base)) / denom
  fit <- feols(as.formula(paste0(y, " ~ won | key")), data = DT, cluster = ~sch_year, fixef.rm = "none")
  ct <- summary(fit)$coeftable
  stopifnot("closed form must equal the regression point" = abs(unname(ct["won", 1]) - obs) < 1e-8)
  data.table(itt = obs, se = ct["won", 2],
             ri_p = (1 + sum(abs(perm) >= abs(obs) - 1e-12)) / (B_RI + 1),
             n = nrow(DT), k = uniqueN(DT$key),
             lm = DT[won == 0L, mean(get(y))], win_mean = DT[won == 1L, mean(get(y))])
}
```

## Design

Piece 2 established that winning a contested 9th-grade seat reduces on-time
entry into the college system. Piece 3 asks why, with the same frame, the same
estimator, and the same inference -- only the outcomes change. The frame is the
certified Piece 2 headline frame (first-choice applications, at least five
winners per stratum, the 10% violation screen, nearest losers capped, one
stratum = classroom by year by income tier), asserted below to contain exactly
the certified 45,123 applications in 1,215 strata. The outcomes come in two
kinds. The transcript outcomes are read from the education ministry's school
records, which cover every enrolled child in Chile, so they need no college
participation of any kind: the strength of the room the child enters, their own
grades, their position in their class, and their standing against the bar their
school's last three graduating classes set -- the object the admission
formula's rank component prices. The score outcomes are read from the college
admission records and exist only for children who show up to that system;
winning itself moves who shows up, so those columns sit in a separate table,
their denominators are reported per arm, and a registered check compares the arms
on their predetermined records inside each denominator, which is what licenses
reading the exam column. Two further registered checks -- scoring every absent
child at the scale floor, and the score at which an assumed value for the absent
would flip the comparison -- bound how far the absent could move it. Standard errors cluster on school by year
and are reported for scale; significance comes from randomization inference,
reshuffling winners within each stratum 9,999 times with winner counts fixed.

## The room and the transcript

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
P <- as.data.table(readRDS(panel_file))
stopifnot("every child must carry the income-tier flag" = P[, all(!is.na(prioritario))])
P[, has_nem := as.integer(!is.na(nem))]   # holds an official NEM score at proc+5
Ps <- copy(P)[, key := paste(key, prioritario)]   # stratum = classroom x year x tier
As <- make_frame(Ps, 5L, 0.10, TRUE, TRUE)
stopifnot("frame must equal the certified headline frame" =
            nrow(As) == 45123L && uniqueN(As$key) == 1215L,
          "no test-track winners may sit in the frame" =
            As[won == 1L & test_track == TRUE, .N] == 0L)
cat(sprintf("Frame: %s applications, %s strata (asserted equal to the certified headline frame).\n\n",
            format(nrow(As), big.mark = ","), format(uniqueN(As$key), big.mark = ",")))

OUT1 <- c("peer_gpa8_z", "gpa9_z", "pct9", "gpa4_z", "pct12", "rank_hist_z")
res1 <- rbindlist(lapply(OUT1, function(y) core(As, y)[, outcome := y]))
fwrite(res1, file.path("tables", "mech_transcript_values.csv"))
tab1 <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect of Winning an SAE Lottery on the Classroom and the Transcript}",
  "\\label{tab: mech transcript}\n",
  "   \\centerline{\\scalebox{0.80}{\\begin{tabular}{lcccccc}\n      \\toprule\n",
  "                   & Peer 8th GPA & 9th GPA & 9th pctile & 4-yr GPA & 12th pctile & Rank vs.\\ hist.\\\\\n",
  "                   & (1) & (2) & (3) & (4) & (5) & (6)\\\\\n      \\midrule\n",
  row_tex("Won the lottery", paste0(fmt(res1$itt), star(res1$ri_p))),
  row_tex("", paste0("(", fmt(res1$se), ")")),
  "      \\midrule\n",
  row_tex("Loser-arm mean", fmt(res1$lm)),
  row_tex("Observations", format(res1$n, big.mark = ",")),
  row_tex("Strata (classroom x year x tier)", format(res1$k, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: same frame, estimator, clustering, and ",
  "randomization inference as the on-time headline (first-choice applications, cohorts ",
  "2016--2021; within-stratum comparison of winners and nearest losers). Column (1): ",
  "leave-one-out mean 8th-grade GPA (national z, application year) of the 9th-grade ",
  "cohort at the school the child attends in the entry year. (2): own 9th-grade GPA, ",
  "national z. (3): position within the child's school-grade graded cohort, 0--1. (4): mean ",
  "GPA over the four on-time high-school years (all four required), national ",
  "entry-cohort z. (5): position within the 12th-grade school cohort, on-time year. ",
  "(6): 4-year GPA measured against the pooled final GPAs of the three classes ",
  "promoted from the child's 12th-grade school just before their own, in that ",
  "group's SDs. Observations vary by column: each outcome exists only where its ",
  "underlying record exists, and strata left with only winners or only losers after ",
  "that restriction drop out. The non-NA share of every column by arm is in the ",
  "accompanying coverage file (mech_column_coverage_values.csv). School-by-year ",
  "clustered standard errors in parentheses. Stars from randomization ",
  "inference (9,999 within-stratum reshuffles): $^{***}$p$<$0.01, $^{**}$p$<$0.05, ",
  "$^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab1, file.path("tables", "mech_transcript.tex"))
cat(tab1)
```

## The exam and the official scores

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
OUT2 <- c("exam_z", "nem_z", "ranking_z")
res2 <- rbindlist(lapply(OUT2, function(y) core(As, y)[, outcome := y]))
fwrite(res2, file.path("tables", "mech_scores_values.csv"))
tab2 <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect of Winning an SAE Lottery on the Exam and the Official Admission Scores}",
  "\\label{tab: mech scores}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lccc}\n      \\toprule\n",
  "                   & Exam score & Official NEM & Official rank\\\\\n",
  "                   & (1) & (2) & (3)\\\\\n      \\midrule\n",
  row_tex("Won the lottery", paste0(fmt(res2$itt), star(res2$ri_p))),
  row_tex("", paste0("(", fmt(res2$se), ")")),
  "      \\midrule\n",
  row_tex("Loser-arm mean", fmt(res2$lm)),
  row_tex("Observations", format(res2$n, big.mark = ",")),
  row_tex("Strata (classroom x year x tier)", format(res2$k, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: same frame, estimator, clustering, and ",
  "randomization inference as the on-time headline. All three columns in national ",
  "within-process SDs, measured at the cohort's on-time college process (application ",
  "year + 5). Column (1): average of the language and mathematics exam scores, ",
  "existing only for children who sat an exam. (2)--(3): DEMRE's official NEM and ",
  "rank scores, existing only for score-carrying college-process registrants. ",
  "Winning moves these denominators (per-arm shares and their estimated differences ",
  "are in the accompanying coverage files). Strata left with only winners or only ",
  "losers after the restriction drop out. School-by-year clustered standard errors ",
  "in parentheses. Stars from randomization inference (9,999 within-stratum ",
  "reshuffles): $^{***}$p$<$0.01, $^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab2, file.path("tables", "mech_scores.tex"))
cat(tab2)
```

## Coverage: who each column can see

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
COVV <- c("in_room9", "has_g9", "has_4yr", "has_exam", "has_nem")
cov <- rbindlist(lapply(COVV, function(y) core(As, y)[, margin := y]))
fwrite(cov, file.path("tables", "mech_coverage_values.csv"))
for (i in seq_len(nrow(cov))) { b <- cov[i]
  cat(sprintf("%s: winners %.4f, losers %.4f, difference %+.4f%s (%.4f), RI p %.4f.\n\n",
              b$margin, b$win_mean, b$lm, b$itt, gsub("[$^{}]", "", star(b$ri_p)), b$se, b$ri_p)) }
csh <- rbindlist(lapply(c(OUT1, OUT2), function(y) data.table(
  column = y,
  win_share  = As[won == 1L, mean(!is.na(get(y)))],
  lose_share = As[won == 0L, mean(!is.na(get(y)))])))
fwrite(csh, file.path("tables", "mech_column_coverage_values.csv"))
for (i in seq_len(nrow(csh)))
  cat(sprintf("Column %s: non-NA share winners %.4f, losers %.4f.\n\n",
              csh$column[i], csh$win_share[i], csh$lose_share[i]))
```

## The exam column: who it compares, and how far the absent could move it

The takers-only exam column conditions on a decision the child made after the
draw, so the first question is whether the two arms inside that column are still
comparable. The tilt check answers it: the predetermined traits of winners
versus losers inside the sitter pool and inside the no-show pool -- a
composition measurement of post-draw-defined groups, stated as such, never a
causal effect. The remaining two exhibits bound how far the absent children
could move the column. The floor scores every absent child as if they sat and
answered nothing (the scale minimum, recovered from the records below), so every
frame child counts. The crossing point uses the fact that the estimate is
exactly linear in the common score assigned to the absent, so one number
says what the absent would all have to score for the exam difference to reach
zero.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
ef <- core(As, "exam_z_floor")[, outcome := "exam_z_floor"]
fwrite(ef, file.path("tables", "mech_exam_floor_values.csv"))
cat(sprintf("Exam, no-show = floor: %+.4f%s (%.4f), RI p %.4f, n %s, strata %s, loser-arm mean %.4f.\n\n",
            ef$itt, gsub("[$^{}]", "", star(ef$ri_p)), ef$se, ef$ri_p,
            format(ef$n, big.mark = ","), format(ef$k, big.mark = ","), ef$lm))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
tl <- rbind(
  core(As[has_exam == 1L], "gpa8_z")[, `:=`(trait = "gpa8_z", subset = "sitters")],
  core(As[has_exam == 0L], "gpa8_z")[, `:=`(trait = "gpa8_z", subset = "no-shows")],
  core(As[has_exam == 1L], "female")[, `:=`(trait = "female", subset = "sitters")],
  core(As[has_exam == 0L], "female")[, `:=`(trait = "female", subset = "no-shows")])
fwrite(tl, file.path("tables", "mech_tilt_values.csv"))
for (i in seq_len(nrow(tl))) { b <- tl[i]
  cat(sprintf("Tilt, %s among %s: winners %.4f, losers %.4f, difference %+.4f%s (%.4f), RI p %.4f, n %s.\n\n",
              b$trait, b$subset, b$win_mean, b$lm, b$itt,
              gsub("[$^{}]", "", star(b$ri_p)), b$se, b$ri_p, format(b$n, big.mark = ","))) }
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
grid <- c(-4, -3, -2, -1, 0)
cr <- rbindlist(lapply(grid, function(v) {
  Av <- copy(As)[, ev := fifelse(!is.na(exam_z), exam_z, v)]
  core(Av, "ev")[, v := v]
}))
fwrite(cr, file.path("tables", "mech_crossing_values.csv"))
b_slope <- (cr[v == 0, itt] - cr[v == -4, itt]) / 4
a_int   <- cr[v == 0, itt]
v_star  <- -a_int / b_slope
stopifnot("estimate must be exactly linear in v" =
            max(abs(cr$itt - (a_int + b_slope * cr$v))) < 1e-10)
for (i in seq_len(nrow(cr))) { b <- cr[i]
  cat(sprintf("Assumed score for every absent child, %+.0f taker-SD: %+.4f%s (%.4f), RI p %.4f.\n\n",
              b$v, b$itt, gsub("[$^{}]", "", star(b$ri_p)), b$se, b$ri_p)) }

# Raw-point translation of the crossing point and of the floor, recovered from
# the panel itself: within a process, raw = mu + sd * z holds row by row, so two
# rows with distinct z identify the national mean and SD (asserted exact against
# every scored row). No scale constant is typed by hand.
SC <- As[!is.na(exam_lm) & !is.na(exam_z), .(po = proc + 5L, exam_lm, exam_z)]
sc <- SC[, {
  i <- which.min(exam_z); j <- which.max(exam_z)
  sdv <- (exam_lm[j] - exam_lm[i]) / (exam_z[j] - exam_z[i])
  mu  <- exam_lm[i] - sdv * exam_z[i]
  stopifnot("raw = mu + sd*z must hold exactly on every scored row" =
              max(abs(exam_lm - (mu + sdv * exam_z))) < 1e-6)
  .(mu = mu, sdv = sdv, raw_vstar = mu + v_star * sdv)
}, by = po][order(po)]
fz <- unique(As[has_exam == 0L, .(po = proc + 5L, floor_z = exam_z_floor)])
sc <- merge(sc, fz, by = "po")[, raw_floor := mu + floor_z * sdv]
fwrite(sc, file.path("tables", "mech_scale_translation_values.csv"))
cat(sprintf("The estimate is exactly linear in the assumed score (slope %+.6f per SD); it reaches zero at %+.3f taker-SD.\n\n",
            b_slope, v_star))
for (i in seq_len(nrow(sc))) cat(sprintf(
  "Process %d: scale floor %.0f raw points (%.3f taker-SD); the crossing point is %.0f raw points.\n\n",
  sc$po[i], sc$raw_floor[i], sc$floor_z[i], sc$raw_vstar[i]))
```
