---
title: "Piece 2 -- Building the Lottery Panel"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

This document builds `lottery_panel.rds`: one row per child per school lottery at the
9th-grade door, application years 2016--2021, with each child's on-time outcomes --
promoted out of 12th grade on schedule, and the college outcomes five years after
application. The analysis file `1_effect_lottery.Rmd` reads only this panel.

**Inputs.** All resolve through one chain: the shared data folder
(`~/Dropbox/Effect SAE on SUA/data`), then the project's `code/data` mirror, then the
Piece 1 `data/` folder. The packed files are built from the public releases by
`code/piece1/00a_pack_raw_data.R`.

- `sae_admissions.rds` (packed) -- SAE applications with lottery numbers and priority
  flags (C1), placements (D1), and applicant characteristics (B1), processes 2016--2025.
- `paes_2026.rds` (packed) -- the 2026 college process: registrants with scores,
  selection states, enrolment.
- `data_students_2004_2025.rds` (shared, canonical) -- the college system's student
  file: applied/assigned/enrolled plus scores, processes through 2025.
- `entrant_composition_panel.rds` (shared; Piece 1's shipped panel) -- supplies the
  over-demanded school list, so both pieces use one list from one artifact.
- `panel_performance_2002_2023.rds` (shared, canonical) -- the graded panel: 8th-grade
  GPA in the application year, and 12th-grade promotion records through 2023.
- `rendimiento_2024_2025.rds` (packed) -- 12th-grade promotion records for 2024--2025,
  so grad on time reaches the last cohorts' on-time years.
- `matricula_2011_2023.rds` and `matricula_2024_2025.rds` (packed) -- the annual
  enrolment census and its supplement: the take-up and persistence flags.

**Output.** `lottery_panel.rds`, one row per contested regular application at a tagged
9th-grade school: lottery id, child id, won/lost, preference rank, lottery number,
application year, grad on time (promoted out of 12th grade in the on-time year,
application + 4; no record or not promoted = 0), the three college outcomes at
application year + 5 (non-appearance = 0), a tracked flag (the child has any
college-process record that year), the
low-income and gender flags, the child's 8th-grade GPA in the application year
(z-scored nationally within year; missing where no graded record matches), and the
census take-up flags: enrolled at the lottery school in each high-school year (entry
through 12th grade), enrolled at the child's own SAE-assigned school at entry (missing
where unassigned), enrolled at any school at entry, and the years-attended count. The winner/purity/cap filters are applied at analysis time,
so every robustness dial reads this one file.

```{r echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({ library(data.table) })
setDTthreads(4); options(warn = 1)
t0 <- proc.time()
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
ROOT <- file.path(home, "Dropbox", "Effect of Centralization")
OUTF <- "lottery_panel.rds"
data_path <- function(f) {
  for (p in c(file.path(home, "Dropbox", "Effect SAE on SUA", "data", f),
              file.path(ROOT, "code", "data", f),
              file.path(ROOT, "Updated Paper Spine", "Piece 1", "data", f)))
    if (file.exists(p)) return(p)
  stop(sprintf("input file not found in any data location: %s", f))
}
say <- function(...) { cat(sprintf(...), "\n"); flush.console() }
```

## Stage 1 -- the college outcome table

One row per student per college process year 2018--2026 with three 0/1 outcomes.
*Applied*: the student submitted a preference list (or enrolled). *Assigned*: the
system offered them a seat. *Enrolled*: they matriculated. Years 2018--2025 come from
the canonical student file; 2026 comes from the packed raw files under definitions
validated against process 2025 (assigned 158,218/158,218 exact; enrolled exact). The
exam score (`exam_lm`) recodes each component to missing before averaging, so a student
absent from one exam carries their sat score; earlier builds averaged first, which
halved the score of single-exam absentees in the 2018--2025 rows (diagnosed in the
2026-08-16 review). The three outcome columns are asserted unchanged against the
verified panel wherever that reference exists.

```{r echo=FALSE, warning=FALSE, message=FALSE}
D <- as.data.table(readRDS(data_path("data_students_2004_2025.rds"))); setnames(D, toupper(names(D)))
stopifnot("student file has APPLIED/ASSIGNED/ENROLLED" =
            all(c("APPLIED", "ASSIGNED", "ENROLLED") %in% names(D)))
P1 <- D[YEAR %in% 2018:2025 & !is.na(MRUN),
        .(mrun = as.integer(MRUN), proc = as.integer(YEAR),
          applied = as.integer(APPLIED == 1),
          assigned = as.integer(ASSIGNED == 1), enrolled = as.integer(ENROLLED == 1),
          L = as.numeric(LENGUAJE_ACTUAL), M = as.numeric(MATEMATICA_ACTUAL),
          nem = as.numeric(NEM), ranking = as.numeric(RANKING))]
rm(D); invisible(gc())
P1[is.na(applied), applied := 0L]
# a 0 is absence, never a score: recode each exam COMPONENT before averaging
P1[!is.na(L) & L <= 0, L := NA_real_][!is.na(M) & M <= 0, M := NA_real_]
P1[, exam_lm := rowMeans(cbind(L, M), na.rm = TRUE)][is.nan(exam_lm), exam_lm := NA_real_]
P1[, c("L", "M") := NULL]
P1[!is.na(nem) & nem <= 0, nem := NA_real_]
P1[!is.na(ranking) & ranking <= 0, ranking := NA_real_]
P1 <- unique(P1, by = c("mrun", "proc"))
say("student-file rows 2018-2025: %s", format(nrow(P1), big.mark = ","))

PAES <- readRDS(data_path("paes_2026.rds"))
A <- PAES$a
num <- function(x) as.numeric(sub(",", ".", as.character(x), fixed = TRUE))
for (c in c("PTJE_NEM", "PTJE_RANKING", "CLEC_REG_ACTUAL", "MATE1_REG_ACTUAL")) set(A, j = c, value = num(A[[c]]))
A[CLEC_REG_ACTUAL <= 0, CLEC_REG_ACTUAL := NA_real_][MATE1_REG_ACTUAL <= 0, MATE1_REG_ACTUAL := NA_real_]
A[PTJE_NEM <= 0, PTJE_NEM := NA_real_][PTJE_RANKING <= 0, PTJE_RANKING := NA_real_]
A <- unique(A, by = "MRUN")
asg26  <- PAES$c[ESTADO_PREF == 24, unique(MRUN)]
appl26 <- unique(PAES$c$MRUN)
enr26  <- unique(PAES$d$MRUN)
P2 <- A[, .(mrun = as.integer(MRUN), proc = 2026L,
            applied  = as.integer(MRUN %in% appl26 | MRUN %in% enr26),
            assigned = as.integer(MRUN %in% asg26),
            enrolled = as.integer(MRUN %in% enr26),
            nem = PTJE_NEM, ranking = PTJE_RANKING,
            exam_lm = rowMeans(cbind(CLEC_REG_ACTUAL, MATE1_REG_ACTUAL), na.rm = TRUE))]
P2[is.nan(exam_lm), exam_lm := NA_real_]
rm(PAES, A); invisible(gc())

CO <- rbind(P1, P2, fill = TRUE); rm(P1, P2)
stopifnot("proc range 2018-2026" = min(CO$proc) == 2018L && max(CO$proc) == 2026L,
          "outcomes are 0/1" = all(CO$applied %in% 0:1) && all(CO$assigned %in% 0:1) && all(CO$enrolled %in% 0:1),
          "applied >= assigned on totals" = sum(CO$applied) >= sum(CO$assigned))
setkey(CO, mrun, proc)
say("outcome table: %s rows | applied %s assigned %s enrolled %s",
    format(nrow(CO), big.mark = ","), format(sum(CO$applied), big.mark = ","),
    format(sum(CO$assigned), big.mark = ","), format(sum(CO$enrolled), big.mark = ","))

# guard: the three outcome columns must equal the verified panel wherever it exists
ref <- file.path(ROOT, "explorations", "2026-08-12_enrollment_panel_probe", "output",
                 "college_outcomes_2018_2026.rds")
if (file.exists(ref)) {
  R <- as.data.table(readRDS(ref))
  MM <- merge(CO[, .(mrun, proc, applied, assigned, enrolled)],
              R[, .(mrun, proc, a = applied, s = assigned, e = enrolled)],
              by = c("mrun", "proc"), all = TRUE)
  stopifnot("outcomes must match the verified panel row for row" =
              MM[applied != a | assigned != s | enrolled != e, .N] == 0L &&
              MM[is.na(applied) | is.na(a), .N] == 0L)
  say("outcome columns identical to the verified reference panel (%s rows)", format(nrow(MM), big.mark = ","))
  rm(R, MM); invisible(gc())
} else say("reference panel not on this machine; identity guard skipped")
saveRDS(CO, "college_outcomes_2018_2026.rds")
```

## Stage 2 -- the over-demanded school list (from Piece 1)

Both pieces use one list: the schools in Piece 1's shipped panel. Its 9th-grade cells
are the schools whose 9th-grade seats were over-demanded in every observed year.

This list supersedes the earlier exploration-era list (`panel_always_overdemanded_v2`)
by the consistency order of 2026-08-16: both pieces read one list from one shipped
artifact. The delta against the earlier list is printed whenever that file is present.

```{r echo=FALSE, warning=FALSE, message=FALSE}
EC <- as.data.table(readRDS(data_path("entrant_composition_panel.rds")))
CELLS <- unique(EC[cod_nivel == 9L, .(rbd)]); rm(EC); invisible(gc())
say("over-demanded 9th-grade schools (Piece 1 list): %s", format(nrow(CELLS), big.mark = ","))
v2 <- file.path(ROOT, "explorations", "2026-08-05_priority_quota", "output",
                "panel_always_overdemanded_v2.csv")
if (file.exists(v2)) {
  O <- unique(fread(v2)[cod_nivel == 9L, .(rbd)])
  say("list provenance: earlier v2 list %d schools | this list %d | removed %d | added %d",
      nrow(O), nrow(CELLS), length(setdiff(O$rbd, CELLS$rbd)), length(setdiff(CELLS$rbd, O$rbd)))
  rm(O)
}
```

## Stage 3 -- the lottery frame

For each application year 2016--2021: read the packed applications and placements; keep
9th-grade applications at listed schools; keep the applications that were a fair coin
flip (drop rows the system auto-added for continuing students, and children holding
sibling, staff-child, returning-student, already-enrolled, or special-education
priority -- the 2016 pilot files carry one priority code instead, kept for its
prioritario and no-priority values); keep contested applications (ranked at least as
high as the child's final placement). A lottery is one classroom in one year (one
school in one year for the 2016 pilot). Winner = the placement file put the child
there. One case needs a stated rule: a placed child whose placement school appears on
none of their application rows (this occurs only in the 2016 pilot file) counts as
having lost every listed lottery -- the contested screen cannot bind for them, so all
their listed applications are kept as losses; the build prints how many children and
frame rows this rule touches. Each child's outcomes are looked up at application year + 5; a child with no
record that year has all three outcomes 0 and tracked 0. The low-income flag and
gender come from the applicant-characteristics table of the same application year.

```{r echo=FALSE, warning=FALSE, message=FALSE}
SAE <- readRDS(data_path("sae_admissions.rds"))
stopifnot("packed SAE file carries the lottery columns" =
            all(c("loteria_original", "tipo_prioridad", "es_pie", "vintage") %in% names(SAE$c1)),
          "packed SAE file carries the admission-ordering columns" =
            all(c("orden_pie", "orden_alta_exigencia_transicion") %in% names(SAE$c1)),
          "packed SAE file carries the placed course" = "cod_curso_admitido" %in% names(SAE$d1),
          "B1 must cover every application year" = all(2016:2021 %in% unique(SAE$b1$proc)))

build_raw <- function(pr) {
  C <- SAE$c1[proc == pr & cod_nivel == 9L]; D <- SAE$d1[proc == pr & cod_nivel == 9L]
  is16 <- C$vintage[1] == "2016"
  C[is.na(agregada_por_continuidad), agregada_por_continuidad := 0L]
  if (is16) {
    C[, regular := agregada_por_continuidad != 1L & tipo_prioridad %in% c(4L, 7L)]
    DA <- unique(D[!is.na(rbd_admitido), .(mrun, rbd_a = rbd_admitido)], by = "mrun")
    AR <- merge(C[, .(mrun, rbd, PREF = preferencia_postulante)],
                DA[, .(mrun, rbd = rbd_a, hit = 1L)], by = c("mrun", "rbd"), all.x = TRUE)
    ARr <- AR[hit == 1L, .(a_rank = min(PREF)), by = mrun]
    C <- merge(C, ARr, by = "mrun", all.x = TRUE); C[, a_rank := as.double(a_rank)][is.na(a_rank), a_rank := Inf]
    C <- merge(C, DA, by = "mrun", all.x = TRUE)
    C[, won := !is.na(rbd_a) & rbd == rbd_a]; C[, contested := preferencia_postulante <= a_rank]
    C[, key := paste(pr, rbd)]
  } else {
    DA <- unique(D[!is.na(rbd_admitido), .(mrun, rbd_a = rbd_admitido, curso_a = cod_curso_admitido)], by = "mrun")
    AR <- merge(C[, .(mrun, rbd, cod_curso, PREF = preferencia_postulante)],
                DA[, .(mrun, rbd = rbd_a, cod_curso = curso_a, hit = 1L)],
                by = c("mrun", "rbd", "cod_curso"), all.x = TRUE)
    ARr <- AR[hit == 1L, .(a_rank = min(PREF)), by = mrun]
    C <- merge(C, ARr, by = "mrun", all.x = TRUE); C[, a_rank := as.double(a_rank)][is.na(a_rank), a_rank := Inf]
    C[, regular := agregada_por_continuidad != 1L &
        (is.na(prioridad_matriculado) | prioridad_matriculado == 0) &
        (is.na(prioridad_hermano) | prioridad_hermano == 0) &
        (is.na(prioridad_hijo_funcionario) | prioridad_hijo_funcionario == 0) &
        (is.na(prioridad_exalumno) | prioridad_exalumno == 0) &
        (is.na(es_pie) | es_pie == 0)]
    C <- merge(C, DA[, .(mrun, rbd_a, curso_a)], by = "mrun", all.x = TRUE)
    C[, won := !is.na(rbd_a) & rbd == rbd_a & cod_curso == curso_a]
    C[, contested := preferencia_postulante <= a_rank]
    C[, key := paste(pr, rbd, cod_curso)]
  }
  M <- C[regular == TRUE & contested == TRUE & rbd %in% CELLS$rbd & !is.na(loteria_original)]
  ol <- unique(C[!is.na(rbd_a) & is.infinite(a_rank), mrun])   # placed at an off-list school
  if (length(ol)) say(paste0("  %d off-list placements: %d placed children have a placement school on ",
    "none of their application rows; kept as losses on every listed application (%d frame rows)"),
    pr, length(ol), M[mrun %in% ol, .N])
  nb <- function(x) !is.na(x) & trimws(as.character(x)) != "" & trimws(as.character(x)) != "0"
  M[, .(key, MRUN = as.integer(mrun), RBD = rbd, won = as.integer(won),
        PREF = as.integer(preferencia_postulante), loteria = as.numeric(loteria_original),
        proc = pr, vintage = fifelse(is16, "2016", "std"),
        rbd_a,                                        # where the child was actually placed
        tipo = tipo_prioridad,                        # 2016 pilot's processed priority code
        test_track = nb(orden_alta_exigencia_transicion),  # academic test-track ordering present
        pie_track  = nb(orden_pie))]                       # special-education ordering present
}
RAW <- rbindlist(lapply(2016:2021, build_raw))
say("raw frame: %s applications | %s lotteries", format(nrow(RAW), big.mark = ","),
    format(uniqueN(RAW$key), big.mark = ","))
print(RAW[, .(apps = .N, lotteries = uniqueN(key), first_choice = sum(PREF == 1L)), by = proc][order(proc)])

# outcomes at application year + 5; tracked = any college-process record that year
q <- data.table(mrun = RAW$MRUN, proc = RAW$proc + 5L); o <- CO[q, on = .(mrun, proc)]
RAW[, tracked := as.integer(!is.na(o$applied))]
RAW[, `:=`(applied = fifelse(is.na(o$applied), 0L, o$applied),
           assigned = fifelse(is.na(o$assigned), 0L, o$assigned),
           enrolled = fifelse(is.na(o$enrolled), 0L, o$enrolled),
           sch_year = paste(proc, RBD))]
# low-income flag and gender from the applicant table of the same application year.
# An NA flag would mean a frame child missing from the applicant table -- that fails
# loudly below; duplicates must agree on the flags before deduplication keeps one row.
bdup <- SAE$b1[, .(nfl = uniqueN(paste(prioritario, es_mujer))), by = .(proc, mrun)][nfl > 1L, .N]
stopifnot("duplicate B1 rows must agree on the flags" = bdup == 0L)
B <- unique(SAE$b1[, .(proc, MRUN = as.integer(mrun), prioritario = as.integer(prioritario == 1),
                       female = as.integer(es_mujer == 1))], by = c("proc", "MRUN"))
RAW <- merge(RAW, B, by = c("proc", "MRUN"), all.x = TRUE)
rm(SAE, B, o, q, CO); invisible(gc())
stopifnot("every frame child must carry the income-tier flag" = !anyNA(RAW$prioritario),
          "every frame child must carry the gender flag" = !anyNA(RAW$female))

# 2016 pilot: the analysis stratifies on the predetermined applicant flag; the pilot's
# own processed priority code (4 = prioritario tier, 7 = no-priority tier) is printed
# against it so the alignment is on the record, then the code column is dropped.
say("2016 pilot crosstab, processed priority code x applicant low-income flag:")
print(RAW[vintage == "2016", .N, by = .(tipo, prioritario)][order(tipo, prioritario)])
RAW[, tipo := NULL]

# predetermined achievement: the child's 8th-grade GPA in the application year, from the
# canonical graded panel, z-scored nationally within year. Missing where no graded
# record matches (match rates print; missingness is legitimate, never filled).
PF <- as.data.table(readRDS(data_path("panel_performance_2002_2023.rds")))
G8 <- PF[COD_GRADO == 8L & COD_ENSE == 110L & AGNO %in% 2016:2021 &
         !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7, .(MRUN, AGNO, PROM_GRAL)]
YM <- c(310L, 410L, 510L, 610L, 710L, 810L, 910L)     # 12th-grade school types (media)
G12 <- unique(PF[COD_ENSE %in% YM & COD_GRADO == 4L & AGNO %in% 2020:2023,
                 .(MRUN, agno = AGNO, sit = SIT_FIN_R)])
rm(PF); invisible(gc())
G8 <- G8[, .(gpa8 = mean(PROM_GRAL)), by = .(MRUN, AGNO)]
G8[, gpa8_z := (gpa8 - mean(gpa8)) / sd(gpa8), by = AGNO]
RAW[, gpa8_z := G8[RAW, on = .(MRUN, AGNO = proc), x.gpa8_z]]
rm(G8); invisible(gc())
say("8th-grade GPA matched: winners %.4f | losers %.4f",
    RAW[won == 1L, mean(!is.na(gpa8_z))], RAW[won == 0L, mean(!is.na(gpa8_z))])

# take-up and persistence: enrolment at the lottery school in each high-school year
# (entry = application+1 = 9th grade, through entry+3 = 12th grade), enrolment at the
# child's own SAE-assigned school at entry (NA where unassigned), and enrolment at any
# school at entry. All from the census April snapshots; the last cohorts' 12th-grade
# years come from the 2024-2025 supplement pack.
MATC <- readRDS(data_path("matricula_2011_2023.rds"))
MATS <- readRDS(data_path("matricula_2024_2025.rds"))
ENT <- unique(rbind(MATC[agno %in% 2017:2023, .(agno, MRUN, RBD)],
                    MATS[, .(agno, MRUN, RBD)]))
rm(MATC, MATS); invisible(gc())
for (k in 0:3) {
  qk <- data.table(MRUN = RAW$MRUN, RBD = RAW$RBD, agno = RAW$proc + 1L + k)
  RAW[, (paste0("at_y", k + 1L)) := as.integer(!is.na(ENT[qk, on = .(MRUN, RBD, agno), which = TRUE]))]
}
qa <- data.table(MRUN = RAW$MRUN, RBD = RAW$rbd_a, agno = RAW$proc + 1L)
RAW[, at_assigned := fifelse(is.na(rbd_a), NA_integer_,
                             as.integer(!is.na(ENT[qa, on = .(MRUN, RBD, agno), which = TRUE])))]
E2 <- unique(ENT[, .(MRUN, agno)])
RAW[, in_any := as.integer(!is.na(E2[data.table(MRUN = RAW$MRUN, agno = RAW$proc + 1L),
                                     on = .(MRUN, agno), which = TRUE]))]
RAW[, yrs_at_school := at_y1 + at_y2 + at_y3 + at_y4]
rm(ENT, E2, qa, qk); invisible(gc())

# graduation on time: promoted out of 12th grade (final status P) in the on-time year
# (application + 4). 12th-grade records 2020-2023 come from the canonical graded panel,
# 2024-2025 from the rendimiento supplement pack. No 12th-grade record in that year, or
# a record without promotion, counts 0. Mirrors the audited 40_final_three.R
# construction (2026-08-13).
RND <- as.data.table(readRDS(data_path("rendimiento_2024_2025.rds")))
G12 <- unique(rbind(G12, RND[COD_ENSE %in% YM & COD_GRADO == 4L,
                             .(MRUN, agno, sit = SIT_FIN_R)]))
rm(RND); invisible(gc())
stopifnot("12th-grade records must cover every on-time year 2020-2025" =
            all(2020:2025 %in% unique(G12$agno)))
GPm <- unique(G12[sit == "P", .(MRUN, agno)])
RAW[, grad_ontime := as.integer(!is.na(GPm[data.table(MRUN = RAW$MRUN, agno = RAW$proc + 4L),
                                           on = .(MRUN, agno), which = TRUE]))]
rm(G12, GPm); invisible(gc())
say("promoted out of 12th on time: winners %.4f | losers %.4f",
    RAW[won == 1L, mean(grad_ontime)], RAW[won == 0L, mean(grad_ontime)])
say("at the lottery school, years 1-4: winners %.4f %.4f %.4f %.4f | losers %.4f %.4f %.4f %.4f",
    RAW[won == 1L, mean(at_y1)], RAW[won == 1L, mean(at_y2)],
    RAW[won == 1L, mean(at_y3)], RAW[won == 1L, mean(at_y4)],
    RAW[won == 0L, mean(at_y1)], RAW[won == 0L, mean(at_y2)],
    RAW[won == 0L, mean(at_y3)], RAW[won == 0L, mean(at_y4)])

stopifnot("one row per child per lottery" = nrow(RAW) == uniqueN(RAW, by = c("key", "MRUN")),
          "every cohort present" = identical(sort(unique(RAW$proc)), 2016:2021),
          "outcomes are 0/1" = all(RAW$applied %in% 0:1) && all(RAW$assigned %in% 0:1) && all(RAW$enrolled %in% 0:1))
saveRDS(RAW, OUTF)
say("wrote %s: %s rows | %s lotteries | prioritario known %.4f | female known %.4f | %.1f min",
    OUTF, format(nrow(RAW), big.mark = ","), format(uniqueN(RAW$key), big.mark = ","),
    RAW[, mean(!is.na(prioritario))], RAW[, mean(!is.na(female))],
    (proc.time() - t0)[["elapsed"]] / 60)
```
