---
title: "Piece 2 -- Building the Lottery Panel, 7th-Grade Door"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

The 7th-grade-door companion to `00_build_lottery_panel.Rmd`: identical construction
with four parameter changes. Door grade = 7° básico. College horizon = application
year + 7 (entry + 1, then six grades to finish 12th). Application cohorts = 2016--2019,
the cohorts whose college year falls inside the outcome data (2023--2026). The
predetermined GPA is 6th grade in the application year (8th grade happens after this
door's treatment). The school list is Piece 1's 7th-grade over-demanded cells. Reads
the college outcome table built by `00_build_lottery_panel.Rmd` -- run that first.
Output: `lottery_panel_7th.rds` (22 columns): the 9th-grade panel's columns except that
the predetermined GPA is named `gpa_pre_z` (6th grade, in place of `gpa8_z`) and the
persistence columns `at_y2..at_y4`, `yrs_at_school`, `at_assigned` are not built --
the take-up columns here are `at_y1` and `in_any` only. Both panels carry
`grad_ontime` (promoted out of 12th grade in the on-time year).

```{r echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({ library(data.table) })
setDTthreads(4); options(warn = 1)
t0 <- proc.time()
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
ROOT <- file.path(home, "Dropbox", "Effect of Centralization")
OUTF <- "lottery_panel_7th.rds"
DOOR <- 7L; COHORTS <- 2016:2019; HORIZON <- 7L; GPA_GRADE <- 6L
data_path <- function(f) {
  for (p in c(file.path(home, "Dropbox", "Effect SAE on SUA", "data", f),
              file.path(ROOT, "code", "data", f),
              file.path(ROOT, "Updated Paper Spine", "Piece 1", "data", f)))
    if (file.exists(p)) return(p)
  stop(sprintf("input file not found in any data location: %s", f))
}
say <- function(...) { cat(sprintf(...), "\n"); flush.console() }

stopifnot("college outcome table missing -- run 00_build_lottery_panel.Rmd first" =
            file.exists("college_outcomes_2018_2026.rds"))
CO <- as.data.table(readRDS("college_outcomes_2018_2026.rds")); setkey(CO, mrun, proc)

EC <- as.data.table(readRDS(data_path("entrant_composition_panel.rds")))
CELLS <- unique(EC[cod_nivel == DOOR, .(rbd)]); rm(EC); invisible(gc())
say("over-demanded %d-grade schools (Piece 1 list): %s", DOOR, format(nrow(CELLS), big.mark = ","))

SAE <- readRDS(data_path("sae_admissions.rds"))
stopifnot(all(c("loteria_original", "tipo_prioridad", "es_pie", "vintage",
                "orden_pie", "orden_alta_exigencia_transicion") %in% names(SAE$c1)),
          "cod_curso_admitido" %in% names(SAE$d1),
          all(COHORTS %in% unique(SAE$b1$proc)))

build_raw <- function(pr) {
  C <- SAE$c1[proc == pr & cod_nivel == DOOR]; D <- SAE$d1[proc == pr & cod_nivel == DOOR]
  is16 <- C$vintage[1] == "2016"
  C[is.na(agregada_por_continuidad), agregada_por_continuidad := 0L]
  if (is16) {
    C[, regular := agregada_por_continuidad != 1L & tipo_prioridad %in% c(4L, 7L)]
    DA <- unique(D[!is.na(rbd_admitido), .(mrun, rbd_a = rbd_admitido)], by = "mrun")
    AR <- merge(C[, .(mrun, rbd, PREF = preferencia_postulante)],
                DA[, .(mrun, rbd = rbd_a, hit = 1L)], by = c("mrun", "rbd"), all.x = TRUE)
    ARr <- AR[hit == 1L, .(a_rank = min(PREF)), by = mrun]
    C <- merge(C, ARr, by = "mrun", all.x = TRUE); C[, a_rank := as.double(a_rank)][is.na(a_rank), a_rank := Inf]
    C <- merge(C, DA, by = "mrun", all.x = TRUE)
    C[, won := !is.na(rbd_a) & rbd == rbd_a]; C[, contested := preferencia_postulante <= a_rank]
    C[, key := paste(pr, rbd)]
  } else {
    DA <- unique(D[!is.na(rbd_admitido), .(mrun, rbd_a = rbd_admitido, curso_a = cod_curso_admitido)], by = "mrun")
    AR <- merge(C[, .(mrun, rbd, cod_curso, PREF = preferencia_postulante)],
                DA[, .(mrun, rbd = rbd_a, cod_curso = curso_a, hit = 1L)],
                by = c("mrun", "rbd", "cod_curso"), all.x = TRUE)
    ARr <- AR[hit == 1L, .(a_rank = min(PREF)), by = mrun]
    C <- merge(C, ARr, by = "mrun", all.x = TRUE); C[, a_rank := as.double(a_rank)][is.na(a_rank), a_rank := Inf]
    C[, regular := agregada_por_continuidad != 1L &
        (is.na(prioridad_matriculado) | prioridad_matriculado == 0) &
        (is.na(prioridad_hermano) | prioridad_hermano == 0) &
        (is.na(prioridad_hijo_funcionario) | prioridad_hijo_funcionario == 0) &
        (is.na(prioridad_exalumno) | prioridad_exalumno == 0) &
        (is.na(es_pie) | es_pie == 0)]
    C <- merge(C, DA[, .(mrun, rbd_a, curso_a)], by = "mrun", all.x = TRUE)
    C[, won := !is.na(rbd_a) & rbd == rbd_a & cod_curso == curso_a]
    C[, contested := preferencia_postulante <= a_rank]
    C[, key := paste(pr, rbd, cod_curso)]
  }
  M <- C[regular == TRUE & contested == TRUE & rbd %in% CELLS$rbd & !is.na(loteria_original)]
  ol <- unique(C[!is.na(rbd_a) & is.infinite(a_rank), mrun])   # placed at an off-list school
  if (length(ol)) say(paste0("  %d off-list placements: %d placed children have a placement school on ",
    "none of their application rows; kept as losses on every listed application (%d frame rows)"),
    pr, length(ol), M[mrun %in% ol, .N])
  nb <- function(x) !is.na(x) & trimws(as.character(x)) != "" & trimws(as.character(x)) != "0"
  M[, .(key, MRUN = as.integer(mrun), RBD = rbd, won = as.integer(won),
        PREF = as.integer(preferencia_postulante), loteria = as.numeric(loteria_original),
        proc = pr, vintage = fifelse(is16, "2016", "std"),
        rbd_a, tipo = tipo_prioridad,
        test_track = nb(orden_alta_exigencia_transicion),
        pie_track  = nb(orden_pie))]
}
RAW <- rbindlist(lapply(COHORTS, build_raw))
say("raw frame: %s applications | %s lotteries", format(nrow(RAW), big.mark = ","),
    format(uniqueN(RAW$key), big.mark = ","))
print(RAW[, .(apps = .N, lotteries = uniqueN(key), first_choice = sum(PREF == 1L)), by = proc][order(proc)])

q <- data.table(mrun = RAW$MRUN, proc = RAW$proc + HORIZON); o <- CO[q, on = .(mrun, proc)]
RAW[, tracked := as.integer(!is.na(o$applied))]
RAW[, `:=`(applied = fifelse(is.na(o$applied), 0L, o$applied),
           assigned = fifelse(is.na(o$assigned), 0L, o$assigned),
           enrolled = fifelse(is.na(o$enrolled), 0L, o$enrolled),
           sch_year = paste(proc, RBD))]

bdup <- SAE$b1[, .(nfl = uniqueN(paste(prioritario, es_mujer))), by = .(proc, mrun)][nfl > 1L, .N]
stopifnot("duplicate B1 rows must agree on the flags" = bdup == 0L)
B <- unique(SAE$b1[, .(proc, MRUN = as.integer(mrun), prioritario = as.integer(prioritario == 1),
                       female = as.integer(es_mujer == 1))], by = c("proc", "MRUN"))
RAW <- merge(RAW, B, by = c("proc", "MRUN"), all.x = TRUE)
rm(SAE, B, o, q, CO); invisible(gc())
stopifnot("every frame child must carry the income-tier flag" = !anyNA(RAW$prioritario),
          "every frame child must carry the gender flag" = !anyNA(RAW$female))

say("2016 pilot crosstab, processed priority code x applicant low-income flag:")
print(RAW[vintage == "2016", .N, by = .(tipo, prioritario)][order(tipo, prioritario)])
RAW[, tipo := NULL]

# predetermined achievement at this door: 6th-grade GPA in the application year
PF <- as.data.table(readRDS(data_path("panel_performance_2002_2023.rds")))
G6 <- PF[COD_GRADO == GPA_GRADE & COD_ENSE == 110L & AGNO %in% COHORTS &
         !is.na(PROM_GRAL) & PROM_GRAL >= 1 & PROM_GRAL <= 7, .(MRUN, AGNO, PROM_GRAL)]
YM <- c(310L, 410L, 510L, 610L, 710L, 810L, 910L)     # 12th-grade school types (media)
G12 <- unique(PF[COD_ENSE %in% YM & COD_GRADO == 4L & AGNO %in% (COHORTS + HORIZON - 1L) &
                   AGNO <= 2023L, .(MRUN, agno = AGNO, sit = SIT_FIN_R)])
rm(PF); invisible(gc())
G6 <- G6[, .(gpa = mean(PROM_GRAL)), by = .(MRUN, AGNO)]
G6[, gpa_pre_z := (gpa - mean(gpa)) / sd(gpa), by = AGNO]
RAW[, gpa_pre_z := G6[RAW, on = .(MRUN, AGNO = proc), x.gpa_pre_z]]
rm(G6); invisible(gc())
say("6th-grade GPA matched: winners %.4f | losers %.4f",
    RAW[won == 1L, mean(!is.na(gpa_pre_z))], RAW[won == 0L, mean(!is.na(gpa_pre_z))])

# take-up: entry-year enrolment (application + 1) at the lottery school and anywhere
MATC <- readRDS(data_path("matricula_2011_2023.rds"))
ENT <- unique(MATC[agno %in% (COHORTS + 1L), .(agno, MRUN, RBD)])
rm(MATC); invisible(gc())
qk <- data.table(MRUN = RAW$MRUN, RBD = RAW$RBD, agno = RAW$proc + 1L)
RAW[, at_y1 := as.integer(!is.na(ENT[qk, on = .(MRUN, RBD, agno), which = TRUE]))]
E2 <- unique(ENT[, .(MRUN, agno)])
RAW[, in_any := as.integer(!is.na(E2[data.table(MRUN = RAW$MRUN, agno = RAW$proc + 1L),
                                     on = .(MRUN, agno), which = TRUE]))]
rm(ENT, E2, qk); invisible(gc())
say("take-up at entry: at lottery school winners %.4f losers %.4f | any school %.4f / %.4f",
    RAW[won == 1L, mean(at_y1)], RAW[won == 0L, mean(at_y1)],
    RAW[won == 1L, mean(in_any)], RAW[won == 0L, mean(in_any)])

# graduation on time: promoted out of 12th grade (final status P) in the on-time year
# (application + HORIZON - 1, i.e. + 6 at this door). 12th-grade records through 2023
# from the canonical graded panel, 2024-2025 from the rendimiento supplement pack.
# No 12th-grade record in that year, or a record without promotion, counts 0.
RND <- as.data.table(readRDS(data_path("rendimiento_2024_2025.rds")))
G12 <- unique(rbind(G12, RND[COD_ENSE %in% YM & COD_GRADO == 4L,
                             .(MRUN, agno, sit = SIT_FIN_R)]))
rm(RND); invisible(gc())
stopifnot("12th-grade records must cover every on-time year at this door" =
            all((COHORTS + HORIZON - 1L) %in% unique(G12$agno)))
GPm <- unique(G12[sit == "P", .(MRUN, agno)])
RAW[, grad_ontime := as.integer(!is.na(GPm[data.table(MRUN = RAW$MRUN, agno = RAW$proc + HORIZON - 1L),
                                           on = .(MRUN, agno), which = TRUE]))]
rm(G12, GPm); invisible(gc())
say("promoted out of 12th on time: winners %.4f | losers %.4f",
    RAW[won == 1L, mean(grad_ontime)], RAW[won == 0L, mean(grad_ontime)])

stopifnot("one row per child per lottery" = nrow(RAW) == uniqueN(RAW, by = c("key", "MRUN")),
          "every cohort present" = identical(sort(unique(RAW$proc)), COHORTS),
          "outcomes are 0/1" = all(RAW$applied %in% 0:1) && all(RAW$assigned %in% 0:1) && all(RAW$enrolled %in% 0:1))
saveRDS(RAW, OUTF)
say("wrote %s: %s rows | %s lotteries | prioritario known %.4f | %.1f min",
    OUTF, format(nrow(RAW), big.mark = ","), format(uniqueN(RAW$key), big.mark = ","),
    RAW[, mean(!is.na(prioritario))], (proc.time() - t0)[["elapsed"]] / 60)
```
