---
title: "Piece 2 -- Effect of Winning an SAE Lottery on College Outcomes"
output:
  pdf_document:
    keep_tex: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Clean ---------------------------------------------------------------
rm(list = ls())

# Libraries ---------------------------------------------------------------
library(data.table)
library(fixest)

# RNG contract: every results chunk re-seeds 20260919 at its own top, so each exhibit's
# random draws do not depend on which chunks ran before it. Chunks share the frame
# objects (P, As, A) built upstream, so run the document top to bottom. The RI p-values
# of record are the ones this file writes; earlier scripts consumed the stream in a
# different order.
set.seed(20260919)
B_RI <- 9999L

# The panel (built by 00_build_lottery_panel.Rmd). A copy next to this file wins;
# otherwise the shared data folder's copy of record.
home <- ifelse(Sys.getenv("USERPROFILE") != "", Sys.getenv("USERPROFILE"), path.expand("~"))
panel_file <- "lottery_panel.rds"
if (!file.exists(panel_file))
  panel_file <- file.path(home, "Dropbox", "Effect SAE on SUA", "data", "lottery_panel.rds")
stopifnot("panel not found next to the code or in the shared data folder" =
            file.exists(panel_file))
dir.create("tables", showWarnings = FALSE)

fmt  <- function(x) sprintf("%.4f", x)
star <- function(p) ifelse(p < .01, "$^{***}$",
                    ifelse(p < .05, "$^{**}$",
                    ifelse(p < .1,  "$^{*}$", "")))
row_tex <- function(lbl, vals) paste0("      ", lbl, " & ", paste(vals, collapse = " & "), "\\\\\n")

# the primary frame: first-choice applications, lotteries with >= W winners, loser
# violations <= pur, nearest losers capped at the winner count
make_frame <- function(P, W, pur, cap, fc) {
  M <- copy(P); if (fc) M <- M[PREF == 1L]
  CU <- M[won == 1L, .(cutoff = max(loteria), n_win = .N), by = key]; M <- merge(M, CU, by = "key")[n_win >= W]
  M[, viol := won == 0L & loteria < cutoff]
  PU <- M[won == 0L, .(n_los = .N, n_viol = sum(viol)), by = key]
  M <- merge(M, PU, by = "key")[n_los >= 1L][n_viol / n_los <= pur]
  L <- M[won == 0L & loteria >= cutoff]
  if (cap) { setorder(L, key, loteria); L[, rk := seq_len(.N), by = key]; L <- L[rk <= n_win] }
  rbind(M[won == 1L], L, fill = TRUE)
}
# within-lottery ITT, precision-weighted across lotteries; the closed form is asserted
# equal to the lottery-fixed-effects regression at every run; stars from reshuffling
# winners within each lottery (winner count held fixed)
core <- function(DT, y) {
  DT <- DT[!is.na(get(y))]; setorder(DT, key)
  KX <- DT[, .(i0 = .I[1], n = .N, m = sum(won)), by = key][m > 0L & m < n]
  DT <- DT[key %in% KX$key]; setorder(DT, key)
  KX <- DT[, .(i0 = .I[1], n = .N, m = sum(won)), by = key]
  denom <- KX[, sum(m * (n - m) / n)]; yv <- as.numeric(DT[[y]])
  base <- DT[, .(b = mean(get(y)) * sum(won)), by = key]$b
  obs <- (sum(yv * DT$won) - sum(base)) / denom
  pn <- numeric(B_RI)
  for (r in seq_len(nrow(KX))) { i0 <- KX$i0[r]; n <- KX$n[r]; m <- KX$m[r]
    yk <- yv[i0:(i0 + n - 1L)]; pn <- pn + .colSums(yk[replicate(B_RI, sample.int(n, m))], m, B_RI) }
  perm <- (pn - sum(base)) / denom
  fit <- feols(as.formula(paste0(y, " ~ won | key")), data = DT, cluster = ~sch_year, fixef.rm = "none")
  ct <- summary(fit)$coeftable
  stopifnot("closed form must equal the regression point" = abs(unname(ct["won", 1]) - obs) < 1e-8)
  data.table(itt = obs, se = ct["won", 2],
             ri_p = (1 + sum(abs(perm) >= abs(obs) - 1e-12)) / (B_RI + 1),
             n = nrow(DT), k = uniqueN(DT$key),
             lm = DT[won == 0L, mean(get(y))], win_mean = DT[won == 1L, mean(get(y))])
}
```

## Design

Inside an over-demanded school, SAE runs a lottery among the ordinary applicants. SAE
reserves 15% of each school's seats for low-income (prioritario) children, so one
classroom really holds two draws: one among the low-income applicants, one among the
rest, each with its own effective cutoff. The comparison therefore happens **within
tier**: the unit is classroom by year by income tier, and inside that unit every child
faced the same doors and the same odds, so winning is a coin flip by construction (the
marginal-priority-group design of Abdulkadiroglu, Angrist, Narita and Pathak 2017; one
school per stratum in the 2016 pilot, which predates classroom identifiers).
The panel (one row per child per application, built and documented in
`00_build_lottery_panel.Rmd`) supplies the tier flag for every child. The question of
record: **does winning a contested seat reduce on-time entry into the college
system?** Outcomes are four yes/no facts on the cohort's own schedule: promoted out of
12th grade in the on-time year (application + 4, from the school records), and -- five
years after the application, when the cohort reaches the college admission process --
applied, assigned a seat, enrolled. A child with no record carries zeros. The on-time
year is the one point every cohort reaches inside the data (the college records end at
process 2026, which is the 2021 cohort's on-time year), so every stratum is asked the
identical question. Each stratum keeps at least five winners, a loser
pool consistent with its winners' cutoff (up to 10% violations), and the nearest losers
capped at the winner count -- all computed within the stratum. Standard errors cluster
on school by year; stars come from randomization inference, reshuffling the winners
within each stratum 9,999 times with the winner count held fixed.

## Headline

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
P <- as.data.table(readRDS(panel_file))
stopifnot("every child must carry the income-tier flag" = P[, all(!is.na(prioritario))])
Ps <- copy(P)[, key := paste(key, prioritario)]   # stratum = classroom x year x tier
As <- make_frame(Ps, 5L, 0.10, TRUE, TRUE)
stopifnot("no test-track winners may sit in the frame" = As[won == 1L & test_track == TRUE, .N] == 0L)

# single-cutoff diagnostics. Within a tier every child faces one effective cutoff, so a
# loser with a strictly better number than the worst same-tier winner should not exist;
# the violation mass is the observable test of that claim, printed before it is filtered.
Mf <- copy(Ps)[PREF == 1L]
CU <- Mf[won == 1L, .(cutoff = max(loteria), n_win = .N), by = key]
Mf <- merge(Mf, CU, by = "key")
f1 <- uniqueN(Mf$key); Mf <- Mf[n_win >= 5L]; f2 <- uniqueN(Mf$key)
Mf[, viol := won == 0L & loteria < cutoff]
PU <- Mf[won == 0L, .(n_los = .N, n_viol = sum(viol)), by = key]
Mf <- merge(Mf, PU, by = "key")[n_los >= 1L]; f3 <- uniqueN(Mf$key)
keep <- Mf[n_viol / n_los <= 0.10]; f4 <- uniqueN(keep$key)
cat(sprintf(paste0("Stratum funnel: %d with a winner; %d with at least 5 winners; %d also ",
                   "with a loser; %d pass the 10%% violation screen.\n\n"), f1, f2, f3, f4))
vd <- unique(keep[, .(key, vintage, n_viol, n_los)])
vs <- vd[, .(strata = .N, with_any_violation = sum(n_viol > 0),
             violating_losers = sum(n_viol), losers = sum(n_los)), by = vintage]
for (i in seq_len(nrow(vs))) cat(sprintf(
  "Vintage %s: %d surviving strata, %d contain a violation; %d violating losers of %d (share %.4f).\n\n",
  vs$vintage[i], vs$strata[i], vs$with_any_violation[i], vs$violating_losers[i],
  vs$losers[i], vs$violating_losers[i] / vs$losers[i]))
rm(Mf, CU, PU, keep, vd, vs)

res <- rbindlist(lapply(c("grad_ontime", "applied", "assigned", "enrolled"),
                        function(y) core(As, y)[, outcome := y]))
fwrite(res, "tables/lottery_headline_values.csv")

tab <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Effect of Winning an SAE Lottery on On-Time Outcomes}",
  "\\label{tab: lottery headline}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lcccc}\n      \\toprule\n",
  "                   & Grad.\\ on time & Applied & Assigned & Enrolled\\\\\n",
  "                   & (1) & (2) & (3) & (4)\\\\\n      \\midrule\n",
  row_tex("Won the lottery", paste0(fmt(res$itt), star(res$ri_p))),
  row_tex("", paste0("(", fmt(res$se), ")")),
  "      \\midrule\n",
  row_tex("Loser-arm mean", fmt(res$lm)),
  row_tex("Observations", format(res$n, big.mark = ",")),
  row_tex("Strata (classroom x year x tier)", format(res$k, big.mark = ",")),
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: within-stratum comparison of winners and ",
  "nearest losers; a stratum is one classroom (one school in the 2016 pilot) in one ",
  "application year within one income tier, so every compared child faced the same ",
  "admission doors. First-choice applications, cohorts 2016--2021. Grad.\\ on time = ",
  "promoted out of 12th grade in the on-time year (application + 4, school records); ",
  "college outcomes at application year + 5; no record = 0 throughout. School-by-year ",
  "clustered standard errors in parentheses. Stars from ",
  "randomization inference (9,999 within-stratum reshuffles): $^{***}$p$<$0.01, ",
  "$^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab, "tables/lottery_headline.tex")
cat(tab)
cat(sprintf("\nHeadline frame: %s applications, %s strata.\n",
            format(nrow(As), big.mark = ","), format(uniqueN(As$key), big.mark = ",")))
```

## The treatment happened

Two facts from the enrolment census establish that the draw did what the headline
assumes. First, winners and losers attend different schools: in the entry year, winners
are at the lottery school and losers are overwhelmingly elsewhere. Second, both arms are
enrolled in some school, so the headline compares attending this school against
attending other schools -- with the difference in schools caused by the random draw and
nothing else.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
tk <- rbind(
  core(As, "at_y1")[, m := "at the lottery school, entry year"],
  core(As, "in_any")[, m := "enrolled at any school, entry year"], fill = TRUE)
fwrite(tk, "tables/lottery_takeup_values.csv")
for (i in seq_len(nrow(tk))) { b <- tk[i]
  cat(sprintf("%s: winners %.4f, losers %.4f, difference %+.4f%s (%.4f), RI p %.4f.\n\n",
              b$m, b$win_mean, b$lm, b$itt, gsub("[$^{}]", "", star(b$ri_p)), b$se, b$ri_p)) }
```

## Balance

On the pooled frame (one classroom = one lottery, tiers mixed), is winning related to
being low-income or female? A clean draw puts both differences at zero; this is the
direct measurement of whether the quota tilts the winner group.

Inside a stratum the tier is constant by construction, so the stratified frame's own
balance tests use the predetermined traits that still vary within stratum: gender, and
the strongest one the data offers, the child's 8th-grade GPA in the application year.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
A <- make_frame(P, 5L, 0.10, TRUE, TRUE)   # pooled: tiers mixed inside each lottery
stopifnot("no test-track winners may sit in the pooled frame" = A[won == 1L & test_track == TRUE, .N] == 0L)
bal <- rbindlist(lapply(c("prioritario", "female"), function(y) core(A, y)[, covar := y]))
bs  <- core(As, "female")[, covar := "female (within stratum)"]
bg8 <- core(As[!is.na(gpa8_z)], "gpa8_z")[, covar := "8th-grade GPA (z, within stratum)"]
bal <- rbind(bal, bs, bg8, fill = TRUE)
fwrite(bal, "tables/lottery_balance_values.csv")
cat(sprintf("8th-grade GPA matched: winners %.4f, losers %.4f.\n\n",
            As[won == 1L, mean(!is.na(gpa8_z))], As[won == 0L, mean(!is.na(gpa8_z))]))
for (i in seq_len(nrow(bal))) { b <- bal[i]
  cat(sprintf("%s: winner-arm mean %.4f, loser-arm mean %.4f, difference %+.4f%s (%.4f), RI p %.4f.\n\n",
              b$covar, b$win_mean, b$lm, b$itt, gsub("[$^{}]", "", star(b$ri_p)), b$se, b$ri_p)) }
```

## Robustness

Five checks, one change each, all against the stratified headline. *Pooled*: one
classroom = one lottery with tiers mixed, the simpler frame the stratified design
refines. *No-priority tier only* and *Low-income tier only*: the headline frame
restricted to one tier at a time, so each tier's answer is visible on its own. *Drop
special-admission strata*: some schools admit part of the class through a
special-education ordering the fair-coin-flip filter cannot see; a child admitted that
way who also drew a lottery number would sit in the winner group without having won a
coin flip -- every stratum containing such a winner is dropped. *Drop same-school
losers*: a child coded as a loser at their first-choice classroom can have been placed
in a different classroom of the same school; those children are dropped. The 2016 pilot
files carry no admission-ordering columns; the pilot's screen is its priority-code
filter, applied in the build.

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
set.seed(20260919)
pk <- As[won == 1L & pie_track == TRUE, unique(key)]
cat(sprintf("Special-education-ordering winners: %d, in %d strata. Same-school losers: %d.\n\n",
            As[won == 1L & pie_track == TRUE, .N], length(pk),
            As[won == 0L & !is.na(rbd_a) & rbd_a == RBD, .N]))
specs <- list(
  list(lab = "Pooled (tiers mixed)",             D = A),
  list(lab = "No-priority tier only",            D = As[prioritario == 0L]),
  list(lab = "Low-income tier only",             D = As[prioritario == 1L]),
  list(lab = "Drop special-admission strata",    D = As[!key %in% pk]),
  list(lab = "Drop same-school losers",          D = As[!(won == 0L & !is.na(rbd_a) & rbd_a == RBD)]))
rows <- ""; rvals <- list()
for (s in specs) {
  r <- rbindlist(lapply(c("grad_ontime", "applied", "assigned", "enrolled"),
                        function(y) core(s$D, y)[, outcome := y]))
  rvals[[s$lab]] <- copy(r)[, check := s$lab]
  rows <- paste0(rows,
    row_tex(s$lab, paste0(fmt(r$itt), star(r$ri_p))),
    row_tex("", paste0("(", fmt(r$se), ")")),
    row_tex("\\quad Observations / strata",
            c(paste0(format(r$n[1], big.mark = ","), " / ", format(r$k[1], big.mark = ",")), "", "", "")))
}
fwrite(rbindlist(rvals), "tables/lottery_robustness_values.csv")
tab_r <- paste0(
  "\\begin{table}[htbp]\n   \\caption{Robustness: Frame and Arm-Composition Checks}",
  "\\label{tab: lottery robustness}\n",
  "   \\centerline{\\scalebox{0.85}{\\begin{tabular}{lcccc}\n      \\toprule\n",
  "                   & Grad.\\ on time & Applied & Assigned & Enrolled\\\\\n",
  "                   & (1) & (2) & (3) & (4)\\\\\n      \\midrule\n",
  rows,
  "      \\bottomrule\n   \\end{tabular}}}\n",
  "   \\par\\smallskip{\\footnotesize Notes: same estimator, clustering, and ",
  "randomization inference as the headline; outcomes as defined there; one row block ",
  "per check. The pooled row ",
  "counts lotteries; all other rows count strata. ",
  "$^{***}$p$<$0.01, $^{**}$p$<$0.05, $^{*}$p$<$0.1.}\n\\end{table}\n")
writeLines(tab_r, "tables/lottery_robustness.tex")
cat(tab_r)
```
